---
title: "Two Donor scRNAseq"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here('..','html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(tidyr)
library(Seurat)
library(patchwork)
library(cluster)
library(parallelDist)
library(ggplot2)
library(data.table)
library(cowplot)
library(future)
library(SeuratDisk)


reload_data=F
source(here::here('r','fig_colors.R'))
source(here::here('r','scrna','HTOdemux.R'))
source(here::here('r','scrna','scrna_analysis_fxns.R'))

options(future.globals.maxSize = 14000 * 1024^2)
plan("sequential")
plan()
```

## Load all four wells individually

```{r}

#rename some adt features
hto_rows <- paste('Hashtag',1:12,sep='_')

adt_feature_names <- fread(here::here(
    '..','..','scrnaseq_tcsl154',
    'meta','adt_names.tsv'))

make_sample_seurat_obj <- function(expt_num, well) {

  scrna.data <- Read10X(data.dir = here::here(
    '..', '..', paste0('scrnaseq_tcsl',expt_num),
    'cellranger', paste('TCSL', expt_num, well, sep='_'), 'outs','raw_feature_bc_matrix'))

  rownames(scrna.data$`Antibody Capture`) <- c(paste0(adt_feature_names$vis_name,'.adt'), hto_rows)
  
  #rna features
  scrna <- CreateSeuratObject(
      counts = scrna.data$`Gene Expression`, 
      project = "tcsl154", 
      min.cells = 3,
      min.features = 200)
  
  adt_rows <- setdiff(rownames(scrna.data$`Antibody Capture`), hto_rows)
  
  #adt features
  scrna[['ADT']] <- CreateAssayObject(
      counts = scrna.data$`Antibody Capture`[adt_rows, colnames(scrna)])
  
  #hto features
  scrna[['HTO']] <- CreateAssayObject(
    counts = scrna.data$`Antibody Capture`[hto_rows, colnames(scrna)])
  
  scrna@meta.data$sample <- paste(expt_num, well, sep='_')
  scrna@meta.data$donor <- expt_num
  scrna@meta.data$well <- well
  
  return(scrna)
}

scrna.list <- list(
  make_sample_seurat_obj(154,1),
  make_sample_seurat_obj(154,2),
  make_sample_seurat_obj(170,1),
  make_sample_seurat_obj(170,2)
)
```



## Remove MT-high cells and individually perform SCTransform on all 3 assay types.

```{r}
mt_rna_pct_cutoff <- 25
read_cutoff <- 1e3

run_sct <- function(seurat_obj, mt_rna_pct_cutoff, read_cutoff) {
  
  # remove high MT-RNA cells (dying/dead?)
  DefaultAssay(seurat_obj) <- 'RNA'
  seurat_obj <- PercentageFeatureSet(seurat_obj, pattern = "^MT-", col.name = "percent.mt")
  seurat_obj <- subset(seurat_obj, subset=(percent.mt < mt_rna_pct_cutoff & nCount_RNA > read_cutoff))
  
  # get cell cycle scores
  DefaultAssay(seurat_obj) <- 'RNA'
  s.genes <- cc.genes$s.genes
  g2m.genes <- cc.genes$g2m.genes
  #https://satijalab.org/seurat/v3.1/cell_cycle_vignette.html
  seurat_obj <- CellCycleScoring(seurat_obj, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
  #seurat_obj <- RunPCA(seurat_obj, features = c(s.genes, g2m.genes), reduction.name = 'cc_pca')
  message('  Cell Cycle Scoring Complete.')
  
  #SCTransform for RNA
  DefaultAssay(seurat_obj) <- 'RNA'
  seurat_obj <- SCTransform(seurat_obj, 
    vars.to.regress = c("S.Score", "G2M.Score"), verbose = FALSE)
  seurat_obj <- RunPCA(seurat_obj, reduction.name = 'rna_pca')
  message('  SCTransform RNA complete.')
  
  #SCTransform for ADT
  DefaultAssay(seurat_obj) <- 'ADT'
  seurat_obj <- SCTransform(
    seurat_obj, assay='ADT', new.assay.name = 'SCT_ADT', 
    vars.to.regress = c("S.Score", "G2M.Score"), verbose = FALSE)
  seurat_obj <- RunPCA(
    seurat_obj, features = rownames(seurat_obj[["SCT_ADT"]]), 
    nfeatures.print = 10, reduction.name='adt_pca')
  message('  SCTransform ADT complete.')
  
  #SCTransform for HTO
  DefaultAssay(seurat_obj) <- 'HTO'
  seurat_obj <- SCTransform(seurat_obj, assay='HTO', new.assay.name = 'SCT_HTO')
  message('  SCTransform HTO complete.')
  
  return(seurat_obj)
}

for (i in 1:4) {
  message(paste('Starting Sample',i))
  scrna.list[[i]] <- run_sct(scrna.list[[i]], mt_rna_pct_cutoff, read_cutoff)
   message('  Sample completed.')
}
```

```{r}


combined_metadata <- rbindlist(lapply(1:4, function(i) data.table(scrna.list[[i]]@meta.data)))
mt_summary <- combined_metadata[,
  mt_thresh := percent.mt < mt_rna_pct_cutoff & nCount_RNA > read_cutoff][,
  list(
    neg= .N-sum(mt_thresh),
    pos=sum(mt_thresh),
    pct_pos=round(sum(mt_thresh)/.N, 2)), by=c('donor')]

plots <- list()

plots$percent.mt <- ggplot(combined_metadata) + 
  aes(x=nCount_RNA, y=percent.mt) + 
  scale_x_log10() + 
  stat_density2d(geom="tile", aes(fill=..density.., alpha=1), contour=FALSE) + 
  geom_point(size=0.5, color=blues9[9], alpha=0.25) +
  stat_density2d(geom="tile", aes(fill=..density..,     alpha=ifelse(..density..<0.01,0,1)), contour=FALSE) + 
  scale_fill_gradientn(colours = colorRampPalette(c("white", blues9))(256)) +
  facet_grid(~donor) +
  geom_hline(aes(yintercept=mt_rna_pct_cutoff)) +
  geom_vline(aes(xintercept=read_cutoff)) +
  geom_text(data=mt_summary, aes(x=10^4,y=100, label=paste(pct_pos,":",pos,'kept,',neg,'removed'))) +
  theme_bw()
```

Percent MT-RNA is higher for 170, and more cells are lost. May want to consider a higher % cutoff for that sample, but
this is already quite permissive according to some literature (10% is considered optimal by Osorio et al. 2020).

### HTO calling for each individual well

Getting unclear error messages with HTO SCT integration, so calling cells separately for each well, then combining
via metadata.

```{r}

HTO_sample_metadata <- fread(here::here(
    '..','..','scrnaseq_tcsl154','meta','sample_metadata.csv'))

for (i in 1:4) {
  DefaultAssay(scrna.list[[i]]) <- 'SCT_HTO'
  
  if (!('hto.tsne' %in% names(scrna.list[[i]]@reductions))) {
    scrna.list[[i]] <- RunTSNE(
      scrna.list[[i]],
      perplexity=100, dims=NULL,
      features=paste0('Hashtag-',1:12),
      reduction.name='hto.tsne', tsne.method='FIt-SNE', nthreads=12)
  }
  
  donor_i <- ifelse(i < 3, 'TCSL154','TCSL170')
  sample_metadata <- HTO_sample_metadata[donor == donor_i]
  sample_metadata[,  sample := paste(car, ifelse(k_type=='baseline', 'unstim','stim'), sep='.')]
  sample_metadata[k_type=='bead_stim',  sample := paste(car, k_type, sep='.')]
  
  scrna.list[[i]]@meta.data$well <- scrna.list[[i]]@meta.data$sample
  scrna.list[[i]]@meta.data$sample <- NULL
  
  scrna.list[[i]] <- assign_HTO_sample(
    scrna.list[[i]], metadata=sample_metadata,
    assay='SCT_HTO', slot='data', save_plots=T)

}

plots$well_tsne_maps <- plot_grid(
  plotlist=unlist(
    lapply(1:4, function (i) list(
      DimPlot(scrna.list[[i]], group.by='sample', reduction='hto.tsne'),
      DimPlot(scrna.list[[i]], group.by='HTO_status', reduction='hto.tsne'))), 
    recursive=F), ncol=2)

plots$well_tsne_maps_pruned <- plot_grid(
  plotlist=unlist(
    lapply(1:4, function (i) {
      scrna_subset <- subset(scrna.list[[i]], subset = (HTO_status %in% c('HighQual','MidQual')))
      list(
      DimPlot(scrna_subset, group.by='sample', reduction='hto.tsne'),
      DimPlot(scrna_subset, group.by='HTO_status', reduction='hto.tsne'))
    }), 
    recursive=F), 
  ncol=2)

initial.rds.file <- here::here(
     '..','..','scrnaseq_tcsl154',
     'rds',
     paste0('combined.initial.data.rds'))

initial.plots.rds.file <- here::here(
     '..','..','scrnaseq_tcsl154',
     'rds',
     paste0('01_loading_plots.rds'))

saveRDS(scrna.list,initial.rds.file)
saveRDS(plots, initial.rds.file)
```
