---
title: "Two Donor scRNAseq"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here('..','html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(tidyr)
library(Seurat)
library(patchwork)
library(cluster)
library(parallelDist)
library(ggplot2)
library(data.table)
library(cowplot)
library(future)
library(SeuratDisk)


reload_data=F
source(here::here('r','fig_colors.R'))
source(here::here('r','scrna','HTOdemux.R'))
source(here::here('r','scrna','scrna_analysis_fxns.R'))

options(future.globals.maxSize = 14000 * 1024^2)
plan("sequential")
plan()
```

## Load all four wells individually

```{r}

#rename some adt features
hto_rows <- paste('Hashtag',1:12,sep='_')

adt_feature_names <- fread(here::here(
    '..','..','scrnaseq_tcsl154',
    'meta','adt_names.tsv'))

make_sample_seurat_obj <- function(expt_num, well) {

  scrna.data <- Read10X(data.dir = here::here(
    '..', '..', paste0('scrnaseq_tcsl',expt_num),
    'cellranger', paste('TCSL', expt_num, well, sep='_'), 'outs','raw_feature_bc_matrix'))

  rownames(scrna.data$`Antibody Capture`) <- c(paste0(adt_feature_names$vis_name,'.adt'), hto_rows)
  
  #rna features
  scrna <- CreateSeuratObject(
      counts = scrna.data$`Gene Expression`, 
      project = "tcsl154", 
      min.cells = 3,
      min.features = 200)
  
  adt_rows <- setdiff(rownames(scrna.data$`Antibody Capture`), hto_rows)
  
  #adt features
  scrna[['ADT']] <- CreateAssayObject(
      counts = scrna.data$`Antibody Capture`[adt_rows, colnames(scrna)])
  
  #hto features
  scrna[['HTO']] <- CreateAssayObject(
    counts = scrna.data$`Antibody Capture`[hto_rows, colnames(scrna)])
  
  scrna@meta.data$sample <- paste(expt_num, well, sep='_')
  scrna@meta.data$donor <- expt_num
  scrna@meta.data$well <- well
  
  return(scrna)
}

scrna.list <- list(
  make_sample_seurat_obj(154,1),
  make_sample_seurat_obj(154,2),
  make_sample_seurat_obj(170,1),
  make_sample_seurat_obj(170,2)
)
```



## Remove MT-high cells and individually perform SCTransform on all 3 assay types.

```{r}
mt_rna_pct_cutoff <- 25
read_cutoff <- 1e3

run_sct <- function(seurat_obj, mt_rna_pct_cutoff, read_cutoff) {
  
  # remove high MT-RNA cells (dying/dead?)
  DefaultAssay(seurat_obj) <- 'RNA'
  seurat_obj <- PercentageFeatureSet(seurat_obj, pattern = "^MT-", col.name = "percent.mt")
  seurat_obj <- subset(seurat_obj, subset=(percent.mt < mt_rna_pct_cutoff & nCount_RNA > read_cutoff))
  
  # get cell cycle scores
  DefaultAssay(seurat_obj) <- 'RNA'
  s.genes <- cc.genes$s.genes
  g2m.genes <- cc.genes$g2m.genes
  #https://satijalab.org/seurat/v3.1/cell_cycle_vignette.html
  seurat_obj <- CellCycleScoring(seurat_obj, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
  #seurat_obj <- RunPCA(seurat_obj, features = c(s.genes, g2m.genes), reduction.name = 'cc_pca')
  message('  Cell Cycle Scoring Complete.')
  
  #SCTransform for RNA
  DefaultAssay(seurat_obj) <- 'RNA'
  seurat_obj <- SCTransform(seurat_obj, 
    vars.to.regress = c("S.Score", "G2M.Score"), verbose = FALSE)
  seurat_obj <- RunPCA(seurat_obj, reduction.name = 'rna_pca')
  message('  SCTransform RNA complete.')
  
  #SCTransform for ADT
  DefaultAssay(seurat_obj) <- 'ADT'
  seurat_obj <- SCTransform(
    seurat_obj, assay='ADT', new.assay.name = 'SCT_ADT', 
    vars.to.regress = c("S.Score", "G2M.Score"), verbose = FALSE)
  seurat_obj <- RunPCA(
    seurat_obj, features = rownames(seurat_obj[["SCT_ADT"]]), 
    nfeatures.print = 10, reduction.name='adt_pca')
  message('  SCTransform ADT complete.')
  
  #SCTransform for HTO
  DefaultAssay(seurat_obj) <- 'HTO'
  seurat_obj <- SCTransform(seurat_obj, assay='HTO', new.assay.name = 'SCT_HTO')
  message('  SCTransform HTO complete.')
  
  return(seurat_obj)
}

for (i in 1:4) {
  message(paste('Starting Sample',i))
  scrna.list[[i]] <- run_sct(scrna.list[[i]], mt_rna_pct_cutoff, read_cutoff)
   message('  Sample completed.')
}

initial.rds.file <- here::here(
     '..','..','scrnaseq_tcsl154',
     'rds',
     paste0('combined.initial.data.rds'))

#if (!file.exists(initial.rds.file)) {
saveRDS(scrna.list,initial.rds.file)

```

```{r}


combined_metadata <- rbindlist(lapply(1:4, function(i) data.table(scrna.list[[i]]@meta.data)))
mt_summary <- combined_metadata[,
  mt_thresh := percent.mt < mt_rna_pct_cutoff & nCount_RNA > read_cutoff][,
  list(
    neg= .N-sum(mt_thresh),
    pos=sum(mt_thresh),
    pct_pos=round(sum(mt_thresh)/.N, 2)), by=c('donor')]

ggplot(combined_metadata) + 
  aes(x=nCount_RNA, y=percent.mt) + 
  scale_x_log10() + 
  stat_density2d(geom="tile", aes(fill=..density.., alpha=1), contour=FALSE) + 
  geom_point(size=0.5, color=blues9[9], alpha=0.25) +
  stat_density2d(geom="tile", aes(fill=..density..,     alpha=ifelse(..density..<0.01,0,1)), contour=FALSE) + 
  scale_fill_gradientn(colours = colorRampPalette(c("white", blues9))(256)) +
  facet_grid(~donor) +
  geom_hline(aes(yintercept=mt_rna_pct_cutoff)) +
  geom_vline(aes(xintercept=read_cutoff)) +
  geom_text(data=mt_summary, aes(x=10^4,y=100, label=paste(pct_pos,":",pos,'kept,',neg,'removed'))) +
  theme_bw()
```

Percent MT-RNA is higher for 170, and more cells are lost. May want to consider a higher % cutoff for that sample, but
this is already quite permissive according to some literature (10% is considered optimal by Osorio et al. 2020).

### HTO calling for each individual well

Getting unclear error messages with HTO SCT integration, so calling cells separately for each well, then combining
via metadata.

```{r}

HTO_sample_metadata <- fread(here::here(
    '..','..','scrnaseq_tcsl154','meta','sample_metadata.csv'))

for (i in 1:4) {
  DefaultAssay(scrna.list[[i]]) <- 'SCT_HTO'
  
  if (!('hto.tsne' %in% names(scrna.list[[i]]@reductions))) {
    scrna.list[[i]] <- RunTSNE(
      scrna.list[[i]],
      perplexity=100, dims=NULL,
      features=paste0('Hashtag-',1:12),
      reduction.name='hto.tsne', tsne.method='FIt-SNE', nthreads=12)
  }
  
  donor_i <- ifelse(i < 3, 'TCSL154','TCSL170')
  sample_metadata <- HTO_sample_metadata[donor == donor_i]
  sample_metadata[,  sample := paste(car, ifelse(k_type=='baseline', 'unstim','stim'), sep='.')]
  sample_metadata[k_type=='bead_stim',  sample := paste(car, k_type, sep='.')]
  
  scrna.list[[i]]@meta.data$well <- scrna.list[[i]]@meta.data$sample
  scrna.list[[i]]@meta.data$sample <- NULL
  
  scrna.list[[i]] <- assign_HTO_sample(
    scrna.list[[i]], metadata=sample_metadata,
    assay='SCT_HTO', slot='data', save_plots=T)

}

plots <- list()

plots$well_tsne_maps <- plot_grid(
  plotlist=unlist(
    lapply(1:4, function (i) list(
      DimPlot(scrna.list[[i]], group.by='sample', reduction='hto.tsne'),
      DimPlot(scrna.list[[i]], group.by='HTO_status', reduction='hto.tsne'))), 
    recursive=F), ncol=2)

plots$well_tsne_maps_pruned <- plot_grid(
  plotlist=unlist(
    lapply(1:4, function (i) {
      scrna_subset <- subset(scrna.list[[i]], subset = (HTO_status %in% c('HighQual','MidQual')))
      list(
      DimPlot(scrna_subset, group.by='sample', reduction='hto.tsne'),
      DimPlot(scrna_subset, group.by='HTO_status', reduction='hto.tsne'))
    }), 
    recursive=F), 
  ncol=2)

```

### Integrate SCT RNA

```{r}

integrate_feature <- function(scrna.list, assay_name, initial_assay, assay_only=F, remove_other_assays=F, use_ref=F) {
  
  integrated_assay_name <- paste0(assay_name, "_INT")
  
  # only one assay
  if (remove_other_assays) {
    scrna.assay.list <- lapply(
      1:length(scrna.list),
      function(i) {
        obj <- CreateSeuratObject(
          scrna.list[[i]][[assay_name]],
          assay=assay_name,
          meta.data=scrna.list[[i]]@meta.data)
        obj[[initial_assay]] <- scrna.list[[i]][[initial_assay]]
        return(obj)
      }
    )
  } else {
    scrna.assay.list <- scrna.list
  }
  
  for (i in 1:length(scrna.assay.list)) {
    DefaultAssay(scrna.assay.list[[i]]) <- assay_name
  }
  
  integrate.features <- SelectIntegrationFeatures(
    object.list = scrna.assay.list, 
    nfeatures = 3000,
    assay=rep(assay_name, length(scrna.assay.list)))
  gc()
  
  scrna.assay.list <- PrepSCTIntegration(
    assay=assay_name, object.list = scrna.assay.list, 
    anchor.features = integrate.features, 
    verbose = TRUE)
  gc()
  
  if (use_ref) {
    scrna.anchors <- FindIntegrationAnchors(
      object.list = scrna.assay.list,
      normalization.method = "SCT",
      anchor.features = integrate.features,
      reference = use_ref)
  } else {
    scrna.anchors <- FindIntegrationAnchors(
      object.list = scrna.assay.list,
      normalization.method = "SCT", 
      anchor.features = integrate.features,
      verbose = TRUE)
  }
  gc()
  
  scrna.integrated <- IntegrateData(
    anchorset = scrna.anchors,
    normalization.method = "SCT", 
    verbose = TRUE,
    new.assay.name = integrated_assay_name)
  gc()
    
  if (assay_only) {
    return(scrna.integrated[[integrated_assay_name]])
  } else {
    return(scrna.integrated)
  }
}

scrna.integrated <- integrate_feature(
  scrna.list, 
  assay_name='SCT',
  initial_assay='RNA',
  assay_only=F,
  remove_other_assays=F)
scrna.integrated[['SCT_ADT_INT']] <- integrate_feature(
  scrna.list,
  assay_name='SCT_ADT',
  initial_assay='ADT',
  assay_only=T,
  remove_other_assays=T)

integrated.rds.file <- here::here(
     '..','..','scrnaseq_tcsl154',
     'rds',
     paste0('combined.integrated.umap.data.rds'))

saveRDS(scrna.integrated,integrated.rds.file)
```

### Prune HTO and CD4/CD8 Doublets, assign CD4/CD8

```{r}

scrna.integrated <- subset(scrna.integrated, subset = (HTO_status %in% c('HighQual','MidQual')))

hto_subset <- subset(scrna.integrated, cells=sample(Cells(scrna.integrated), 8000))

t_out <- reassign_t_type(scrna.integrated, n_clust=3, plots)
plots <- t_out$plot_obj
scrna.integrated <- t_out$seurat_obj
rm(t_out)

plots$t_type_bars <- ggplot(
  data.table(scrna.integrated@meta.data)) +
  geom_bar(aes(x=car, fill=t_type)) +
  facet_grid(k_type~donor, space='free', scales='free') +
  coord_flip()
```
## Run WNN Clustering on all cells

```{r}

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
scrna.integrated <- RunPCA(scrna.integrated, features = c(s.genes, g2m.genes), reduction.name = 'cc_pca')

scrna.integrated <- RunPCA(scrna.integrated, assay='SCT_INT', reduction.name = 'rna_pca')

# identify unique TCSL170 ADTs
adt_count_ratios <- (
  Matrix::rowSums(subset(scrna.integrated, subset=(donor == 'TCSL154'))[['SCT_ADT_INT']]@counts) / 
    Matrix::rowSums(subset(scrna.integrated, subset=(donor == 'TCSL170'))[['SCT_ADT_INT']]@counts))
tcsl170_adts <- names(which(-log10(adt_count_ratios) > 2))
shared_adts <- names(which(-log10(adt_count_ratios) < 2))

scrna.integrated <- RunPCA(scrna.integrated, assay='SCT_ADT_INT', features= shared_adts, 
    nfeatures.print = 10, reduction.name='adt_pca')
scrna_output <- sct_umap(scrna.integrated, title='All cells')

scrna.integrated <- scrna_output$obj
plots <- c(plots, scrna_output$plots)
rm(scrna_output)

plots$combined_umap_2 <- (
  DimPlot(scrna.integrated, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5) + 
    NoLegend() +
  DimPlot(scrna.integrated, reduction = 'wnn.umap', group.by = 't_type', label = TRUE, repel = TRUE, label.size = 2.5) + 
    NoLegend() +
  DimPlot(scrna.integrated, reduction = 'wnn.umap', group.by = 'k_type', label = TRUE, repel = TRUE, label.size = 2.5) +
    NoLegend() +
  DimPlot(scrna.integrated, reduction = 'wnn.umap', group.by = 'Phase', label = TRUE, repel = TRUE, label.size = 2.5) + 
    NoLegend() +
  DimPlot(scrna.integrated, reduction = 'wnn.umap', group.by = 'donor', label = TRUE, repel = TRUE, label.size = 2.5) + 
    NoLegend())
```


## Map TCSL170 subtypes onto new UMAP

```{r}

scrna.wnn.cd4_all_metadata <- readRDS("~/scrnaseq_tcsl154/rds/scrna.wnn.cd4_all_metadata.rds")
scrna.wnn.cd8_all_metadata <- readRDS("~/scrnaseq_tcsl154/rds/scrna.wnn.cd8_all_metadata.rds")

original_metadata <- rbind(
    data.table(scrna.wnn.cd4_all_metadata, keep.rownames=T)[, list(rn_154=rn, sample, t_type, Subtype)],
    data.table(scrna.wnn.cd8_all_metadata, keep.rownames=T)[, list(rn_154=rn, sample, t_type, Subtype)])[,
      rn := gsub('(.*)-\\d','\\1',rn_154)]

integrated_metadata <- data.table(
  scrna.integrated@meta.data, keep.rownames=T)[,
    list(rn_orig=rn, sample, donor, orig_t_type=t_type)][
      donor == 'TCSL154', rn := gsub('(.*)-\\d_\\d','\\1',rn_orig)]

sample_subtype_map <- merge(integrated_metadata, original_metadata, by='rn', all=T)[!is.na(Subtype) & !is.na(rn_orig)]

scrna.integrated@meta.data[,'subtype_154'] <- factor(NA, levels=levels(sample_subtype_map$Subtype))
scrna.integrated@meta.data[sample_subtype_map$rn_orig,]$subtype_154 <- sample_subtype_map$Subtype

DimPlot(scrna.integrated, reduction = 'wnn.umap', group.by = 'subtype_154', label = TRUE, repel = TRUE, label.size = 2.5)
```

## Donor, T, and Stim amounts per cluster
```{r}
ggplot(data.table(scrna.integrated@meta.data)[t_type != 'DP'][, 
    .N, by=c('seurat_clusters','donor','k_type','t_type')][N > 100]) + 
  geom_bar(aes(x=seurat_clusters, y=N, fill=interaction(donor,k_type)), stat='identity') +
  facet_wrap(~t_type, scales='free')

## END OF DATA LOADING

```
------------------------------------------------------------------------------------------------------------------------

Finally, ran UMAP creation for all subsets made HD5 seurat objects for each with `07_run_subset_job_indiv.r`.

```{r}




```

------------------------------------------------------------------------------------------------------------------------

## Map new clusters onto old subtypes


Easiest way is to match based on most frequently assigned subtype. Do CD8 only first.

```{r}
scrna_cd8 <- LoadH5Seurat(here::here(
  '..','..','scrnaseq_tcsl154',
  'rds',
  paste0('scrna.wnn.int.CD8.stim.clean.h5Seurat')))

cd8_metadata <- data.table(
  scrna_cd8@meta.data, keep.rownames=T)[,
    list(rn_orig=rn, sample, donor, orig_t_type=t_type)][
      donor == 'TCSL154', rn := gsub('(.*)-\\d_\\d','\\1',rn_orig)]

sample_subtype_map <- merge(cd8_metadata, original_metadata, by='rn', all=T)[!is.na(Subtype) & !is.na(rn_orig)]
get_cluster_enrich_plot(scrna_cd8, 'seurat_clusters', hide='', facet_formula=formula(~car))

scrna_cd8@meta.data[,'subtype_154'] <- factor(NA, levels=levels(sample_subtype_map$Subtype))
scrna_cd8@meta.data[sample_subtype_map$rn_orig,]$subtype_154 <- sample_subtype_map$Subtype

DimPlot(scrna_cd8, reduction = 'wnn.umap', group.by = 'subtype_154', label = TRUE, repel = TRUE, label.size = 2.5) +
  DimPlot(scrna_cd8, reduction = 'wnn.umap', group.by = 'seurat_clusters', label = TRUE, repel = TRUE, label.size = 2.5)

cd8_umap_plots <- umap_plot(scrna_cd8, assay='SCT_INT', cluster_name='seurat_clusters', reduction='wnn.umap')
cd8_umap_plots_adt <- umap_plot(scrna_cd8, assay='SCT_ADT_INT', cluster_name='seurat_clusters', reduction='wnn.umap')

clust_mapping_154 <- compare_clusters(scrna_cd8, 'seurat_clusters', 'subtype_154')

ggplot(data.table(scrna_cd8@meta.data)[, 
    list(N=.N, seurat_clusters), by=c('donor','car','k_type')][, 
      pct := .N/N, by=c('donor','car','seurat_clusters','k_type')]) + 
  geom_bar(aes(x=seurat_clusters, fill=donor, y=pct), stat='identity', position='dodge') + 
  facet_grid(car + k_type ~ .)
```

### Sheih Et all IRF Gene Signature

```{r}

# Genes from IRF cluster in Fig 7b of Shieh et al 2020 

scrna_cd8 <- AddModuleScore(scrna_cd8, features=list('1'=c('CCL4','CCL5','CD27','CD74','FGFBP2','GNLY','GZMA','GZMB','GZMH','GZMK','HLA-DRA','HLA-DRB1','HLA-DRB5','IFNG','LAG3','NKG7','STMN1')),name='Sheih_et_al_IRF')


```
