---
title: "pooled_05"
author: "Daniel Goodman"
date: "4/7/2020"
output:
  html_document:
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, 
  output_dir = here::here('..','html','pooled'))})
---

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(ggrepel)
library(RColorBrewer)
library(DESeq2)
#library(rnaseqGene)
library(plotly)
library(Rtsne)
library(scales)
library(dplyr)
library(ggtext)
library(rstatix)

rm(list = ls())

setDTthreads(8)

s3.data.dir <- '/Volumes/s3/roybal-tcsl/lenti_screen_compiled_data/data'

#sudo ln -s /Users/dbg/Library/Group\ Containers/G69SCX94XU.duck/Library/Application\ Support/duck/Volumes/s3.amazonaws.com\ –\ S3 /Volumes/S3
data.output.dir <- file.path(s3.data.dir)

load(file=file.path(data.output.dir, 'pooled_deseq2_analysis_data.Rdata'))
s3.data.dir <- '/Volumes/s3/roybal-tcsl/lenti_screen_compiled_data/data'
data.output.dir <- file.path(s3.data.dir)
load(file.path(data.output.dir, 'pooled_cytokine_repeat_data.Rdata'))

# arrayed list of CARs
array.list <- c('BAFF-R','TNR8','4-1BB','TACI','CD28','KLRG1','CD40')

# load colors
source(here::here('r','fig_colors.R'))

`%ni%` <- Negate(`%in%`)

names(car_colors) <- c(
  "CD28", "4-1BB", "BAFF-R", "TACI", "CD40", "CD30",
  "KLRG1", "zeta", "Unt", "None")

```

## deSeq2 normalization followed by CAR score calculation

```{r}
all.deseq.results.dt[CAR.align == 'TNR8', CAR.align := 'CD30']
all.deseq.results.dt[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']

all.deseq.results.dt[, padj.disp := -log10(padj)]
all.deseq.results.dt[, lfc.disp := log2FoldChange]
all.deseq.results.dt[padj.disp > 10, padj.disp := Inf]
all.deseq.results.dt[abs(lfc.disp) > 5, lfc.disp := sign(lfc.disp) * Inf]

data.dt <- all.deseq.results.dt

data.dt[, CAR.type := 'other']
data.dt[CAR.align == '4-1BB', CAR.type := '4-1BB']
data.dt[CAR.align == 'CD28', CAR.type := 'CD28']
data.dt[, CAR.type := factor(
  CAR.type,levels=c('other','4-1BB','CD28'))]
```

### Single heatmap, CD4 plus CD8.

```{r fig.height=4, fig.width=2.5, units='inches'}

# MAD: constant * median(abs(x-median))
# robust z-scores: x - median / median(abs(x - median))
robust_z_score = function(value) 0.6745*(value - median(value))/ mad(value)

ggnormal_dist <- stat_function(
  fun = dnorm, args = list(mean = 0, sd = 1),
  col = "#1b98e0", size = 1, linetype=2)

# stouffer's method: z score over sqrt of number of values
stouffer.z <- function(z) sum(z, na.rm=T) / sqrt(sum(!is.na(z)))
z.to.p.adj.bh <- function(z) p.adjust(pnorm(abs(z), lower.tail=F), "BH")

#current stats workflow is:
# get a Z-score for every group
# combine Z scores using stouffers
# use Benjamani Hochberg to correct for multiple hypotheses

##deseq2
all.deseq.results.dt[, comparison := paste(test_set,ref_set,sep= '.')]

deseq2_combined <- all.deseq.results.dt[,
    comparison := paste(test_set, ref_set, sep='.')][,
    combined_var := paste(comparison, assay)][, value := log2FoldChange]

deseq2_combined[, value.scale := scale(value), by = .(t.type, k.type, assay, comparison)]

deseq2_combined[,
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]

ggplot(deseq2_combined[k.type == 'pos' & comparison %in% c('A.baseline')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align, value.scale),
  color=interaction(assay, comparison), size=padj.disp > -log(0.05))) +
  facet_wrap(~t.type)

## rlog car score
############################

# combine z scores using stouffer's method: z score over sqrt of number of values
car_score_means <- unique(read.counts[assay != 'baseline' & batch != 'post-cytof',
    list(rlog_car_score, CAR.align, assay, t.type, k.type, donor, sort.group, batch)])[,
  value.scale := scale(rlog_car_score), by=sort.group]
car_score_means[,
    value.t.type := stouffer.z(unique(value.scale)),
      by=.(CAR.align, t.type, k.type, assay)]
car_score_means[,
    value.overall := stouffer.z(unique(value.scale)),
      by=.(CAR.align, k.type, assay)]

car_score_means[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, k.type, assay)]
car_score_means[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, k.type, assay)]

ggplot(car_score_means[k.type == 'pos' & assay %in% c('CTV1','CTV2')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(car_score_means[k.type == 'pos' & assay %in% c('CTV1','CTV2')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(car_score_means[assay != 'CTV3' & k.type == 'pos']) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)

car_score_means[, 
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]
car_score_means[, comparison := 'rlog_car_score']
car_score_means[, 
  combined_var := paste(assay,comparison, sep='.'), by=.(t.type)]

# fc in proliferation
##########################

rel_abund_means <- read.counts[assay != 'baseline' & batch != 'post-cytof'][,
  value.scale := scale(log(car.abund.rel.baseline)), by=sort.group]
rel_abund_means[,
    value.t.type :=  stouffer.z(unique(value.scale)), 
      by=.(CAR.align, t.type, k.type, assay)]
rel_abund_means[,
    value.overall :=  stouffer.z(unique(value.scale)), 
      by=.(CAR.align, k.type, assay)]

rel_abund_means[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, k.type, assay)]
rel_abund_means[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, k.type, assay)]


rel_abund_means[, comparison := 'abund_rel_baseline']
rel_abund_means[, 
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]
rel_abund_means[, 
  combined_var := paste(assay,comparison, sep='.'), by=.(t.type)]

ggplot(rel_abund_means[k.type == 'pos' & assay %in% c('CTV2','CTV3')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(rel_abund_means[k.type == 'pos' & assay %in% c('CTV2','CTV3')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(rel_abund_means[assay != 'CTV3' & k.type == 'pos']) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)


### CD69 
cd69.ranked <- melt(
  read.counts[batch == 'post-cytof' & assay == 'CD69'], 
  measure.vars = c('min.max.ratio.norm','all.max.ratio.norm','rlog_car_score'),
  variable.name='comparison')[bin == 'A' & k.type == 'pos']
cd69.ranked[, value.scale := scale(value),
  by=.(sort.group,comparison,t.type)]
cd69.ranked <- cd69.ranked[comparison=='rlog_car_score']

cd69.ranked[, value.overall :=  stouffer.z(unique(value.scale)), 
  by=.(CAR.align, k.type, assay, comparison)]

cd69.ranked[, value.t.type :=  stouffer.z(unique(value.scale)), 
  by=.(CAR.align, t.type, k.type, assay, comparison)]

cd69.ranked[, 
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]
cd69.ranked[, 
  combined_var := paste(assay,comparison, sep='.'), by=.(t.type)]

cd69.ranked[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, k.type, assay)]
cd69.ranked[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, k.type, assay)]

ggplot(cd69.ranked) + geom_point(aes(x=value.scale,  y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type))) + geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type)), pch=21, size=2) + facet_wrap(~k.type+t.type+comparison) + geom_vline(xintercept=c(-1.96,1.96))

ggplot(cd69.ranked) + geom_point(aes(x=value.scale,  y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type))) + geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type)), pch=21, size=2) + facet_wrap(~k.type+comparison) + geom_vline(xintercept=c(-1.96,1.96))

ggplot(cd69.ranked) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(cd69.ranked) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(cd69.ranked) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)

### Cytokines - IL2 and IFNG with new and old donors

old.read.counts <- read.counts[assay %in% c('IL2','IFNy')]

old.read.counts[assay == 'IFNy', assay := 'ifng']
old.read.counts[assay == 'IL2', assay := 'il2']

old.read.counts[assay == 'ifng', sort.cond := ifelse(bin %in% c('A','B'), 'neg','ifng')]
old.read.counts[assay == 'il2', sort.cond := ifelse(bin %in% c('A','B'), 'neg','il2')]

old.read.counts <- old.read.counts[, list(
    CAR.align,
    cell.count,
    bin.pct,
    car.bin.pct,
    car.bin.cells= cell.count*car.bin.read.pct),
  by=.(donor, assay, bin, sort.cond, k.type, t.type)]

old.read.counts <- old.read.counts[, list(
  cell.count= cell.count[1],
  car.bin.cells = sum(car.bin.cells),
  bin.pct= bin.pct[1],
  car.bin.pct.incl= sum(car.bin.pct),
  car.bin.pct= sum(car.bin.pct)),
  by=.(sort.cond, CAR.align, donor, assay, t.type, k.type)]

old.read.counts[, sort.group := paste(donor, assay, t.type, k.type, sep='_')]
old.read.counts[, car.sort.group.cells := sum(car.bin.cells), 
    by=.(CAR.align, sort.group)]

shared_cols <- intersect(
  names(old.read.counts), 
  names(cyto.read.counts.summed))

cyto.read.counts.all <- rbind(
  cyto.read.counts.summed[, ..shared_cols],
  old.read.counts[, ..shared_cols])

cyto.ranked.all <- cyto.read.counts.all[
  k.type == 'pos' & assay %in% c('il2.ifng', 'il2','ifng') &
  sort.cond %in% c('il2','ifng') &
  car.sort.group.cells > 500]

cyto.ranked.all[, value.scale := scale(
  log(car.bin.pct.incl)), by=.(sort.cond, donor, t.type, k.type)]

cyto.ranked.all[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']
cyto.ranked.all[CAR.align == 'BTLA.a', CAR.align := 'BTLA']
cyto.ranked.all[, assay := factor(sort.cond, labels=c('IFNy','IL2'))]
cyto.ranked.all <- cyto.ranked.all[CAR.align %ni% c('BTLA.b','BTLA.a')]

cyto.ranked.all[, group := paste(sort.cond,t.type,k.type)]
cyto.ranked.all[, comparison := 'pct_pos']
cyto.ranked.all[, combined_var := paste(comparison, sort.cond)]
cyto.ranked.all[, car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]

cyto.ranked.all[, value.t.type := stouffer.z(value.scale), by=.(CAR.align, t.type)]
cyto.ranked.all[, value.assay := stouffer.z(value.scale), by=.(CAR.align, assay)]
cyto.ranked.all[, value.ta := stouffer.z(value.scale), by=.(CAR.align, t.type, assay)]
cyto.ranked.all[, value.overall := stouffer.z(value.scale), by=.(CAR.align)]

cyto.ranked.all[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type)]

cyto.ranked.all[, p.adj.a := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, assay)]

cyto.ranked.all[, p.adj.ta := p.adjust(2*pnorm(abs(value.ta), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, assay)]

cyto.ranked.all[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align)]


ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.assay, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.a < 0.05), pch=21) +
  facet_wrap(~sort.cond)

ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.ta, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.ta < 0.05), pch=21) +
  facet_wrap(~t.type+assay)

ggplot(cyto.ranked.all) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)

### Combined columns, also pvalue
shared_cols <- c(intersect(
  intersect(
    intersect(names(rel_abund_means),names(cd69.ranked)),
    intersect(names(cyto.ranked.all),names(car_score_means))),
  names(deseq2_combined)), 'padj')

total_combined <- unique(rbind(
  #deseq2 measurements for CTV1
  deseq2_combined[, shared_cols, with=F],
  
  #new and old cytokine combined with z-scores and p-values
  unique(copy(cyto.ranked.all)[, padj := p.adj.ta][, value.scale := value.ta][, 
    shared_cols, with=F]),
  
  #cd69 from both donors
  copy(cd69.ranked)[, padj := p.adj.t][
    comparison == 'rlog_car_score', shared_cols, with=F],
  
  #car scores for CTV
  copy(car_score_means)[, padj := p.adj.t][, 
    comparison := 'rlog_car_score'][, shared_cols, with=F],
  
  #relative abundance/expansion
  copy(rel_abund_means)[, padj := p.adj.t][, value.scale := value.t.type][
    comparison == 'abund_rel_baseline', shared_cols, with=F],
  
  use.names=F))

```

### Together and separate CD4 and CD8 heatmaps.
```{r fig.height=8, fig.width=6}
library(dendextend)
library(ggdendro)
library(cowplot)
library(gtable)
library(grid)


remove_t_type_sep_col_all_cars <- function (breaks) {
  car_name <- unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))
  
  car_axis_colors <- rep('black', length(car_name))
  for (car_i in names(car_colors)) {
    if (car_i %in% car_name) 
      car_axis_colors[(car_i == car_name)] <- car_colors[car_i]
  }

  # bold CAR names if colored
  car_name[car_axis_colors != 'black'] <- paste0('**',car_name[car_axis_colors != 'black'],'**')
  
  car_color_span <- paste("<span style = 'color: ",
    car_axis_colors,
    ";'>",
    car_name,
    "</span>", sep = "")
  
  return(car_color_span)
}

make_dendroheatmap <- function(
  df, title,
  legend_theme=theme(legend.position = "none"), boxnumbers=F)
{
  
  # Make dendrogram
  car_mean_cast <- dcast(
    df[k.type == 'pos'], 
    CAR.align ~ assay + t.type, 
    value.var='value.scale')
  
  car_dendro_m <- as.matrix(car_mean_cast[, -1])
  rownames(car_dendro_m) <- unlist(car_mean_cast[,1])
  car_dendro <- as.dendrogram(hclust(d = dist(x = car_dendro_m)))
  
  # order cars based only on significant values
  car_order <- rownames(car_dendro_m)[
    order(rowMeans(car_mean_cast * as.numeric(dcast(
      df[k.type == 'pos'], 
      CAR.align ~ assay + t.type, 
      value.var='padj')[, -1] < 0.05), na.rm=T))]

  car_dendro <- rotate(car_dendro, car_order)

  # Create dendrogram plot
  dendro_vars_plot <- ggdendrogram(data = car_dendro, rotate = FALSE) + 
    theme(axis.text.y = element_text(size = 6))
  
  # row order
  df$CAR.align <- factor(
      df$CAR.align,
      levels = labels(car_dendro))
  
  #manual x labels
  y_lab_man <- c(
    IFNy.CD4='IFNy 18h', IL2.CD4='IL2 18h',
    CD69.CD4='CD69 18h', CD69.CD8='CD69 18h',
    IFNy.CD8='IFNy 18h', IL2.CD8='IL2 18h',
    CTV1.CD4='Cell divisions, d0-d3/d4', CTV1.CD8='Cell divisions, d0-d3/d4',
    CTV2.CD4='Cell divisions, d9-14/16', CTV2.CD8='Cell divisions, d9-14/16',
    CTV3.CD4='Total expansion d0-24', CTV3.CD8='Total expansion d0-24'
    )
  
  # heatmap
  var_heatmap <-   ggplot(df[,
      t.type := factor(t.type, levels=c('CD8','CD4'))][,
        assay := factor(assay, levels= c('IL2','IFNy','CD69','CTV1','CTV2','CTV3'))]) + 
    geom_tile(aes(y = interaction(assay,t.type), x = CAR.align, 
                  fill = value.scale), colour = "black", size = .20, alpha=ifelse(boxnumbers, 0.5, 1)) + 
    geom_tile(data=df[padj < 0.05],
      aes(y = interaction(assay,t.type), x = CAR.align),
      fill=NA,
      color='black', size=0.5,
      alpha=ifelse(boxnumbers, 0.5, 1)) + 
    scale_fill_distiller('Z-score', 
      palette='PiYG', limits=c(-4,4), oob=scales::squish, direction=1) +
    scale_y_discrete(expand=c(0,0),
      labels=y_lab_man) +
    scale_x_discrete(labels= remove_t_type_sep_col_all_cars, expand=c(0,0)) +
    labs(title=title) + 
    facet_grid(t.type~., scales='free', space='free', switch='y') + 
    coord_cartesian(clip="off") +
    theme(
        strip.placement = "outside",
        axis.text.x = element_markdown(angle = 90, hjust = 1, size=unit(8,'pt')),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        plot.margin=margin(-19,5.5,5.5,5.5, 'points'),
        strip.background = element_blank(),
        panel.background = element_rect(fill='grey90', color='black', size=0.2),
        panel.grid = element_blank(),
        axis.title.x = element_blank())
  
    if (boxnumbers) {
      var_heatmap <- var_heatmap + 
        geom_text(aes(x = interaction(assay,t.type), y = CAR.align, 
          label=round(value.scale,1)), colour = "black")
    }
  
  var_heatmap <- var_heatmap + legend_theme
  
  # col dendro
  dendro_data_col <- dendro_data(car_dendro, type = "rectangle")
  
  dendro_col <- axis_canvas(var_heatmap, axis = "x") +
    geom_segment(data = segment(dendro_data_col),
        aes(x = x, y = y*0.99, xend = xend, yend = yend*0.99))
  # plot_dendroheat <- insert_xaxis_grob(var_heatmap, 
  #   dendro_col, dendro_units, position = "top")
  
  plot_dendroheat <- (
    dendro_col+theme(plot.margin = margin(10,0,0,0))) / 
    (var_heatmap + theme(
        plot.margin= margin(0,40,0,0), panel.spacing= unit(5,'pt'), plot.title=element_blank())
    ) +
    plot_layout(heights=c(1,6))
  
  return(plot_dendroheat)
}

chosen_metrics <- total_combined[k.type == 'pos' &
  ((assay %in% c('CTV1','CTV2') & comparison == 'AB.D') |
    (assay %in% c('CTV3') & comparison == 'abund_rel_baseline') |
    (assay %in% c('IL2','IFNy') & comparison == 'pct_pos') |
      (!(assay %in% c('CTV1','CTV2','CTV3')) & comparison == 'rlog_car_score'))]

fig_3_single_dendro <- make_dendroheatmap(
  df=chosen_metrics,
  title='',
  legend_theme=theme(
      legend.position = c(1, 0.5), 
      legend.justification = c(0, 0.5),
      plot.margin = margin(5.5,66,5.5,44, unit="pt")))

fig_3_single_dendro
```

###  redo deseq2 with combined 4v8
```{r}

# in testing_deseq2.R, we found that AB.D is the best within-CTV comparison.
# (for testing proliferation explicitly, not expansion (via baseline))

#run_deseq fxn
#--------
run_deseq_combined48 <- function(data.dt, ref_bin, test_bin, 
    control_replicates = T,
    interaction = T, group.control = F, weight.bins = F) 
{
  # identify inputs
  assay_input <- as.character(unique(data.dt[, assay]))
  k_type <- as.character(unique(data.dt[, k.type]))
  t_type <- as.character(unique(data.dt[, t.type]))
  
  ## 1. Do bin normalization weights =======
  # bin normalization weights
  data.weights <- unique(data.dt[, list(
    batch, donor, timepoint, assay, t.type, k.type, 
    sort.group, bin, bin.pct, bin.reads)])[, 
      list(bin, bin.reads, bin.pct, sort.group,
        read.weight=bin.pct * bin.reads / sum(bin.pct * bin.reads)),
      by=.(batch, donor, timepoint, assay, t.type, k.type)][, 
        read.weight.norm := read.weight/exp(mean(log(read.weight))), 
        by=.(batch, donor, timepoint, assay, t.type, k.type)]
  
  #stopifnot(nrow(interaction(assay_input, k_type, t_type)) == 1)
  
  # prepare cts and coldata dataframes
  
  ## 2. Prepare Ref Bins ============
  if(length(ref_bin) == 1 & ref_bin[1] == 'baseline') {
    # reference is baseline
    # get baseline counts per donor/assay replicate
    ref.bin.dt <- dcast(
      data.dt[
        bin == 'D', 
        .(
          CAR.align, 
          bin.sort.group = paste(
            batch, donor, timepoint, assay, t.type, 'base', sep = '_'), 
          k.type, t.type, batch, assay, donor, 
          sort.group, 
          bin = 'base',
          counts = baseline.counts)], 
      CAR.align ~ bin.sort.group, value.var='counts')
    
    ref.weights <- rep(1, ncol(ref.bin.dt))
    
  } else {
    # reference is a specified bin
    ref.bin.dt <- dcast(
      data.dt[
        bin %in% ref_bin, 
        .(
          CAR.align, 
          bin.sort.group = paste(sort.group, bin, sep = '_'), 
          k.type, t.type, batch, assay, donor, 
          sort.group, 
          bin,
          counts)], 
      CAR.align ~ bin.sort.group, value.var='counts')
    
    ref.weights <- dcast(data.weights[
        bin %in% ref_bin,
        .(
          bin.sort.group = paste(sort.group, bin, sep = '_'), 
          k.type, t.type, batch, assay, donor, 
          sort.group,
          bin,
          read.weight.norm)],
        . ~ bin.sort.group, 
        value.var = 'read.weight.norm')
    
    stopifnot(nrow(ref.bin.dt) == nrow(unique(ref.bin.dt)))
  }
  
  # copy the ref bin columns for each of the test bin columns
  if (interaction == T) {
    
    num.ref.reps <- ncol(ref.bin.dt) - 1
    
    ref.bin.dt <- cbind(ref.bin.dt[, 1], 
      do.call("cbind", replicate(length(test_bin), 
      ref.bin.dt[, -1], simplify = FALSE)))
    
    names(ref.bin.dt) <- c(names(ref.bin.dt[, 1]), 
      paste(names(ref.bin.dt[, -1]), rep(test_bin, each=num.ref.reps), 
        sep = '_'))
  }
  
  ## 3. Prepare Test Bins ============
  test.bin.dt <- dcast(
      data.dt[
        bin %in% test_bin,
        .(
          CAR.align, 
          bin.sort.group = paste(sort.group, bin, sep = '_'), 
          k.type, t.type, batch, assay, donor, 
          sort.group, 
          bin,
          counts)], 
      CAR.align ~ bin.sort.group, value.var='counts')
  
  # check that replicate counts match
  stopifnot(nrow(ref.bin.dt) == nrow(unique(ref.bin.dt)))
  stopifnot(nrow(test.bin.dt) == nrow(unique(test.bin.dt)))
  
  ## 4. Merge and create design matrix ============
  cts <- merge(ref.bin.dt, test.bin.dt, by = 'CAR.align')
  cts <- data.frame(cts[, -1], row.names = cts[, CAR.align])
  cts[is.na(cts)] <- 0
  
  coldata <- data.frame(
    condition = c(
      rep('reference', ncol(ref.bin.dt) - 1), 
      rep('test', ncol(test.bin.dt) - 1)),
    rep = data.table(t(sapply(strsplit(c(
      names(ref.bin.dt)[-1], 
      names(test.bin.dt)[-1]),"_"), `[`, c(1,2))))[,
        paste(V1, V2, sep='_')],
    bin = sapply(strsplit(c(
      names(ref.bin.dt)[-1], 
      names(test.bin.dt)[-1]),"_"), `[`, 7),
    t.type = gsub('.*(CD[48]).*','\\1',names(ref.bin.dt)[-1]),
    row.names = c(names(ref.bin.dt)[-1], names(test.bin.dt)[-1]))
  
  dds <- DESeqDataSetFromMatrix(countData = cts,
                                colData = coldata,
                                design =  ~ condition + rep + t.type)
  
  # set reference
  dds$condition <- relevel(dds$condition, ref = "reference")
  
  print(coldata)
  
  # pre-filtering
  keep <- rowSums(counts(dds)) >= 10
  dds <- dds[keep,]
  
  dds <- DESeq2::DESeq(dds)
  
  # shrink log fold change & convert to data.table, merge contrasts
  results.dt <- rbindlist(
      lapply(resultsNames(dds)[-1], function(result_i) 
          as.data.table(lfcShrink(dds, coef=result_i, type="apeglm"), 
        keep.rownames='CAR.align')[, coef := result_i]))
  
  results.dt[, assay := assay_input][, k.type := k_type]
  
  return(results.dt)
}

ref_set <- 'D'
test_set <- c('A','B')

ref_str <- paste0(ref_set, collapse='')
test_str <- paste0(test_set, collapse='')
inter <- "F"

deseq.results.dt <- read.counts[batch != 'post-cytof' & k.type == 'pos',
  {
    message(paste(c(ref_str, test_str, inter, .BY[1],"\n"), collapse= ' - '));
    tryCatch(
      run_deseq_combined48(
        data.dt = data.table(.SD)[, assay := .BY[1]], 
        ref_bin = ref_set,
        test_bin = test_set,
        interaction = inter,
        group.control = T),
      error= function(e) {message(e); return(data.table())}
    )
  },
  by = .(assay)]
```

### heatmap that combines CD4 and CD8
```{r}

total_combined_scores_cd4_cd8 <- rbind(
  # 'proliferation' - deseq2 CTV1 and 2
  deseq.results.dt[
    assay %in% c('CTV1', 'CTV2') &
    coef == 'condition_test_vs_reference',
    list(CAR.align, padj, value.scale= as.vector(scale(log2FoldChange))),
    by=.(assay)][, assay := paste('proliferation', assay, sep='.'),],
  
  #cytokines combined 4 & 8
  unique(cyto.ranked.all[k.type == 'pos' & car.sort.group.cells > 400, list(
    assay=sort.cond,
    CAR.align, padj= p.adj.a, value.scale= value.assay)]),
  
  #cd69 combined 4 & 8
  unique(cd69.ranked[, list(
    assay, CAR.align, padj=p.adj.o, value.scale= value.overall)]),
  
  #'expansion' - rel abund CTV2 and CTV3
  unique(rel_abund_means[
    comparison == 'abund_rel_baseline' &
      k.type == 'pos' &
      assay %in% c('CTV3'), list(
        assay=paste('expansion', assay, sep='.'),
        CAR.align, padj=p.adj.o, value.scale=value.overall)])
)

#reorder assay levels
total_combined_scores_cd4_cd8[, assay := factor(
  assay, levels=c(
    'il2','ifng','CD69','expansion.CTV3',
    'proliferation.CTV2','proliferation.CTV1'))]

make_dendroheatmap_cd4cd8_combined <- function(
  df, title,
  legend_theme=theme(legend.position = "none"), boxnumbers=F)
{
  
  # Make dendrogram
  car_mean_cast <- dcast(
    df, 
    CAR.align ~ assay, 
    value.var='value.scale')
  
  car_dendro_m <- as.matrix(car_mean_cast[, -1])
  rownames(car_dendro_m) <- unlist(car_mean_cast[,1])
  car_dendro <- as.dendrogram(hclust(d = dist(x = car_dendro_m)))
  
  # COMMENTED, not weighting
  # rotate dendrogram leaves according to CAR ranking.
  # weight measurement types 
  # - weight late expansion more highly in ranking of x axis, cytokines lower

  # ctv3_weight <- 1.5
  # ctv2_weight <- 2
  # cytokine_weight <- 0.5
  # 
  # col_weights <- 1+(ctv2_weight-1)*grepl('CTV2', colnames(car_dendro_m))
  # col_weights <- col_weights+(ctv3_weight-1)*grepl('CTV3', colnames(car_dendro_m))
  # col_weights <- col_weights+(cytokine_weight-1)*grepl('IL2|IFNy', colnames(car_dendro_m))
  # 
  # car_order <- names(sort(
  #   rowSums(car_dendro_m * rep(col_weights, each=nrow(car_dendro_m)), na.rm=T) /
  #     sum(col_weights)))
  # car_dendro <- rotate(car_dendro, car_order)
  
  # order cars based only on significant values
  car_order <- rownames(car_dendro_m)[
    order(rowMeans(car_mean_cast * as.numeric(dcast(
      df, 
      CAR.align ~ assay, 
      value.var='padj')[, -1] < 0.05), na.rm=T))]

  car_dendro <- rotate(car_dendro, car_order)
  

  # Create dendrogram plot
  dendro_vars_plot <- ggdendrogram(data = car_dendro, rotate = FALSE) + 
    theme(axis.text.y = element_text(size = 6))
  
  # row order
  df$CAR.align <- factor(
      df$CAR.align,
      levels = labels(car_dendro))
  
  #manual x labels
  y_lab_man <- c(
    il2='IL2 18h', ifng='IFNγ 18h',
    CD69='CD69 18h',
    proliferation.CTV1='Cell divisions, d0-d3/4',
    proliferation.CTV2='Cell divisions, d9-d14/16',
    expansion.CTV3='Total expansion, d0-d24'
    )
  
  # heatmap
  var_heatmap <- ggplot(df) + 
    geom_tile(
      aes(y = assay, x = CAR.align, fill = value.scale),
      color='black', size=0.2,
      alpha=ifelse(boxnumbers, 0.5, 1)) + 
    geom_tile(data=df[padj < 0.05],
      aes(y = assay, x = CAR.align),
      fill=NA,
      color='black', size=0.5,
      alpha=ifelse(boxnumbers, 0.5, 1)) + 
    scale_fill_distiller('Z-score', 
      palette='PiYG', limits=c(-4,4), oob=scales::squish, direction=1) +
    scale_y_discrete(expand=c(0,0), labels=y_lab_man) +
    scale_x_discrete() +
    coord_cartesian(clip="off") +
    theme(
        strip.placement = "outside",
        axis.text.x = element_markdown(angle = 90, hjust = 1),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        plot.margin=margin(-19,5.5,5.5,5.5, 'points'),
        strip.background = element_blank(),
        panel.background = element_rect(fill='white'),
        axis.title.x = element_blank())
  
  var_heatmap <- var_heatmap + legend_theme
  
  # col dendro
  dendro_data_col <- dendro_data(car_dendro, type = "rectangle")
  
  dendro_col <- axis_canvas(var_heatmap, axis = "x") +
    geom_segment(data = segment(dendro_data_col),
        aes(x = x, y = y*0.99, xend = xend, yend = yend*0.99))
  # plot_dendroheat <- insert_xaxis_grob(var_heatmap, 
  #   dendro_col, dendro_units, position = "top")
  
  plot_dendroheat <- (
    dendro_col+theme(plot.margin = margin(10,0,0,0))) / 
    (var_heatmap + theme(
        plot.margin= margin(0,40,0,0), panel.spacing= unit(5,'pt'), plot.title=element_blank())
    ) +
    plot_layout(heights=c(1,6))
  
  return(plot_dendroheat)
}

fig_2a_combined_dendro <- make_dendroheatmap_cd4cd8_combined(total_combined_scores_cd4_cd8,
  legend_theme=theme(
      legend.position = c(1, 0.5), 
      legend.justification = c(0, 0.5),
      plot.margin = margin(5.5,66,5.5,44, unit="pt")))
```

### new PCA
```{r}

read.counts[CAR.align == 'TNR8', CAR.align := 'CD30']
read.counts[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']

pca_shared_cols <- c(intersect(
  intersect(
    intersect(names(rel_abund_means),names(cd69.ranked)),
    intersect(names(cyto.ranked.all),names(car_score_means))),
  names(deseq2_combined)))

pc_total_combined <- unique(rbind(
  #deseq2 measurements for CTV1
  deseq2_combined[, pca_shared_cols, with=F],
  
  #new and old cytokine combined with z-scores and p-values
  #unique(copy(cyto.ranked.all)[,  c('sort.group', pca_shared_cols), with=F]),
  unique(copy(cyto.ranked.all)[, padj := p.adj.ta][, value.scale := value.ta][, 
    shared_cols, with=F]),
  
  #cd69 from both donors
  copy(cd69.ranked)[comparison == 'rlog_car_score',
    c('sort.group', pca_shared_cols), with=F],
  
  #car scores for CTV
  copy(car_score_means)[, comparison := 'rlog_car_score'][, 
    c('sort.group', pca_shared_cols), with=F],
  
  #relative abundance/expansion
  copy(rel_abund_means)[comparison == 'abund_rel_baseline', 
    c('sort.group',pca_shared_cols), with=F],

  fill=T))

cast_metrics <- dcast(
  pc_total_combined[!is.na(value.scale) & 
    !(assay == 'CTV3' & !grepl('baseline', comparison))],
  CAR.align ~ assay + k.type + t.type + comparison + sort.group, 
  value.var='value.scale')

cast_metrics[is.na(cast_metrics)] <- 0
#prcomp_metrics <- data.frame(cast_metrics[, -1])
#prcomp_metrics[is.na(prcomp_metrics)] <- 0
combined_pca <- prcomp(cast_metrics[, -1])

# calculate pca stats
pca.dt <- data.table(pc = data.table(colnames(combined_pca$rotation))[, 
    PC := as.integer(gsub("[A-Z]", "", V1))][, PC], 
    sd = combined_pca$sdev, 
    var = combined_pca$sdev^2, 
    var.norm = combined_pca$sdev^2/sum(combined_pca$sdev^2), 
    var.acc = cumsum(combined_pca$sdev^2/sum(combined_pca$sdev^2)))

melt.pca.dt <- melt(
  pca.dt, measure.vars = c("var.norm", "var.acc"), variable.name = "metric")

pca.stats.plot <- ggplot(melt.pca.dt) +
  geom_line(aes(x = pc, y = value, color = metric)) +
  geom_point(aes(x = pc, y = value, color = metric)) +
  scale_x_continuous(limits = c(1, NA)) +
  labs(title = "Fraction of Variance Captured by Principal Components, Every Sort Group", 
       x = "Principal Component", y = "Fraction of Variance") +
  theme_bw()

# project data onto principal components
projected.metrics <- scale(cast_metrics[, setdiff(names(cast_metrics), 
  c("CAR.align")), with = FALSE],
  combined_pca$center, combined_pca$scale) %*% combined_pca$rotation

projected.metrics <- cbind.data.frame(CAR.align=cast_metrics[, CAR.align],
  projected.metrics)

projected.metrics <- data.table(merge(projected.metrics, 
  unique(read.counts[, .(CAR.align, len)]), 
  by = "CAR.align"))

projected.metrics[, CAR.annot := 'none']
projected.metrics[CAR.align %in% names(car_colors), CAR.annot := CAR.align]
projected.metrics[, CAR.annot := factor(
  CAR.annot, levels=c(names(car_colors)[1:7], 'none'))]

pct_exp <- scales::label_percent()(pca.dt[pc %in% c(1,2), var.norm])

# by name
pca.pos.sort.group.plot <- ggplot(projected.metrics,
    aes(x = -PC1, y = -PC2, label = CAR.annot,
      size=(CAR.align %in% names(car_colors)),
      color=CAR.annot)) +
  geom_point() +
  geom_text_repel(data=copy(projected.metrics)[
      CAR.annot == 'none', CAR.annot := ' '],
    size=3, color='grey20', box.padding = 0.5, segment.alpha=0) +
  scale_size_discrete(guide=F) +
  scale_color_manual('',
    labels=c(names(car_colors)[1:7], 'other costims'),
    values=c(car_colors[1:7], 'grey')) +
  labs(
    x=paste0('PC1 (',pct_exp[1],')'),
    y=paste0('PC2 (',pct_exp[2],')')) +
  theme_bw() +
  theme(axis.text = element_blank(), panel.grid=element_blank())

# by size
pca.pos.sort.size.plot <- ggplot(
    projected.metrics[,
      CAR.label := ifelse(CAR.annot == 'none', '', as.character(CAR.align))],
    aes(
      x = -PC1, y = -PC2, label = CAR.label, 
      size=(CAR.align %in% names(car_colors)), 
      fill=len/3)) +
  geom_point(pch=21, color='black') +
  scale_size_manual(guide=F, values=c(2,6)) +
  scale_fill_distiller(palette='RdYlBu') +
  geom_text_repel(data=projected.metrics,
    size=3, color='grey20', box.padding = 0.5, segment.alpha=0) +
  labs(
    x=paste0('PC1 (',pct_exp[1],')'),
    y=paste0('PC2 (',pct_exp[2],')')) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    panel.grid=element_blank(),
    legend.key.width=unit(0.4, 'lines'))

# by type

types <- fread(text="
  CAR.align,Activity,Family
  CRTAM,Costimulatory,IgSF
  CD28,Costimulatory,IgSF
  ICOS,Costimulatory,IgSF
  CXADR,Costimulatory,IgSF
  CD244,Costimulatory,IgSF
  CD7,Costimulatory,IgSF
  NTB-A,Costimulatory,IgSF
  CD200R,Costimulatory,IgSF
  NKR-P1A,Costimulatory,IgSF
  GITR,Costimulatory,TNFRSF
  CD30,Costimulatory,TNFRSF
  4-1BB,Costimulatory,TNFRSF
  CD40,Costimulatory,TNFRSF
  CD96,Unknown,IgSF
  CD300a,Unknown,CD300
  CD300f,Unknown,CD300
  CD2,Mixed,IgSF
  TIM1,Mixed,TIM
  NKG2D,Mixed,Lectin
  CD72,Mixed,Lectin
  DC-SIGN,Mixed,Lectin
  Siglec-3,Mixed,Lectin
  TACI,Mixed,TNFRSF
  BAFF-R,Mixed,TNFRSF
  BCMA,Mixed,TNFRSF
  LAG3,Inhibitory,IgSF
  CRACC,Inhibitory,IgSF
  TIGIT,Inhibitory,IgSF
  BTLA,Inhibitory,IgSF
  LAIR1,Inhibitory,IgSF
  PD1,Inhibitory,IgSF
  CTLA4,Inhibitory,IgSF
  ILT2,Inhibitory,IgSF
  ILT3,Inhibitory,IgSF
  ILT4,Inhibitory,IgSF
  KLRG1,Inhibitory,IgSF
  KIR3DL1,Inhibitory,IgSF
  KIR2DL1,Inhibitory,IgSF
  HAVCR2,Inhibitory,TIM
  TLT-1,Inhibitory,TREM")

pca_type_combined <- merge(types, projected.metrics, by='CAR.align', all=T)
pca_type_combined[,
  CAR.label := ifelse(CAR.annot == 'none', ' ', as.character(CAR.align))]

pca.pos.sort.type.plot <- ggplot(pca_type_combined,
    aes(x = -PC1, y = -PC2, label = CAR.label, shape=Activity, fill=Family,
      size=(CAR.align %in% names(car_colors)))) +
  geom_point() +
  scale_shape_manual(values=c(24,25,23,21)) +
  scale_size_manual(guide=F, values=c(2,5)) +
  guides(
    fill = guide_legend(override.aes = list(shape = 21, size=5)),
    shape = guide_legend(override.aes = list(size=5))) +
  scale_fill_manual(values=c(
    setNames(
      brewer.pal(12,'Paired')[c(5,3,7,9,10)],
      c('CD300','IgSF','Lectin','TIM','TREM')),
    setNames(brewer.pal(5,'Blues')[3],'TNFRSF')
    )) +
  geom_text_repel(data=pca_type_combined,
    size=3, color='grey20', box.padding = 0.4, segment.alpha=0) +
  labs(
    x=paste0('PC1 (',pct_exp[1],')'),
    y=paste0('PC2 (',pct_exp[2],')')) +
  theme_bw() +
  theme(axis.text = element_blank(), panel.grid=element_blank())

#explain components
# map component names back to variables
comp_names <- unique(cbind(
    pc_total_combined[,paste(assay, k.type, t.type, comparison, sort.group, sep='_')],
    pc_total_combined[,paste(assay, k.type, t.type, comparison, sort.group, sep='|')]))
indiv_components <- data.table(matrix(unlist(strsplit(comp_names[,2], '\\|')), ncol=5, byrow=T))
names(indiv_components) <- c('assay', 'k.type', 't.type', 'comparison', 'sort.group')
indiv_components <- data.table(measure=comp_names[,1], indiv_components)

component_rotations <- indiv_components[
  melt.data.table(data.table(combined_pca$rotation, keep.rownames=T), id.vars='rn'),
  on=c(measure='rn')]

indiv_component_plot <- ggplot(
  component_rotations[
    variable %in% c('PC1','PC2')][,
      list(diff_1v2= value[variable == 'PC1'] - value[variable == 'PC2']),
      by=.(assay, comparison, k.type, t.type)][, assay_group := assay][
        assay %in% c('CD69','IFNy','IL2'), assay_group := 'markers'][,
          list(value= sum(diff_1v2)), by=.(t.type, k.type, assay, assay_group)]) + 
  geom_bar(aes(x=assay, y=value, fill=assay_group), color='grey50', stat='identity') +
  facet_grid(k.type + assay_group ~ t.type, scales='free', space='free') +
  geom_hline(yintercept=0) +
  coord_flip() +
  labs(x='Pooled Assay Scores, grouped', y='<- More PC2     More PC1 ->')

plot_rotations <- component_rotations[variable %in% c('PC1','PC2')][,
  `:=`(min_val= min(value), max_val=max(value)), by=.(k.type, assay, t.type, comparison)][,
    measure_group := assay][
      grep('baseline', comparison), measure_group := 'expansion'][
        assay %in% c('CD69','IFNy','IL2'), measure_group := 'markers']

ggplot(plot_rotations) +
  geom_point(aes(x=interaction(assay,comparison), y=-value, color=variable)) +
  geom_linerange(data=plot_rotations[variable=='PC1'], aes(x=interaction(assay,comparison),
      ymin=-min_val, ymax=-max_val, color=ifelse(min_val == value, 'PC1','PC2'))) +
  facet_grid(k.type + measure_group ~ t.type, scales='free', space='free') +
  geom_hline(yintercept=0) + coord_flip()


plot_summary <- copy(plot_rotations)[
  measure_group == 'markers', measure_group := assay][
  measure_group == 'expansion', measure_group := paste(assay, measure_group, sep='.')][,
  list(val_mean=mean(-value)), by=.(measure_group, k.type, t.type, variable)]

plot_summary[, measure_group := factor(
  measure_group,
  levels=c('IL2','IFNy','CD69','CTV3.expansion',
           'CTV2.expansion','CTV1.expansion','CTV2','CTV1'),
  labels=c('IL2','IFNy','CD69','Expansion:\nd0-d24','Expansion:\nd0-d16',
           'Expansion:\nd0-d4','Proliferation:\nd9-d16','Proliferation:\nd0-d3/4'))]

pca_factors <- ggplot(plot_summary[, k.type := factor(k.type, labels=c('CD19-','CD19+'))]) + 
  geom_bar(
    aes(x=measure_group, y=val_mean, fill=variable),
    stat='identity', position=position_dodge(0.7), color='grey30', size=0.5, alpha=0.8, width=0.7) +
  facet_grid(k.type~t.type, scales='free_y', space='free_y') +
  geom_hline(yintercept=0, size=0.5, linetype=2) +
  scale_y_continuous(breaks=c(-0.10, 0, 0.10)) +
  coord_flip() +
  labs(x='Measurement (mean DESeq2 LFC)',y="Principal Component\nContribution (rotation)") +
  theme_minimal() +
  theme(
    panel.border=element_rect(color='black', fill=NA),
    panel.grid=element_blank())

ratio.values <- pca_type_combined[, (max(PC1)-min(PC1))/(max(PC2)-min(PC2))]
```

### expansion over time plot, 8 cars

```{r}

car.abund.order <- c('CD28','4-1BB','BAFF-R','TACI','CD40','CD30','KLRG1')

# add relative baseline measurement if not added
#   (preserve idempotence for debug)

if (!('car.abund.rel.baseline' %in% names(read.counts))) {
  car.abund.set <- read.counts[assay=='baseline' & batch=='prolif2',
            list(car.abund.baseline= car.abund, donor, CAR.align, t.type)]
  
  read.counts <- read.counts[car.abund.set, on=.(donor, CAR.align, t.type)]
  
  read.counts[, car.abund.rel.baseline := car.abund/car.abund.baseline,
              by=.(donor, t.type, batch)]
  
  read.counts[assay == 'baseline',
              car.abund.rel.baseline := 1,
              by=.(batch, donor, t.type, CAR.align)]
}

car.abund.data <- rbind(
  read.counts[k.type == 'pos' & batch != 'post-cytof'], 
  copy(read.counts[batch != 'post-cytof' & assay == 'baseline'])[,
    k.type := 'pos'])[k.type != 'NA'][CAR.align %in% array.list]

car.abund.data[CAR.align == 'TNR8', CAR.align := 'CD30']

car.abund.data[, CAR.align.ord := factor(CAR.align, levels=car.abund.order)]

# add significance
car.abund.data <- (pooled_prolif_p_values[t.type == 'both'][, -'t.type'] %>% 
    add_significance(p.col='padj'))[
      car.abund.data, on='CAR.align']

# add **s as a parsed label

car.abund.order.pval_labels <- unique(
  car.abund.data[order(CAR.align.ord), list(CAR.align.ord, padj.signif)])[,
    paste0('"',car.abund.order,'"^"', padj.signif, '"')]

car.abund.data[,
  CAR.align.ord.pval := factor(CAR.align.ord,
    levels=car.abund.order,
    labels=car.abund.order.pval_labels)]

car.abund.plot <- ggplot(car.abund.data,
    aes(
      x=day, y=log2(car.abund.rel.baseline),
      color= interaction(t.type), group= interaction(k.type, t.type, batch, donor))) + 
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks=c(0,12,24), limits=c(0,24.5)) +
  scale_color_manual('', values=brewer.pal(9, 'YlOrRd')[c(4,6)]) +
  facet_wrap(~CAR.align.ord.pval, ncol=8,
    labeller="label_parsed") + 
  scale_linetype_manual(values=c(2,1)) +
  labs(x='Timepoint (days)', y='Relative Expansion, log2 FC') + 
  geom_hline(linetype=2, yintercept=0) + 
  theme_minimal() +
  theme(
    legend.position='bottom',
    panel.grid = element_blank(),
    panel.border = element_rect(fill=NA),
    strip.background=element_rect(linetype=1, color='white', size=3),
    strip.text = element_text(color='white',face='bold'))

color_car_strips <- function(plot) {

  car.abund.gtable <- ggplot_gtable(ggplot_build(plot))
  strip_both <- which(grepl('strip-', car.abund.gtable$layout$name))
  fills <- car_colors
  k <- 1
  
  for (i in strip_both) {
    j <- which(grepl('rect', car.abund.gtable$grobs[[i]]$grobs[[1]]$childrenOrder))
    car.abund.gtable$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }
  return(car.abund.gtable)
}
```

### Fig 3 assembled

```{r}
fig3 <- plot_grid(
  fig_3_single_dendro,
  plot_grid(
    pca.pos.sort.type.plot + theme(
      axis.title=element_text(size=unit(9,'pt')),
      legend.margin=unit(0,"cm"),
      plot.margin=margin(5,5,5,25, unit='pt'),
      axis.title.y = element_blank(),
      legend.text = element_text(size=unit(7,'pt')),
      legend.title = element_text(size=unit(8,'pt'))),
    color_car_strips(car.abund.plot + theme(
        axis.title=element_text(size=unit(9,'pt')),
        panel.border=element_rect(fill=NA),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.box='horizontal', 
        strip.background = element_rect(colour="white", fill="white"),
        legend.justification=c(1,0),
        legend.position=c(1,0.8))),
    ncol=2, nrow=1, rel_widths = c(1.55,3), labels=c('B','C')),
  rel_heights=c(1.4,1),
  nrow=2, ncol=1, labels=c('A',''))

ggsave(here::here('..','figs','compiled','fig3.newdraft.pdf'), fig3,
       height=7, width=11, useDingbats=FALSE)
```
# new cytokine panels - v1, by donor

```{r}
# cyto.plot.data <- cyto.read.counts.all[k.type == 'pos' & sort.cond != 'neg' &
#   assay %in% c('il2.ifng', 'il2','ifng') &
#   sort.cond %in% c('il2','ifng') &
#   car.sort.group.cells > 400]
# 
# cyto.plot.data[, CAR.axis := factor(paste(CAR.align,donor, sep='|'))]
# # cyto.plot.data[, CAR.axis := factor(
# #   CAR.axis, 
# #   levels=levels(CAR.axis),
# #   labels=gsub('(.*)\\|.*', '\\1', levels(CAR.axis)))]
# 
# #fold-change from mean
# cyto.plot.data[, l2fc.mean := log2(car.bin.pct.incl/mean(car.bin.pct.incl, na.rm=T)),
#   by=.(sort.cond, donor, t.type)]
# 
# cytokine_panel <- ggplot(cyto.plot.data, aes(
#     x = reorder(CAR.axis, l2fc.mean), 
#     y = l2fc.mean, color=interaction(t.type, sort.cond))) +
#   geom_point(size=2.5) +
#   # scale_color_manual(
#   #     values=c('CD4'=cd19_val[2], 'CD8'='grey30')) + 
#   facet_wrap(donor ~ ., scales='free') +
#   scale_y_continuous() +
#   scale_x_discrete(labels= remove_t_type_sep_col) +
#   labs(y = 'Cytokine Secretion,\nLog2 Fold Change from mean') +
#   theme_minimal() +
#   theme(panel.border=element_rect(fill=NA),
#         axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
#         axis.title.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         strip.background = element_rect(colour="white", fill="white"),
#         legend.title=element_blank(),
#         legend.justification=c(1,1))#,
#         #legend.position=c(1,1.03))

# ggsave(here::here('..','figs','pooled','cytokine_plot.pdf'), cytokine_panel,
#        height=3, width=8, useDingbats=FALSE)
# 
# cytokine_panel
```

# new cytokine panels - v2, by cytokine

```{r}
# cyto.plot.data <- cyto.read.counts.all[k.type == 'pos' & sort.cond != 'neg' &
#   assay %in% c('il2.ifng', 'il2','ifng') &
#   sort.cond %in% c('il2','ifng') &
#   car.sort.group.cells > 500]
# cyto.plot.data[CAR.align == 'BTLA.a', CAR.align := 'BTLA']
# 
# 
# cyto.plot.data[, CAR.axis := factor(paste(CAR.align,sort.cond, sep='|'))]
# cyto.plot.data[, CAR.axis := factor(
#   CAR.axis, cytokine_panel
#   levels=levels(CAR.axis),
#   labels=gsub('(.*)\\|.*', '\\1', levels(CAR.axis)))]

#fold-change from mean
# cyto.plot.data[, l2fc.mean := log2(car.bin.pct.incl/mean(car.bin.pct.incl, na.rm=T)),
#   by=.(sort.cond, donor, t.type)]
# 
# lfc.max = 3
# 
# cd4_cd8_colors <- brewer.pal(9, 'YlOrRd')[c(4,6)]
# 
# cyto.plot.data[,
#     l2fc.mean.disp := ifelse(abs(l2fc.mean) < 3,l2fc.mean, lfc.max*.98*sign(l2fc.mean))]
# 
# cytokine_panel <- ggplot(cyto.plot.data, 
#     aes(
#       x = reorder(CAR.axis, l2fc.mean), 
#       y = l2fc.mean.disp, fill=t.type, color=t.type,
#       shape = factor(sign(l2fc.mean.disp)*(abs(l2fc.mean) > 3)))) +
#   geom_point(size=2) +
#   scale_shape_manual(values=c(25,21,24), guide='none') +
#   scale_fill_manual(values=cd4_cd8_colors) +
#   scale_color_manual(values=cd4_cd8_colors) +
#   facet_wrap(factor(sort.cond, labels=c('IFNγ','IL2')) ~ ., scales='free') +
#   geom_hline(yintercept=0, linetype=2, color='grey60') +
#   scale_y_continuous(limits=c(-lfc.max,lfc.max), expand=c(0,0)) +
#   scale_x_discrete(labels= remove_t_type_sep_col) +
#   labs(y = 'Cytokine Secretion,\nLog2 Fold Change from mean') +
#   theme_minimal() +
#   theme(panel.border=element_rect(fill=NA),
#         axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
#         axis.title.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         strip.background = element_rect(colour="white", fill="white"),
#         legend.title=element_blank(),
#         legend.justification=c(1,1))#,
#         #legend.position=c(1,1.03))

# ggsave(here::here('..','figs','pooled','cytokine_plot.pdf'), cytokine_panel,
#        height=3, width=8, useDingbats=FALSE)
# 
# cytokine_panel
```

# new cytokine panels - v3, by cytokine, colored with donor - using this

```{r}
cyto.plot.data <- cyto.read.counts.all[k.type == 'pos' & sort.cond != 'neg' &
  assay %in% c('il2.ifng', 'il2','ifng') &
  sort.cond %in% c('il2','ifng') &
  car.sort.group.cells > 500]
cyto.plot.data[CAR.align == 'BTLA.a', CAR.align := 'BTLA']
cyto.plot.data[, n_measure := .N, by=.(CAR.align, sort.cond)]
cyto.plot.data <- cyto.plot.data[n_measure > 1]

cyto.plot.data[, CAR.axis := factor(paste(CAR.align,sort.cond, sep='|'))]
# cyto.plot.data[, CAR.axis := factor(
#   CAR.axis, cytokine_panel
#   levels=levels(CAR.axis),
#   labels=gsub('(.*)\\|.*', '\\1', levels(CAR.axis)))]

#fold-change from mean
cyto.plot.data[, l2fc.mean := log2(car.bin.pct.incl/mean(car.bin.pct.incl, na.rm=T)),
  by=.(sort.cond, donor, t.type)]

lfc.max = 3

cd4_cd8_colors <- brewer.pal(9, 'YlOrRd')[c(4,6)]

donor_colors = c(
  brewer.pal(11, 'RdBu')[9],
  brewer.pal(11, 'PuOr')[9],
  brewer.pal(11, 'PRGn')[9])

cyto.plot.data[, donor := factor(donor, labels=paste0('Donor ', c('A','B','C')))]

# deal with OOB and CD4/CD8 together
cyto.plot.data[, pt.shape := factor(
  sign(l2fc.mean)*(abs(l2fc.mean) > 3),
  levels=c(-1,0,1),
  labels=c('OOB.lo','IB','OOB.hi'))]
cyto.plot.data[, l2fc.mean.disp := ifelse(
  abs(l2fc.mean) < 3,l2fc.mean, lfc.max*.98*sign(l2fc.mean))]
cyto.plot.data[pt.shape == 'IB', pt.shape := t.type]

pt.shape.types=c('OOB.lo'=25, 'CD4'=21, 'CD8'=22, 'OOB.hi'=24)
pt.shape.types_samesubset=c('OOB.lo'=25, 'CD4'=21, 'CD8'=21, 'OOB.hi'=24)

cyto.plot.data[,
  l2fc.mean.sort.cond := mean(l2fc.mean, na.rm=T),
  by=.(CAR.align, sort.cond)]

do_wilcox_test <- function(df) {
  wilcox_test(df,
    sort.group.cond.rank ~ CAR.align,
    ref.group='all',
    alternative='g') %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
}

car_ranks <- cyto.plot.data[,
      `:=`(
        sort.group.cond.rank= rank(-car.bin.pct.incl),
        CAR.align.sort.cond= factor(CAR.align)),
      by=.(sort.cond, sort.group)][,
        n_measure := .N, by=.(CAR.align)][n_measure > 1][,
            CAR.align := factor(as.character(CAR.align))]

wilcoxon_stats <- car_ranks[, do_wilcox_test(.SD), by='sort.cond']
wilcoxon_stats <- wilcoxon_stats[, list(
  CAR.align=group2, p.adj, p.adj.signif, sort.cond,
  CAR.axis= factor(paste(group2,sort.cond, sep='|')))]

remove_t_type_sep_col_bold_signif <- function (breaks) {
  car_name <- unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))
  
  car_axis_colors <- rep('black', length(car_name))
  car_axis_colors[grep('CD28', car_name)] <- car_colors['CD28']
  car_axis_colors[grep('4-1BB', car_name)] <- car_colors['4-1BB']

  # bold CAR names if colored or if signif in wilcoxon_stats
  bold_cars <- breaks %in% wilcoxon_stats[p.adj.signif != 'ns', CAR.axis]
  car_name[bold_cars] <- 
      paste0('**',car_name[bold_cars],'**')
  
  car_color_span <- paste("<span style = 'color: ",
    car_axis_colors,
    ";'>",
    car_name,
    "</span>", sep = "")
  
  return(car_color_span)
}

cytokine_panel <- ggplot(cyto.plot.data, 
    aes(
      x = reorder(CAR.axis, l2fc.mean), 
      y = l2fc.mean.disp, fill=donor, color=donor,
      shape = pt.shape)) +
  geom_point(size=1.5) +
  geom_point(size=2.5, shape=21, color='grey30', fill=NA,
    aes(y=l2fc.mean.sort.cond)) +
  scale_shape_manual(values=pt.shape.types, breaks=c('CD4','CD8')) +
  scale_fill_manual(values=donor_colors, guide='none') +
  scale_color_manual(values=donor_colors) +
  facet_wrap(factor(sort.cond, labels=c("'IFN'~gamma",'IL2')) ~ ., scales='free', labeller=label_parsed) +
  geom_hline(yintercept=0, linetype=2, color='grey60') +
  scale_y_continuous(limits=c(-lfc.max,lfc.max), expand=c(0,0)) +
  scale_x_discrete(labels= remove_t_type_sep_col_bold_signif) +
  labs(y = 'Relative Cytokine Secretion,\nlog2 FC', 
       color='', shape='') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA),
        axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.box='horizontal', 
        strip.background = element_rect(colour="white", fill="white"),
        legend.justification=c(1,1),
        legend.position=c(0.25,1))

# cytokine_panel_color_subset <- ggplot(cyto.plot.data, 
#     aes(
#       x = reorder(CAR.axis, l2fc.mean), 
#       y = l2fc.mean.disp, fill=t.type, color=t.type,
#       shape = pt.shape)) +
#   geom_point(size=1.5) +
#   geom_point(size=2.5, shape=21, color='grey30', fill=NA,
#     aes(y=l2fc.mean.sort.cond)) +
#   scale_shape_manual(values=pt.shape.types_samesubset, guide='none') +
#   scale_fill_manual(values=cd4_cd8_colors, guide='none') +
#   scale_color_manual(values=cd4_cd8_colors) +
#   facet_wrap(factor(sort.cond, labels=c('IFNγ','IL2')) ~ ., scales='free') +
#   geom_hline(yintercept=0, linetype=2, color='grey60') +
#   scale_y_continuous(limits=c(-lfc.max,lfc.max), expand=c(0,0.01)) +
#   scale_x_discrete(labels= remove_t_type_sep_col_bold_signif) +
#   labs(y = 'Cytokine Secretion,\nLog2 Fold Change from mean', color='Donor') +
#   theme_minimal() +
#   theme(panel.border=element_rect(fill=NA),
#         axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
#         axis.title.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         strip.background = element_rect(colour="white", fill="white"),
#         legend.justification=c(1,1),
#         legend.position=c(0.05,1))

# ggsave(here::here('..','figs','pooled','cytokine_plot.pdf'), cytokine_panel,
#        height=3, width=8, useDingbats=FALSE)
# 
cytokine_panel
```

# new cytokine panels - other, (bad visualization), by raw %

```{r}
# cyto.plot.data <- cyto.read.counts.all[k.type == 'pos' & sort.cond != 'neg' &
#   assay %in% c('il2.ifng', 'il2','ifng') &
#   sort.cond %in% c('il2','ifng') &
#   car.sort.group.cells > 500]
# 
# cyto.plot.data[, CAR.axis := factor(paste(CAR.align,t.type,donor, sep='|'))]
# # cyto.plot.data[, CAR.axis := factor(
# #   CAR.axis, 
# #   levels=levels(CAR.axis),
# #   labels=gsub('(.*)\\|.*', '\\1', levels(CAR.axis)))]
# 
# #fold-change from mean
# cyto.plot.data[, l2fc.mean := log2(car.bin.pct.incl/mean(car.bin.pct.incl, na.rm=T)),
#   by=.(sort.cond, donor, t.type)]
# 
# cytokine_panel <- ggplot(cyto.plot.data, aes(
#     x = reorder(CAR.axis, car.bin.pct.incl), 
#     y = car.bin.pct.incl, color=interaction(sort.cond, donor))) +
#   geom_point(size=2.5) +
#   # scale_color_manual(
#   #     values=c('CD4'=cd19_val[2], 'CD8'='grey30')) + 
#   facet_wrap(t.type ~ donor, scales='free') +
#   scale_y_continuous(labels=percent) +
#   scale_x_discrete(labels= remove_t_type_sep_col) +
#   labs(y = 'Cytokine Secretion,\nLog2 Fold Change from mean') +
#   theme_minimal() +
#   theme(panel.border=element_rect(fill=NA),
#         axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
#         axis.title.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         strip.background = element_rect(colour="white", fill="white"),
#         legend.title=element_blank(),
#         legend.justification=c(1,1))#,
#         #legend.position=c(1,1.03))

# ggsave(here::here('..','figs','pooled','cytokine_plot.pdf'), cytokine_panel,
#        height=3, width=8, useDingbats=FALSE)
# 
# cytokine_panel
```

# new version of fig2 'sprinter' early vs late plot

```{r}
read.counts[, car.abund.log := log10(car.abund)]

#assay_rep_list:
assay_rep_set <- unique(read.counts[
  data.table(
    assay=c('baseline','CTV1','CTV2','CTV3'), 
    prev_assay=c(NA,'baseline','CTV1','CTV2')), on='assay'][, 
  list(assay, sort.group, prev_assay, donor, batch, t.type, k.type)])

#map day0 
assay_rep_set <- rbind(
  assay_rep_set[assay=='baseline'][, k.type := 'pos'], 
  assay_rep_set[assay=='baseline'][, k.type := 'neg'], 
  assay_rep_set[assay!='baseline'])

#merge with prev copy to get prev sort.group correspondence
assay_rep_set <- assay_rep_set[, list(
    donor, batch, t.type, k.type,
    prev_assay=assay,prev.sort.group=sort.group)][
  assay_rep_set,
  on=c('donor','batch','t.type','k.type','prev_assay')]

# use baseline from prolif2 always
assay_rep_set[assay == 'CTV1', 
  prev.sort.group := gsub('prolif1','prolif2',prev.sort.group)]

#merge with prev measurements
prev_measure <- unique(read.counts[, list(
  car.abund.prev=car.abund, prev.sort.group=sort.group, CAR.align)])[
    assay_rep_set, on=c('prev.sort.group')]

#merge with orig read counts
sprinter.read.counts <- prev_measure[, list(
    car.abund.prev, prev.sort.group, CAR.align, sort.group)][
      read.counts, on=c('sort.group','CAR.align')]

#calculate relative prev
sprinter.read.counts[, car.abund.rel.prev := car.abund/car.abund.prev]

# expansion vs proliferation
car_label_set <- c('BAFF-R','CD30','4-1BB','TACI',
                   'CD28','KLRG1','CD40','NTB-A','CD2')

plot_set <- unique(
  sprinter.read.counts[assay %in% c('CTV1', 'CTV2')][, 
    car.lbl := ifelse(
      CAR.align %in% car_label_set, as.character(CAR.align), '')][,
    list(car.lbl, rlog_car_score, car.abund.rel.baseline, 
         car.abund.rel.prev, k.type, assay, t.type, donor, 
         CAR.align, batch, sort.group, avg.divisions)])

early_v_late_indiv_dt <- dcast(
    plot_set, car.lbl+k.type+t.type + donor + CAR.align + batch ~ assay, 
    value.var=c('rlog_car_score','car.abund.rel.baseline','car.abund.rel.prev'))[
      k.type=='pos'][, late_v_early := log2(
        car.abund.rel.prev_CTV1/car.abund.rel.prev_CTV2)]

# early vs late stats
early_v_late_lm_dt <- early_v_late_indiv_dt[,
  data.table(
    summary(lm(late_v_early ~ 0+CAR.align))$coeff, keep.rownames=T)]
setnames(early_v_late_lm_dt, c('rn','Pr(>|t|)'),c('CAR.align','p.val'))
early_v_late_lm_dt[, CAR.align := gsub('CAR.align','',CAR.align)]

# total exp stats
total_exp_lm_dt <- early_v_late_indiv_dt[,
  data.table(
    summary(lm(log2(car.abund.rel.baseline_CTV2) ~ 0+CAR.align))$coeff, keep.rownames=T)]
setnames(total_exp_lm_dt,c('rn','Pr(>|t|)'),c('CAR.align','p.val'))
total_exp_lm_dt[, CAR.align := gsub('CAR.align','',CAR.align)]

sprinter_dt <- merge(
  total_exp_lm_dt,
  early_v_late_lm_dt,
  by=c('CAR.align'),
  suffixes=c(".total",".evl"))

# set most labels to blank
sprinter_dt <- sprinter_dt[,
    car.lbl := ifelse((p.val.total < 0.05 & Estimate.total > 0) | 
    (CAR.align %in% c('4-1BB','CD28','BAFF-R','TACI')), CAR.align, '')]

# add additional blank points to make ggrepel avoid annotation boxes
sprinter_lbls_w_blanks <- rbind(
  sprinter_dt,
  data.table(
    car.lbl='',
    Estimate.evl =rep(c(-1.3,1.3),2),
    Estimate.total =rep(c(-.85,.85), each=2)),
  fill=T)

sprinter_plot <- ggplot(sprinter_dt,
  aes(y=-Estimate.evl, x=Estimate.total, label=car.lbl)) + 
  geom_point(pch=21, aes(
    color=(p.val.evl > 0.05 | car.lbl != ''),
    fill=factor(as.numeric((p.val.evl < 0.05))*sign(Estimate.evl),
      labels=c('later', 'n.s.', 'earlier')),
    size=ifelse(Estimate.total > 0, -log10(p.val.total),0))) +
  scale_color_manual(values=c('white','black'), guide='none') +
  scale_fill_manual(
    values=c(later=hcl.colors(8,'PRGn')[7],ns='white',earlier=hcl.colors(8,'PRGn')[2]),
    na.value='white',
    limits=c('earlier','later')) +
  geom_text_repel(data=sprinter_lbls_w_blanks, seed=10, point.padding=0.1) +
  annotate(geom='label', x=-Inf, y=Inf, label='later/poor', 
    label.r=unit(0,'pt'), size=3, vjust=.99, hjust=.01, color=hcl.colors(8,'PRGn')[7]) +
  annotate(geom='label', x=-Inf, y=-Inf, label='earlier/poor', 
    label.r=unit(0,'pt'), size=3, vjust=.01, hjust=.01, color=hcl.colors(8,'PRGn')[2]) +
  annotate(geom='label', x=Inf, y=Inf, label='later/potent', 
    label.r=unit(0,'pt'), size=3, vjust=.99, hjust=.99, color=hcl.colors(8,'PRGn')[7]) +
  annotate(geom='label', x=Inf, y=-Inf, label='earlier/potent', 
    label.r=unit(0,'pt'), size=3, vjust=.01, hjust=.99, color=hcl.colors(8,'PRGn')[2]) +
  geom_vline(xintercept=0, linetype=2, color='grey60') +
  geom_hline(yintercept=0, linetype=2, color='grey60') +
  scale_size_continuous(range=c(2,6)) +
  #scale_x_continuous(expand=c(0.2,0.1)) +
  scale_y_continuous(expand=c(0.08,0.04)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    legend.position= 'bottom',
    legend.box = "vertical",
    panel.border = element_rect(colour = "black")) +
  labs(
    y='Late vs Early Expansion,\nlog2(d3-14 / d0-3)',
    x='Overall expansion, log2(d0-d14) FC from mean',
    fill='Significant late/early expansion,\np<0.05:',
    size='Overall rel. expansion, -log10(p):')
```

# new fig2c abund with stats

```{r}
library(nlme)

prolif_data_raw <- read.counts[
  (k.type =='pos' | is.na(k.type)) & batch != 'post-cytof'][,
    list(batch, donor, log2.rel.abund= log2(car.abund.rel.baseline), 
      CAR.align, t.type, assay)]
prolif_data_raw[, assay := factor(assay, ordered=T)]

prolif_data_raw[, group := paste(batch, donor, t.type, sep='_')]

new_car_levels  <- prolif_data[, c(
  levels(CAR.align)[!(levels(CAR.align) %in% c('4-1BB','CD28'))], 
  '4-1BB','CD28')]

prolif_data[, CAR.align := factor(CAR.align, new_car_levels)]

with(prolif_data_raw, {
  nestinginfo <- groupedData(log2.rel.abund ~ CAR.align | group, data= prolif_data_raw)
  fit.compsym <- gls(log2.rel.abund ~ factor(CAR.align)*factor(assay),
    data=nestinginfo,
    corr=corCompSymm(, form= ~ 1 | group))
  fit.ar1 <- gls(log2.rel.abund ~ factor(CAR.align)*factor(assay),
    data=nestinginfo,
    corr=corAR1(, form= ~ 1 | group))
  # fit.ar1het <- gls(log2.rel.abund ~ factor(CAR.align)*factor(assay),
  #   data=nestinginfo,
  #   corr=corAR1(, form= ~ 1 | group),
  #   weights=varIdent(form = ~ 1 | assay))
  
  # compare models for AIC and loglik
  print(anova(fit.compsym, fit.ar1)) #, fit.ar1het)) #compares the models
  print(anova(fit.compsym))
  print(anova(fit.ar1))
})

# Unstructured Variance/Covariance Structure gave lowest AIC and loglik, so using that
compute_pairwise_mixed_effects <- function(data, car_choice) {
  glsControl(maxIter=1000, msMaxIter=1000, tolerance=1e-3)
  with(data, tryCatch({
    
    data[, is.car := CAR.align == car_choice]
    print(car_choice)
    nestinginfo <- groupedData(log2.rel.abund ~ is.car | group, data=data)

    fit.ar1 <- gls(log2.rel.abund ~ factor(is.car)*assay, 
      data=nestinginfo,
      corr=corAR1(, form= ~ 1 | group))
        
    return(list(car=car_choice, model=fit.ar1))},
    
    error=function(cond) {
      
      message("Here's the original error message:")
      message(cond)
      return(NA)

    }))
}

models_cd8 <- lapply(prolif_data_raw[,unique(CAR.align)],
  function(x) compute_pairwise_mixed_effects(prolif_data_raw[t.type=='CD8'], x))
for (i in seq(length(models_cd8))) models_cd8[[i]]$t.type = 'CD8'

models_cd4 <- lapply(prolif_data_raw[,unique(CAR.align)],
  function(x) compute_pairwise_mixed_effects(prolif_data_raw[t.type=='CD4'], x))
for (i in seq(length(models_cd4))) models_cd4[[i]]$t.type = 'CD4'

models_combined <- lapply(prolif_data_raw[,unique(CAR.align)],
  function(x) compute_pairwise_mixed_effects(prolif_data_raw, x))
for (i in seq(length(models_combined))) models_combined[[i]]$t.type = 'both'

pooled_prolif_p_values <- data.table(t(sapply(
  c(models_cd8, models_cd4, models_combined),
    function(model_i) c(
      CAR.align=as.character(model_i$car),
      t.type=model_i$t.type,
      pval=as.numeric(anova(model_i$model)$`p-value`[2]),
      coef=as.numeric(model_i$model$coefficients[2])))))
pooled_prolif_p_values[, coef := as.numeric(coef)]
pooled_prolif_p_values[, pval := as.numeric(pval)]
pooled_prolif_p_values[, CAR.align := factor(CAR.align)]
pooled_prolif_p_values[, padj := p.adjust(as.numeric(pval), 'BH'), by='t.type']

#-----------

# library(ggnewscale)
# 
# prolif_data <- read.counts[
#   (k.type =='pos' | is.na(k.type)) & batch != 'post-cytof'][,
#     list(batch, donor, 
#       mean.car.abund.rel.baseline= mean(log2(car.abund.rel.baseline)), 
#       sd.car.abund.rel.baseline= 2^sd(log2(car.abund.rel.baseline))), 
#     by=c('CAR.align', 't.type','assay')]
# 
# prolif_data_combined <- read.counts[
#   (k.type =='pos' | is.na(k.type)) & batch != 'post-cytof'][,
#     list(batch, donor, 
#       mean.car.abund.rel.baseline= mean(log2(car.abund.rel.baseline)), 
#       sd.car.abund.rel.baseline= 2^sd(log2(car.abund.rel.baseline))), 
#     by=c('CAR.align','assay')][, t.type := 'both']
# 
# prolif_data <- rbind(prolif_data_combined, prolif_data)
# 
# # add in mixed linear model p-values
# prolif_data <- unique(prolif_data)[unique(p_values), on=c('CAR.align', 't.type')]
# prolif_data[, signif := padj < 0.01]
# prolif_data[, CAR.align := factor(CAR.align)]
# 
# #reorder CAR names to put CD28 and 41BB in front (at end of levels list)
# new_car_levels  <- prolif_data[, c(
#  levels(factor(CAR.align))[!(levels(CAR.align) %in% c('4-1BB','CD28'))],
#  '4-1BB','CD28')]
# prolif_data[, CAR.align := factor(CAR.align, new_car_levels)]
# 
# prolif_data[, CAR.type := 'other']
# prolif_data[CAR.align == '4-1BB', CAR.type := '4-1BB']
# prolif_data[CAR.align == 'CD28', CAR.type := 'CD28']
# prolif_data[, CAR.type := factor(
#   CAR.type,levels=c('other','4-1BB','CD28'))]
# 
# label_data <- unique(prolif_data[assay=='CTV3', 
#   list(t.type, CAR.align, mean.car.abund.rel.baseline,
#        assay, CAR.type, signif, coef)])[, 
#     car_order := rank(mean.car.abund.rel.baseline), by='t.type'][
#       signif == T | CAR.align %in% c('CD28','4-1BB')]
# 
# #color labels pink vs green
# label_data[, CAR.signif := CAR.type]
# label_data[coef > 0, CAR.topbot :='top']
# label_data[coef < 0, CAR.topbot :='bot']
# label_data[, CAR.topbot := factor(CAR.topbot)]
# 
# # add topbot data from labels to prolif data to color lines
# abund_plot_data <- label_data[,
#   list(CAR.align,t.type,CAR.topbot)][
#     prolif_data[order(CAR.type)], on=c('CAR.align','t.type')]
# abund_plot_data[is.na(CAR.topbot), CAR.topbot := 'other']
# 
# fig2_abundance_over_time <- ggplot(abund_plot_data, aes(
#     y=mean.car.abund.rel.baseline, x=assay,
#     color=CAR.topbot,
#     group=CAR.align,
#     label=CAR.align)) + 
#   # line
#   geom_line(aes(size=CAR.type)) +
#   scale_size_manual('',
#     labels=c('New Receptors', '4-1BB', 'CD28'),
#     values=c(0.5,1.5,1.5), guide=F) +
#   scale_color_manual('',
#     labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
#     values=c(receptor_cols, outlier_cols[1], "#40B549", 'grey30'),
#     guide=F) +
#   #points
#   new_scale('size') +
#   geom_point(aes(size=CAR.type, color=CAR.type, group=CAR.align)) +
#   scale_size_manual('',
#     labels=c('New Receptors', '4-1BB', 'CD28'),
#     values=c(1,2,2), guide=F) +  
#   # text
#   new_scale('color') +
#   geom_text_repel(data=label_data,
#     aes(color=CAR.topbot),
#     nudge_x = 1.2,
#     direction = "y",
#     hjust = 1,
#     size = 3,
#     segment.size = 0.3,
#     segment.color='grey') +
#   scale_color_manual('',
#     labels=c('4-1BB', 'CD28','Top','Bottom'),
#     values=c(receptor_cols, outlier_cols),
#     guide=F) +
#   facet_grid(. ~ t.type) +
#   scale_x_discrete(
#     'Days of continuous CD19 stimulation', 
#     expand = c(0, 0.4, 0, 1.4), 
#     labels=c('Baseline', 'd3-4', 'd14-16', 'd24')) +
#   scale_y_continuous('Relative Expansion,\n log2 FC') +
#   theme_minimal() +
#   theme(panel.border=element_rect(fill=NA))
# 
# 
# fig2_abundance_over_time_vert <- ggplot(
#   abund_plot_data[rev(order(CAR.topbot))], aes(
#     y=mean.car.abund.rel.baseline, x=assay,
#     color=CAR.topbot,
#     group=CAR.align,
#     label=CAR.align)) + 
#   # line
#   geom_line(aes(size=CAR.topbot)) +
#   scale_size_manual('',
#     labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
#     values=c(1.5,1.5,0.5,0.8,0.5), guide=F) +
#   scale_color_manual('',
#     labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
#     values=c(receptor_cols, outlier_cols[1], "#40B549", 'grey40'),
#     guide=F) +
#   #points
#   new_scale('size') +
#   geom_point(aes(size=CAR.topbot, color=CAR.topbot, group=CAR.align)) +
#   scale_size_manual('',
#     labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
#     values=c(2,2,1,1.5,1), guide=F) +
#   # text
#   new_scale('color') +
#   geom_text_repel(data=label_data,
#     aes(color=CAR.topbot),
#     nudge_x = 1.2,
#     direction = "y",
#     hjust = 1,
#     size = 3,
#     segment.size = 0.3,
#     segment.color='grey') +
#   scale_color_manual('',
#     labels=c('4-1BB', 'CD28','Top','Bottom'),
#     values=c(receptor_cols, outlier_cols),
#     guide=F) +
#   facet_grid(t.type ~ .) +
#   scale_x_discrete(
#     'Days of continuous CD19 stimulation', 
#     expand = c(0, 0.4, 0, 1.4), 
#     labels=c('Baseline', 'd3-4', 'd14-16', 'd24')) +
#   scale_y_continuous('Relative Expansion,\n log2 FC') +
#   theme_minimal() +
#   theme(panel.border=element_rect(fill=NA), panel.grid=element_blank())
```

```{r}
prolif_data <- read.counts[
  (k.type =='pos' | is.na(k.type)) & batch != 'post-cytof'][,
    list(batch, donor, 
      mean.car.abund.rel.baseline= mean(log2(car.abund.rel.baseline)), 
      sd.car.abund.rel.baseline= 2^sd(log2(car.abund.rel.baseline))), 
    by=c('CAR.align', 't.type','assay')]

# # add in mixed linear model p-values
prolif_data <- unique(prolif_data)[unique(pooled_prolif_p_values
  [t.type != 'both']), on=c('CAR.align', 't.type')]
prolif_data <- add_significance(prolif_data, p.col='padj')

#reorder CAR names to put CD28 and 41BB in front (at end of levels list)
new_car_levels  <- prolif_data[, c(
  levels(CAR.align)[!(levels(CAR.align) %in% c('4-1BB','CD28'))], 
  '4-1BB','CD28')]

prolif_data[, CAR.align := factor(CAR.align, new_car_levels)]

prolif_data[, CAR.type := 'other']
prolif_data[CAR.align == '4-1BB', CAR.type := '4-1BB']
prolif_data[CAR.align == 'CD28', CAR.type := 'CD28']
prolif_data[, CAR.type := factor(
  CAR.type,levels=c('other','4-1BB','CD28'))]


top_num <- 6

label_data <- unique(prolif_data[assay=='CTV3', 
  list(t.type, CAR.align, mean.car.abund.rel.baseline,
       assay, CAR.type, padj, padj.signif)])[, 
    car_order := rank(mean.car.abund.rel.baseline), by='t.type'][
      car_order %in% c(1:(1+top_num),(40-top_num):40) | 
        CAR.align %in% c('CD28','4-1BB')]

#color labels pink vs green
label_data[, CAR.topbot := CAR.type]
label_data[car_order %in% 1:(1+top_num) & CAR.type == 'other', 
  CAR.topbot :='top']
label_data[car_order %in% (40-top_num):40 & CAR.type == 'other',
  CAR.topbot :='bot']
label_data[, CAR.topbot := factor(CAR.topbot)]
label_data <- label_data[padj.signif != 'ns' | CAR.align %in% c('CD28','4-1BB')]

# add topbot data from labels to prolif data to color lines
abund_plot_data <- label_data[,
  list(CAR.align,t.type,CAR.topbot)][
    prolif_data[order(CAR.type)], on=c('CAR.align','t.type')]
abund_plot_data[is.na(CAR.topbot), CAR.topbot := 'other']


fig2_abundance_over_time_vert <- ggplot(
  abund_plot_data[rev(order(CAR.topbot))], aes(
    y=mean.car.abund.rel.baseline, x=assay,
    color=CAR.topbot,
    group=CAR.align,
    label=CAR.align)) + 
  # line
  geom_line(aes(size=CAR.topbot)) +
  scale_size_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
    values=c(1.5,1.5,0.5,0.8,0.5), guide=F) +
  scale_color_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
    values=c(receptor_cols, outlier_cols[1], "#40B549", 'grey40'),
    guide=F) +
  #points
  new_scale('size') +
  geom_point(aes(size=CAR.topbot, color=CAR.topbot, group=CAR.align)) +
  scale_size_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
    values=c(2,2,1,1.5,1), guide=F) +
  # text
  new_scale('color') +
  geom_text_repel(data=label_data,
    aes(color=CAR.topbot, label=paste0('"',CAR.align,'"^"', padj.signif, '"')),
    parse=T,
    nudge_x = 1.2,
    direction = "y",
    hjust = 1,
    size = 2.7,
    segment.size = 0.3,
    segment.color='grey') +
  scale_color_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom'),
    values=c(receptor_cols, outlier_cols),
    guide=F) +
  facet_grid(t.type ~ .) +
  scale_x_discrete(
    'Days of continuous CD19 stimulation', 
    expand = c(0, 0.4, 0, 1.4), 
    labels=c('Baseline', 'd3-4', 'd14-16', 'd24')) +
  scale_y_continuous('Relative Expansion,\n log2 FC from mean ') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA), panel.grid=element_blank())
```

# t test on ctv3
```{r}

ctv3_abund <- unique(read.counts[assay %in% c('CTV3')][,
    list(car.abund.rel.baseline, k.type, assay, t.type, donor, 
         CAR.align, batch, sort.group, avg.divisions)][
           k.type == 'pos'])
ctv3_abund[, car.abund.rel.baseline := log2(car.abund.rel.baseline)]
ctv3_abund[, nrow := .N, by=c('t.type','CAR.align')]

cd8_pvalues <- data.table(t(sapply(
  unique(ctv3_abund[t.type == 'CD8' & nrow > 1]$CAR.align),
  function (x) list(
    t.type='CD8',
    CAR.align=x,
    pval=t.test(
      ctv3_abund[t.type=='CD8' & CAR.align==x, car.abund.rel.baseline],
      ctv3_abund[t.type=='CD8' & CAR.align != x, car.abund.rel.baseline]
        )$p.value))))

cd4_pvalues <- data.table(t(sapply(
  unique(ctv3_abund[t.type == 'CD4' & nrow > 1]$CAR.align),
  function (x) list(
    t.type='CD4',
    CAR.align=x,
    pval=t.test(
      ctv3_abund[t.type=='CD4' & CAR.align==x, car.abund.rel.baseline],
      ctv3_abund[t.type=='CD4' & CAR.align != x, car.abund.rel.baseline]
        )$p.value))))

total_pvalues <- data.table(t(sapply(
  unique(ctv3_abund$CAR.align),
  function (x) list(
    t.type='total',
    CAR.align=x,
    pval=t.test(
      ctv3_abund[CAR.align==x, car.abund.rel.baseline],
      ctv3_abund[CAR.align != x, car.abund.rel.baseline]
        )$p.value))))

cd8_pvalues[, padj := p.adjust(pval, method= 'BH')]
cd4_pvalues[, padj := p.adjust(pval, method= 'BH')]
total_pvalues[, padj := p.adjust(pval, method= 'BH')]


# aovmodel <- (aov(log2.rel.abund ~ CAR.align*assay + Error(group), data=prolif_data_raw))
# emm <- emmeans(aovmodel, ~ CAR.align)
```

#### Replicate Comparison
```{r}

data.dt <- copy(all.deseq.results.dt)

data.dt[CAR.align == 'TNR8', CAR.align := 'CD30']
data.dt[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']

data.dt[, padj.disp := -log10(padj)]
data.dt[, lfc.disp := log2FoldChange]
data.dt[padj.disp > 10, padj.disp := Inf]
data.dt[abs(lfc.disp) > 5, lfc.disp := sign(lfc.disp) * Inf]

data.dt[, CAR.type := 'other']
data.dt[CAR.align == '4-1BB', CAR.type := '4-1BB']
data.dt[CAR.align == 'CD28', CAR.type := 'CD28']
data.dt[, CAR.type := factor(
  CAR.type,levels=c('other','4-1BB','CD28'))]

max_pval_y <- 6.5

#CD28 and 4-1BB Colors
receptor_cols <- RColorBrewer::brewer.pal(9, 'YlGnBu')[c(6,8)]

# label significant hits
data.dt[, CAR.type.size := CAR.type]
data.dt[CAR.type == 'other' & 
    ((padj < 0.05 & log2FoldChange > 0) | 
    (padj < 0.05 & log2FoldChange > 0)), 
  CAR.type.size := 'other_sig']

data.dt[, CAR.label.sig := '']
data.dt[CAR.type.size != 'other',
    CAR.label.sig := CAR.align]

```

#### CD4 vs CD8s (new version, w/ mean prolif and diff)

```{r, fig.width=7, fig.height=4}

cast_4v8 <- dcast(data.dt[
    k.type == 'pos' & 
        ((ref_set == 'baseline' & test_set == 'A' & assay != 'CTV1') | 
        (ref_set == 'CD' & test_set == 'A' & assay == 'CTV1'))], 
      CAR.align + assay + CAR.type ~ t.type, 
    value.var = c("log2FoldChange", "padj.disp"))

# rename facets
cast_4v8[, assay := factor(assay, 
    labels=assay_fig_labels)]

# label significant hits
cast_4v8[, CAR.type.size := CAR.type]
cast_4v8[
    ((padj.disp_CD4 > -log10(0.05) & log2FoldChange_CD4 > 0) | 
    (padj.disp_CD8 > -log10(0.05) & log2FoldChange_CD8 > 0)), 
  CAR.type.size := 'other_sig']

cast_4v8[, CAR.label.sig := '']
cast_4v8[CAR.type.size != 'other', CAR.label.sig := CAR.align]

cast_4v8[, mean_lfc := rowMeans(cbind(log2FoldChange_CD4, log2FoldChange_CD8))]
cast_4v8[, rel_8 := log2FoldChange_CD8 - log2FoldChange_CD4]

cast_4v8[CAR.type.size == 'other_sig' & xor(padj.disp_CD4 > -log10(0.05), 
  padj.disp_CD8 > -log10(0.05)), CAR.type.size := 'other_sig_diff']

# add false points to get ggrepel to avoid plotting underneath the skewed lbls
cast_4v8_w_blanks <- rbind(
  cast_4v8,
  data.table(CAR.label.sig='', rel_8 =rep(c(-1,1),3), mean_lfc=rep(-1,6),
   assay = rep(levels(cast_4v8$assay),each= 2)),
  fill=T)

prolif_4v8_rel_vert <- ggplot(cast_4v8_w_blanks[order(CAR.type)], 
  aes(y=rel_8, x=mean_lfc,
      color=CAR.type,
      label=CAR.label.sig,
      size=CAR.type.size)) + 
  annotate(geom='label', x=-Inf, y=Inf, label='CD8 Skewed', 
    label.r=unit(0,'pt'), size=3, vjust=.99, hjust=.01) +
  annotate(geom='label', x=-Inf, y=-Inf, label='CD4 Skewed', 
    label.r=unit(0,'pt'), size=3, vjust=.01, hjust=.01) +
  geom_point() +
  facet_grid(assay ~ ., scales='free_x') +
  scale_color_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28'),
    values=c('grey20', receptor_cols), guide=F) +
  scale_size_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28', 'New Receptors', 'New Receptors'),
    values=c(0.8,0.8,0.8,2.5,2.5), guide=F) +
  scale_x_continuous(limits=c(-2.5,3.25), breaks=c(-2:3)) +
  scale_y_continuous(limits=c(-2.25,1.9)) +
  geom_hline(yintercept=0, linetype='dashed', alpha=.7) +
  geom_text_repel(size=3, point.padding=0.3, segment.color='grey', max.overlaps=20) +
  theme_minimal() + 
  labs(y='log2 CD8:CD4 proliferation ratio', 
       x='Rel. proliferation/expansion, log2 FC from mean') +
  theme(panel.border=element_rect(fill=NA), panel.grid=element_blank()) +
  guides(size=F)

ggsave(here::here('..','figs','pooled','mixed_4v8_rel.pdf'), prolif_4v8_rel,
       height=2.5, width=7, useDingbats=FALSE)

```

# compile panels, first pass

```{r}

plot_a <- plot_grid(
  barplot_stacked_plot_horiz+theme(legend.position = "none"), 
  (barplot_stacked_multi_plot / plot_spacer()) + plot_layout(heights=c(4.5,1)), 
  rel_widths=c(5,1),
  ncol = 2)

plot_b <- cytokine_panel + 
  guides(shape = guide_legend(override.aes = list(fill = 'grey30') )) +
  theme(legend.justification = c(0, 1), legend.position = c(0, 1),
        legend.margin = margin(t=0, unit='cm'),
        plot.margin = unit(c(1,1,1,0.8), "cm"),
        panel.grid=element_blank())

plot_c <- fig2_abundance_over_time_vert + 
  theme(panel.grid=element_blank())

plot_d <- sprinter_plot + 
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    legend.position= 'bottom',
    legend.text = element_text(size=9, hjust=0),
    legend.title = element_text(size=9, hjust=0),
    legend.box = "vertical",
    panel.border = element_rect(colour = "black"),
    plot.margin = unit(c(0.2,0.2,0.2,0.2), "cm"),
    legend.title.align=0,
    legend.margin=margin(t = -0.3, unit='cm'),
    legend.box.just = "left",
    legend.justification = 'left')

plot_e <- prolif_4v8_rel_vert + theme(
  plot.margin = unit(c(0,1,0.5,0.8), "cm"),
  panel.grid=element_blank())

plot_abcde <- plot_grid(
  plot_a,
  plot_b,
  plot_grid(
    plot_c, plot_d, plot_e, labels=c('C','D','E'),
    ncol=3, rel_widths=c(1.1, 0.9, 1)), ncol=1, 
  nrow=3, 
  rel_heights=c(0.8, 0.9, 1.3),
  labels=c('A','B',''))

ggsave(here::here('..','figs','compiled','fig2.newdraft.pdf'), plot_abcde,
       height=10, width=11, useDingbats=FALSE)  


```
# Add to supplemental - compare abundance and length to expansion and proliferation

```{r}
# mean measures by car
car_len_abund_comparisons <- read.counts[
  k.type == 'pos'][,
  `:=`(
    mean_rlog_car_score= mean(rlog_car_score[assay=='CTV1'], na.rm=T),
    meanlog2.abund.rel.baseline= mean(log2(car.abund.rel.baseline[assay=='CTV3']), na.rm=T),
    baseline.car.abund= mean(log2(baseline.car.abund), na.rm=T)),
  by=c('CAR.align')]


d3_ctv <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=mean_rlog_car_score, x= baseline.car.abund)) + 
  labs(
    y='CTV score, d3',
    x='log2 initial abundance',
    title='Initial Library Abundance vs d3 Prolif. Score')

d24_exp <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=meanlog2.abund.rel.baseline, x= baseline.car.abund)) + 
  labs(
    y='Relative Expansion, d24',
    x='log2 initial abundance',
    title='Initial Library Abundance vs d24 Rel. Expansion')

d3_ctv_len <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=mean_rlog_car_score, x= log2(len))) + 
  labs(
    y='CTV score, d3',
    x='log2 domain length',
    title='Domain Length vs d3 Prolif. Score')

d24_exp_len <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=meanlog2.abund.rel.baseline, x= log2(len))) + 
  labs(
    y='Relative Expansion, d24',
    x='log2 domain length',
    title='Domain Length vs d24 Rel. Expansion')

len_abund_vs_outcome <- (d3_ctv | d24_exp) / (d3_ctv_len | d24_exp_len)

```