---
title: "fig_1_2_3_compiled.Rmd"
author: "Daniel Goodman"
date: "4/7/2020"
output:
  html_document:
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, 
  output_dir = here::here('..','html','pooled'))})
---

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(ggrepel)
library(RColorBrewer)
library(DESeq2)
#library(rnaseqGene)
library(plotly)
library(Rtsne)
library(scales)
library(dplyr)
library(ggtext)
library(rstatix)
library(ggnewscale)
library(dendextend)
library(ggdendro)
library(cowplot)
library(gtable)
library(grid)

rm(list = ls())

setDTthreads(8)

# get local path to s3 dir and other local settings
source(here::here('r','local_settings.R'))

load(file=file.path(s3.data.dir, 'pooled_deseq2_analysis_data.Rdata'))
load(file.path(s3.data.dir, 'pooled_cytokine_repeat_data.Rdata'))

# set this to true to re-run, else will load from s3
rerun_deseq <- F
reload_data <- F

# arrayed list of CARs
array.list <- c('BAFF-R','TNR8','4-1BB','TACI','CD28','KLRG1','CD40')

# load colors
source(here::here('r','fig_colors.R'))

`%ni%` <- Negate(`%in%`)

names(car_colors) <- c(
  "CD28", "4-1BB", "BAFF-R", "TACI", "CD40", "CD30",
  "KLRG1", "zeta", "Unt", "None")

# map day 0 to both plus and minus
map_day_0 <- function(df) {
  return(rbind(
    df[k562!='none'],
    df[k562=='none'][, k562 := 'cd19+'],
    df[k562=='none'][, k562 := 'cd19-']
  ))
}

```

### Fig3: deSeq2 normalization followed by CAR score calculation

```{r}
all.deseq.results.dt[CAR.align == 'TNR8', CAR.align := 'CD30']
all.deseq.results.dt[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']

all.deseq.results.dt[, padj.disp := -log10(padj)]
all.deseq.results.dt[, lfc.disp := log2FoldChange]
all.deseq.results.dt[padj.disp > 10, padj.disp := Inf]
all.deseq.results.dt[abs(lfc.disp) > 5, lfc.disp := sign(lfc.disp) * Inf]

data.dt <- all.deseq.results.dt

data.dt[, CAR.type := 'other']
data.dt[CAR.align == '4-1BB', CAR.type := '4-1BB']
data.dt[CAR.align == 'CD28', CAR.type := 'CD28']
data.dt[, CAR.type := factor(
  CAR.type,levels=c('other','4-1BB','CD28'))]
```

### Fig3: Single heatmap, CD4 plus CD8.

```{r fig.height=4, fig.width=2.5, units='inches'}

# MAD: constant * median(abs(x-median))
# robust z-scores: x - median / median(abs(x - median))
robust_z_score = function(value) 0.6745*(value - median(value))/ mad(value)

ggnormal_dist <- stat_function(
  fun = dnorm, args = list(mean = 0, sd = 1),
  col = "#1b98e0", size = 1, linetype=2)

# stouffer's method: z score over sqrt of number of values
stouffer.z <- function(z) sum(z, na.rm=T) / sqrt(sum(!is.na(z)))
z.to.p.adj.bh <- function(z) p.adjust(pnorm(abs(z), lower.tail=F), "BH")

#current stats workflow is:
# get a Z-score for every group
# combine Z scores using stouffers
# use Benjamani Hochberg to correct for multiple hypotheses

##deseq2
all.deseq.results.dt[, comparison := paste(test_set,ref_set,sep= '.')]

deseq2_combined <- all.deseq.results.dt[,
    comparison := paste(test_set, ref_set, sep='.')][,
    combined_var := paste(comparison, assay)][, value := log2FoldChange]

deseq2_combined[, value.scale := scale(value), by = .(t.type, k.type, assay, comparison)]

deseq2_combined[,
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]

ggplot(deseq2_combined[k.type == 'pos' & comparison %in% c('A.baseline')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align, value.scale),
  color=interaction(assay, comparison), size=padj.disp > -log(0.05))) +
  facet_wrap(~t.type)

## rlog car score
############################

# combine z scores using stouffer's method: z score over sqrt of number of values
car_score_means <- unique(read.counts[assay != 'baseline' & batch != 'post-cytof',
    list(rlog_car_score, CAR.align, assay, t.type, k.type, donor, sort.group, batch)])[,
  value.scale := scale(rlog_car_score), by=sort.group]
car_score_means[,
    value.t.type := stouffer.z(unique(value.scale)),
      by=.(CAR.align, t.type, k.type, assay)]
car_score_means[,
    value.overall := stouffer.z(unique(value.scale)),
      by=.(CAR.align, k.type, assay)]

car_score_means[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, k.type, assay)]
car_score_means[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, k.type, assay)]

ggplot(car_score_means[k.type == 'pos' & assay %in% c('CTV1','CTV2')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(car_score_means[k.type == 'pos' & assay %in% c('CTV1','CTV2')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(car_score_means[assay != 'CTV3' & k.type == 'pos']) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)

car_score_means[, 
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]
car_score_means[, comparison := 'rlog_car_score']
car_score_means[, 
  combined_var := paste(assay,comparison, sep='.'), by=.(t.type)]

# fc in proliferation
##########################

rel_abund_means <- read.counts[assay != 'baseline' & batch != 'post-cytof'][,
  value.scale := scale(log(car.abund.rel.baseline)), by=sort.group]
rel_abund_means[,
    value.t.type :=  stouffer.z(unique(value.scale)), 
      by=.(CAR.align, t.type, k.type, assay)]
rel_abund_means[,
    value.overall :=  stouffer.z(unique(value.scale)), 
      by=.(CAR.align, k.type, assay)]

rel_abund_means[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, k.type, assay)]
rel_abund_means[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, k.type, assay)]


rel_abund_means[, comparison := 'abund_rel_baseline']
rel_abund_means[, 
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]
rel_abund_means[, 
  combined_var := paste(assay,comparison, sep='.'), by=.(t.type)]

ggplot(rel_abund_means[k.type == 'pos' & assay %in% c('CTV2','CTV3')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(rel_abund_means[k.type == 'pos' & assay %in% c('CTV2','CTV3')]) + 
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(rel_abund_means[assay != 'CTV3' & k.type == 'pos']) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)


### CD69 
cd69.ranked <- melt(
  read.counts[batch == 'post-cytof' & assay == 'CD69'], 
  measure.vars = c('min.max.ratio.norm','all.max.ratio.norm','rlog_car_score'),
  variable.name='comparison')[bin == 'A' & k.type == 'pos']
cd69.ranked[, value.scale := scale(value),
  by=.(sort.group,comparison,t.type)]
cd69.ranked <- cd69.ranked[comparison=='rlog_car_score']

cd69.ranked[, value.overall :=  stouffer.z(unique(value.scale)), 
  by=.(CAR.align, k.type, assay, comparison)]

cd69.ranked[, value.t.type :=  stouffer.z(unique(value.scale)), 
  by=.(CAR.align, t.type, k.type, assay, comparison)]

cd69.ranked[, 
  car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]
cd69.ranked[, 
  combined_var := paste(assay,comparison, sep='.'), by=.(t.type)]

cd69.ranked[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, k.type, assay)]
cd69.ranked[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, k.type, assay)]

ggplot(cd69.ranked) + geom_point(aes(x=value.scale,  y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type))) + geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type)), pch=21, size=2) + facet_wrap(~k.type+t.type+comparison) + geom_vline(xintercept=c(-1.96,1.96))

ggplot(cd69.ranked) + geom_point(aes(x=value.scale,  y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type))) + geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), color=interaction(comparison,t.type)), pch=21, size=2) + facet_wrap(~k.type+comparison) + geom_vline(xintercept=c(-1.96,1.96))

ggplot(cd69.ranked) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(cd69.ranked) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(cd69.ranked) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)

### Cytokines - IL2 and IFNG with new and old donors

old.read.counts <- read.counts[assay %in% c('IL2','IFNy')]

old.read.counts[assay == 'IFNy', assay := 'ifng']
old.read.counts[assay == 'IL2', assay := 'il2']

old.read.counts[assay == 'ifng', sort.cond := ifelse(bin %in% c('A','B'), 'neg','ifng')]
old.read.counts[assay == 'il2', sort.cond := ifelse(bin %in% c('A','B'), 'neg','il2')]

old.read.counts <- old.read.counts[, list(
    CAR.align,
    cell.count,
    bin.pct,
    car.bin.pct,
    car.bin.cells= cell.count*car.bin.read.pct),
  by=.(donor, assay, bin, sort.cond, k.type, t.type)]

old.read.counts <- old.read.counts[, list(
  cell.count= cell.count[1],
  car.bin.cells = sum(car.bin.cells),
  bin.pct= bin.pct[1],
  car.bin.pct.incl= sum(car.bin.pct),
  car.bin.pct= sum(car.bin.pct)),
  by=.(sort.cond, CAR.align, donor, assay, t.type, k.type)]

old.read.counts[, sort.group := paste(donor, assay, t.type, k.type, sep='_')]
old.read.counts[, car.sort.group.cells := sum(car.bin.cells), 
    by=.(CAR.align, sort.group)]

shared_cols <- intersect(
  names(old.read.counts), 
  names(cyto.read.counts.summed))

cyto.read.counts.all <- rbind(
  cyto.read.counts.summed[, ..shared_cols],
  old.read.counts[, ..shared_cols])

cyto.ranked.all <- cyto.read.counts.all[
  k.type == 'pos' & assay %in% c('il2.ifng', 'il2','ifng') &
  sort.cond %in% c('il2','ifng') &
  car.sort.group.cells > 500]

cyto.ranked.all[, value.scale := scale(
  log(car.bin.pct.incl)), by=.(sort.cond, donor, t.type, k.type)]

cyto.ranked.all[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']
cyto.ranked.all[CAR.align == 'BTLA.a', CAR.align := 'BTLA']
cyto.ranked.all[, assay := factor(sort.cond, labels=c('IFNy','IL2'))]
cyto.ranked.all <- cyto.ranked.all[CAR.align %ni% c('BTLA.b','BTLA.a')]

cyto.ranked.all[, group := paste(sort.cond,t.type,k.type)]
cyto.ranked.all[, comparison := 'pct_pos']
cyto.ranked.all[, combined_var := paste(comparison, sort.cond)]
cyto.ranked.all[, car.axis := paste(CAR.align,t.type,k.type, sep='|'), by=.(t.type)]

cyto.ranked.all[, value.t.type := stouffer.z(value.scale), by=.(CAR.align, t.type)]
cyto.ranked.all[, value.assay := stouffer.z(value.scale), by=.(CAR.align, assay)]
cyto.ranked.all[, value.ta := stouffer.z(value.scale), by=.(CAR.align, t.type, assay)]
cyto.ranked.all[, value.overall := stouffer.z(value.scale), by=.(CAR.align)]

cyto.ranked.all[, p.adj.t := p.adjust(2*pnorm(abs(value.t.type), lower.tail=F), "BH"),
  by=.(CAR.align, t.type)]

cyto.ranked.all[, p.adj.a := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align, assay)]

cyto.ranked.all[, p.adj.ta := p.adjust(2*pnorm(abs(value.ta), lower.tail=F), "BH"),
  by=.(CAR.align, t.type, assay)]

cyto.ranked.all[, p.adj.o := p.adjust(2*pnorm(abs(value.overall), lower.tail=F), "BH"),
  by=.(CAR.align)]


ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.overall, y=reorder(CAR.align, value.scale), 
      color=interaction(assay), size=p.adj.o < 0.05), pch=21) 

ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.t.type, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.t < 0.05), pch=21) +
  facet_wrap(~t.type)

ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.assay, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.a < 0.05), pch=21) +
  facet_wrap(~sort.cond)

ggplot(cyto.ranked.all) +
  geom_point(aes(x=value.scale, y=reorder(CAR.align,value.scale), 
      color=interaction(t.type, donor, assay))) + 
  geom_point(aes(x=value.ta, y=reorder(CAR.align, value.scale), 
      color=interaction(assay, t.type), size=p.adj.ta < 0.05), pch=21) +
  facet_wrap(~t.type+assay)

ggplot(cyto.ranked.all) + 
  geom_density(aes(x=value.scale)) + 
  ggnormal_dist + 
  geom_errorbar(aes(x=value.scale, ymin=0, ymax=0.01), size=0.1)

### Combined columns, also pvalue
shared_cols <- c(intersect(
  intersect(
    intersect(names(rel_abund_means),names(cd69.ranked)),
    intersect(names(cyto.ranked.all),names(car_score_means))),
  names(deseq2_combined)), 'padj')

total_combined <- unique(rbind(
  #deseq2 measurements for CTV1
  deseq2_combined[, shared_cols, with=F],
  
  #new and old cytokine combined with z-scores and p-values
  unique(copy(cyto.ranked.all)[, padj := p.adj.ta][, value.scale := value.ta][, 
    shared_cols, with=F]),
  
  #cd69 from both donors
  copy(cd69.ranked)[, padj := p.adj.t][
    comparison == 'rlog_car_score', shared_cols, with=F],
  
  #car scores for CTV
  copy(car_score_means)[, padj := p.adj.t][, 
    comparison := 'rlog_car_score'][, shared_cols, with=F],
  
  #relative abundance/expansion
  copy(rel_abund_means)[, padj := p.adj.t][, value.scale := value.t.type][
    comparison == 'abund_rel_baseline', shared_cols, with=F],
  
  use.names=F))

```

### Fig3: Separate CD4 and CD8 heatmaps.
```{r fig.height=8, fig.width=6}
remove_t_type_sep_col_all_cars <- function (breaks) {
  car_name <- unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))
  
  car_axis_colors <- rep('black', length(car_name))
  for (car_i in names(car_colors)) {
    if (car_i %in% car_name) 
      car_axis_colors[(car_i == car_name)] <- car_colors[car_i]
  }

  # bold CAR names if colored
  car_name[car_axis_colors != 'black'] <- paste0('**',car_name[car_axis_colors != 'black'],'**')
  
  car_color_span <- paste("<span style = 'color: ",
    car_axis_colors,
    ";'>",
    car_name,
    "</span>", sep = "")
  
  return(car_color_span)
}

make_dendroheatmap <- function(
  df, title,
  legend_theme=theme(legend.position = "none"), boxnumbers=F)
{
  
  # Make dendrogram
  car_mean_cast <- dcast(
    df[k.type == 'pos'], 
    CAR.align ~ assay + t.type, 
    value.var='value.scale')
  
  car_dendro_m <- as.matrix(car_mean_cast[, -1])
  rownames(car_dendro_m) <- unlist(car_mean_cast[,1])
  car_dendro <- as.dendrogram(hclust(d = dist(x = car_dendro_m)))
  
  # order cars based only on significant values
  car_order <- rownames(car_dendro_m)[
    order(rowMeans(car_mean_cast * as.numeric(dcast(
      df[k.type == 'pos'], 
      CAR.align ~ assay + t.type, 
      value.var='padj')[, -1] < 0.05), na.rm=T))]

  car_dendro <- rotate(car_dendro, car_order)

  # Create dendrogram plot
  dendro_vars_plot <- ggdendrogram(data = car_dendro, rotate = FALSE) + 
    theme(axis.text.y = element_text(size = 6))
  
  # row order
  df$CAR.align <- factor(
      df$CAR.align,
      levels = labels(car_dendro))
  
  #manual x labels
  y_lab_man <- c(
    IFNy.CD4='IFNy 18h', IL2.CD4='IL2 18h',
    CD69.CD4='CD69 18h', CD69.CD8='CD69 18h',
    IFNy.CD8='IFNy 18h', IL2.CD8='IL2 18h',
    CTV1.CD4='Cell divisions, d0-d3/d4', CTV1.CD8='Cell divisions, d0-d3/d4',
    CTV2.CD4='Cell divisions, d9-14/16', CTV2.CD8='Cell divisions, d9-14/16',
    CTV3.CD4='Total expansion d0-24', CTV3.CD8='Total expansion d0-24'
    )
  
  # heatmap
  var_heatmap <-   ggplot(df[,
      t.type := factor(t.type, levels=c('CD8','CD4'))][,
        assay := factor(assay, levels= c('IL2','IFNy','CD69','CTV1','CTV2','CTV3'))]) + 
    geom_tile(aes(y = interaction(assay,t.type), x = CAR.align, 
                  fill = value.scale), colour = "black", size = .20, alpha=ifelse(boxnumbers, 0.5, 1)) + 
    geom_tile(data=df[padj < 0.05],
      aes(y = interaction(assay,t.type), x = CAR.align),
      fill=NA,
      color='black', size=0.5,
      alpha=ifelse(boxnumbers, 0.5, 1)) + 
    scale_fill_distiller('Z-score', 
      palette='PiYG', limits=c(-4,4), oob=scales::squish, direction=1) +
    scale_y_discrete(expand=c(0,0),
      labels=y_lab_man) +
    scale_x_discrete(labels= remove_t_type_sep_col_all_cars, expand=c(0,0)) +
    labs(title=title) + 
    facet_grid(t.type~., scales='free', space='free', switch='y') + 
    coord_cartesian(clip="off") +
    theme(
        strip.placement = "outside",
        axis.text.x = element_markdown(angle = 90, hjust = 1, size=unit(8,'pt')),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        plot.margin=margin(-19,5.5,5.5,5.5, 'points'),
        strip.background = element_blank(),
        panel.background = element_rect(fill='grey90', color='black', size=0.2),
        panel.grid = element_blank(),
        axis.title.x = element_blank())
  
    if (boxnumbers) {
      var_heatmap <- var_heatmap + 
        geom_text(aes(x = interaction(assay,t.type), y = CAR.align, 
          label=round(value.scale,1)), colour = "black")
    }
  
  var_heatmap <- var_heatmap + legend_theme
  
  # col dendro
  dendro_data_col <- dendro_data(car_dendro, type = "rectangle")
  
  dendro_col <- axis_canvas(var_heatmap, axis = "x") +
    geom_segment(data = segment(dendro_data_col),
        aes(x = x, y = y*0.99, xend = xend, yend = yend*0.99))
  # plot_dendroheat <- insert_xaxis_grob(var_heatmap, 
  #   dendro_col, dendro_units, position = "top")
  
  plot_dendroheat <- (
    dendro_col+theme(plot.margin = margin(10,0,0,0))) / 
    (var_heatmap + theme(
        plot.margin= margin(0,40,0,0), panel.spacing= unit(5,'pt'), plot.title=element_blank())
    ) +
    plot_layout(heights=c(1,6))
  
  return(plot_dendroheat)
}

chosen_metrics <- total_combined[k.type == 'pos' &
  ((assay %in% c('CTV1','CTV2') & comparison == 'AB.D') |
    (assay %in% c('CTV3') & comparison == 'abund_rel_baseline') |
    (assay %in% c('IL2','IFNy') & comparison == 'pct_pos') |
      (!(assay %in% c('CTV1','CTV2','CTV3')) & comparison == 'rlog_car_score'))]

fig_3_single_dendro <- make_dendroheatmap(
  df=chosen_metrics,
  title='',
  legend_theme=theme(
      legend.position = c(1, 0.5), 
      legend.justification = c(0, 0.5),
      plot.margin = margin(5.5,66,5.5,44, unit="pt")))

fig_3_single_dendro
```

## Fig2: abundance over time stats (need for fig2 and 3)

```{r}
library(nlme)

prolif_data_raw <- read.counts[
  (k.type =='pos' | is.na(k.type)) & batch != 'post-cytof'][,
    list(batch, donor, log2.rel.abund= log2(car.abund.rel.baseline), 
      CAR.align, t.type, assay)]
prolif_data_raw[, assay := factor(assay, ordered=T)]

prolif_data_raw[, group := paste(batch, donor, t.type, sep='_')]

new_car_levels  <- prolif_data[, c(
  levels(CAR.align)[!(levels(CAR.align) %in% c('4-1BB','CD28'))], 
  '4-1BB','CD28')]

prolif_data[, CAR.align := factor(CAR.align, new_car_levels)]

# compare models based on AIC/BIC/loglik
with(prolif_data_raw, {
  nestinginfo <- groupedData(log2.rel.abund ~ CAR.align | group, 
      data= prolif_data_raw)
  fit.compsym <- gls(log2.rel.abund ~ factor(CAR.align)*factor(assay),
    data=nestinginfo,
    corr=corCompSymm(, form= ~ 1 | group))
  fit.ar1 <- gls(log2.rel.abund ~ factor(CAR.align)*factor(assay),
    data=nestinginfo,
    corr=corAR1(, form= ~ 1 | group))

  # compare models for AIC and loglik
  print(anova(fit.compsym, fit.ar1)) #, fit.ar1het)) #compares the models
  print(anova(fit.compsym))
  print(anova(fit.ar1))
})

# AR1 Variance/Covariance Structure gave lowest AIC and loglik, so using that
compute_pairwise_mixed_effects <- function(data, car_choice) {
  glsControl(maxIter=1000, msMaxIter=1000, tolerance=1e-3)
  with(data, tryCatch({
    
    data[, is.car := CAR.align == car_choice]
    print(car_choice)
    nestinginfo <- groupedData(log2.rel.abund ~ is.car | group, data=data)

    fit.ar1 <- gls(log2.rel.abund ~ factor(is.car)*assay, 
      data=nestinginfo,
      corr=corAR1(, form= ~ 1 | group))
        
    return(list(car=car_choice, model=fit.ar1))},
    
    error=function(cond) {
      
      message("Here's the original error message:")
      message(cond)
      return(NA)

    }))
}

models_cd8 <- lapply(prolif_data_raw[,unique(CAR.align)],
  function(x) compute_pairwise_mixed_effects(prolif_data_raw[t.type=='CD8'], x))
for (i in seq(length(models_cd8))) models_cd8[[i]]$t.type = 'CD8'

models_cd4 <- lapply(prolif_data_raw[,unique(CAR.align)],
  function(x) compute_pairwise_mixed_effects(prolif_data_raw[t.type=='CD4'], x))
for (i in seq(length(models_cd4))) models_cd4[[i]]$t.type = 'CD4'

models_combined <- lapply(prolif_data_raw[,unique(CAR.align)],
  function(x) compute_pairwise_mixed_effects(prolif_data_raw, x))
for (i in seq(length(models_combined))) models_combined[[i]]$t.type = 'both'

pooled_prolif_p_values <- data.table(t(sapply(
  c(models_cd8, models_cd4, models_combined),
    function(model_i) c(
      CAR.align=as.character(model_i$car),
      t.type=model_i$t.type,
      pval=as.numeric(anova(model_i$model)$`p-value`[2]),
      coef=as.numeric(model_i$model$coefficients[2])))))

pooled_prolif_p_values[, coef := as.numeric(coef)]
pooled_prolif_p_values[, pval := as.numeric(pval)]
pooled_prolif_p_values[, CAR.align := factor(CAR.align)]
pooled_prolif_p_values[, padj := p.adjust(as.numeric(pval), 'BH'), by='t.type']
```

### Fig3: expansion over time plot, 8 cars

```{r}

car.abund.order <- c('CD28','4-1BB','BAFF-R','TACI','CD40','CD30','KLRG1')

# add relative baseline measurement if not added
#   (preserve idempotence for debug)

if (!('car.abund.rel.baseline' %in% names(read.counts))) {
  car.abund.set <- read.counts[assay=='baseline' & batch=='prolif2',
            list(car.abund.baseline= car.abund, donor, CAR.align, t.type)]
  
  read.counts <- read.counts[car.abund.set, on=.(donor, CAR.align, t.type)]
  
  read.counts[, car.abund.rel.baseline := car.abund/car.abund.baseline,
              by=.(donor, t.type, batch)]
  
  read.counts[assay == 'baseline',
              car.abund.rel.baseline := 1,
              by=.(batch, donor, t.type, CAR.align)]
}

car.abund.data <- rbind(
  read.counts[k.type == 'pos' & batch != 'post-cytof'], 
  copy(read.counts[batch != 'post-cytof' & assay == 'baseline'])[,
    k.type := 'pos'])[k.type != 'NA'][CAR.align %in% car.abund.order]

car.abund.data[CAR.align == 'TNR8', CAR.align := 'CD30']

car.abund.data[, CAR.align.ord := factor(CAR.align, levels=car.abund.order)]

# add significance
car.abund.data <- (pooled_prolif_p_values[t.type == 'both'][, -'t.type'] %>% 
    add_significance(p.col='padj'))[
      car.abund.data, on='CAR.align']

# add **s as a parsed label

car.abund.order.pval_labels <- unique(
  car.abund.data[order(CAR.align.ord), list(CAR.align.ord, padj.signif)])[,
    paste0('"',car.abund.order,'"^"', padj.signif, '"')]

car.abund.data[,
  CAR.align.ord.pval := factor(CAR.align.ord,
    levels=car.abund.order,
    labels=car.abund.order.pval_labels)]

car.abund.plot <- ggplot(car.abund.data,
    aes(
      x=day, y=log2(car.abund.rel.baseline),
      color= interaction(t.type), group= interaction(k.type, t.type, batch, donor))) + 
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks=c(0,12,24), limits=c(0,24.5)) +
  scale_color_manual('', values=brewer.pal(9, 'YlOrRd')[c(4,6)]) +
  facet_wrap(~CAR.align.ord.pval, ncol=8,
    labeller="label_parsed") + 
  scale_linetype_manual(values=c(2,1)) +
  labs(x='Timepoint (days)', y='Relative Expansion, log2 FC') + 
  geom_hline(linetype=2, yintercept=0) + 
  theme_minimal() +
  theme(
    legend.position='bottom',
    panel.grid = element_blank(),
    panel.border = element_rect(fill=NA),
    strip.background=element_rect(linetype=1, color='white', size=3),
    strip.text = element_text(color='white',face='bold'))

color_car_strips <- function(plot) {

  car.abund.gtable <- ggplot_gtable(ggplot_build(plot))
  strip_both <- which(grepl('strip-', car.abund.gtable$layout$name))
  fills <- car_colors
  k <- 1
  
  for (i in strip_both) {
    j <- which(grepl('rect', car.abund.gtable$grobs[[i]]$grobs[[1]]$childrenOrder))
    car.abund.gtable$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }
  return(car.abund.gtable)
}
```

### Supp Fig3 - CD19- abund

```{r}

abund.array.list <- c('BAFF-R','CD30','4-1BB','TACI','CD28','KLRG1','CD40','TNR8')

car.abund.data.neg <- rbind(
  read.counts[k.type == 'neg' & batch != 'post-cytof'], 
  copy(read.counts[batch != 'post-cytof' & assay == 'baseline'])[,
    k.type := 'neg'])[k.type != 'NA'][CAR.align %in% abund.array.list]

car.abund.data.neg[CAR.align == 'TNR8', CAR.align := 'CD30']

car.abund.data.neg[, CAR.align.ord := factor(CAR.align, levels=car.abund.order)]

# add significance
car.abund.data.neg <- (pooled_prolif_p_values[t.type == 'both'][, -'t.type'] %>% 
    add_significance(p.col='padj'))[
      car.abund.data.neg, on='CAR.align']

# add **s as a parsed label

car.abund.order.pval_labels <- unique(
  car.abund.data.neg[order(CAR.align.ord), list(CAR.align.ord, padj.signif)])[,
    paste0('"',car.abund.order,'"^"', padj.signif, '"')]

car.abund.data.neg[,
  CAR.align.ord.pval := factor(CAR.align.ord,
    levels=car.abund.order)]

car.abund.data.neg.plot <- ggplot(car.abund.data.neg,
    aes(
      x=day, y=log2(car.abund.rel.baseline),
      color= interaction(t.type), group= interaction(k.type, t.type, batch, donor))) + 
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks=c(0,12,24), limits=c(0,24.5)) +
  scale_color_manual('', values=brewer.pal(9, 'YlOrRd')[c(4,6)]) +
  facet_wrap(~CAR.align.ord.pval, ncol=8) + 
  scale_linetype_manual(values=c(2,1)) +
  labs(x='Timepoint (days)', y='Relative Expansion, log2 FC') + 
  geom_hline(linetype=2, yintercept=0) + 
  theme_minimal() +
  theme(
    legend.position='bottom',
    panel.grid = element_blank(),
    panel.border = element_rect(fill=NA),
    strip.background=element_rect(linetype=1, color='white', size=3),
    strip.text = element_text(color='white',face='bold'))

```

### Fig3: PCA
```{r}

read.counts[CAR.align == 'TNR8', CAR.align := 'CD30']
read.counts[CAR.align == 'HAVCR1b', CAR.align := 'TIM1']

pca_shared_cols <- c(intersect(
  intersect(
    intersect(names(rel_abund_means),names(cd69.ranked)),
    intersect(names(cyto.ranked.all),names(car_score_means))),
  names(deseq2_combined)))

pc_total_combined <- unique(rbind(
  #deseq2 measurements for CTV1
  deseq2_combined[, pca_shared_cols, with=F],
  
  #new and old cytokine combined with z-scores and p-values
  #unique(copy(cyto.ranked.all)[,  c('sort.group', pca_shared_cols), with=F]),
  unique(copy(cyto.ranked.all)[, padj := p.adj.ta][, value.scale := value.ta][, 
    shared_cols, with=F]),
  
  #cd69 from both donors
  copy(cd69.ranked)[comparison == 'rlog_car_score',
    c('sort.group', pca_shared_cols), with=F],
  
  #car scores for CTV
  copy(car_score_means)[, comparison := 'rlog_car_score'][, 
    c('sort.group', pca_shared_cols), with=F],
  
  #relative abundance/expansion
  copy(rel_abund_means)[comparison == 'abund_rel_baseline', 
    c('sort.group',pca_shared_cols), with=F],

  fill=T))

cast_metrics <- dcast(
  pc_total_combined[!is.na(value.scale) & 
    !(assay == 'CTV3' & !grepl('baseline', comparison))],
  CAR.align ~ assay + k.type + t.type + comparison + sort.group, 
  value.var='value.scale')

cast_metrics[is.na(cast_metrics)] <- 0
#prcomp_metrics <- data.frame(cast_metrics[, -1])
#prcomp_metrics[is.na(prcomp_metrics)] <- 0
combined_pca <- prcomp(cast_metrics[, -1])

# calculate pca stats
pca.dt <- data.table(pc = data.table(colnames(combined_pca$rotation))[, 
    PC := as.integer(gsub("[A-Z]", "", V1))][, PC], 
    sd = combined_pca$sdev, 
    var = combined_pca$sdev^2, 
    var.norm = combined_pca$sdev^2/sum(combined_pca$sdev^2), 
    var.acc = cumsum(combined_pca$sdev^2/sum(combined_pca$sdev^2)))

melt.pca.dt <- melt(
  pca.dt, measure.vars = c("var.norm", "var.acc"), variable.name = "metric")

pca.stats.plot <- ggplot(melt.pca.dt) +
  geom_line(aes(x = pc, y = value, color = metric)) +
  geom_point(aes(x = pc, y = value, color = metric)) +
  scale_x_continuous(limits = c(1, NA)) +
  labs(title = "Fraction of Variance Captured by Principal Components, Every Sort Group", 
       x = "Principal Component", y = "Fraction of Variance") +
  theme_bw()

# project data onto principal components
projected.metrics <- scale(cast_metrics[, setdiff(names(cast_metrics), 
  c("CAR.align")), with = FALSE],
  combined_pca$center, combined_pca$scale) %*% combined_pca$rotation

projected.metrics <- cbind.data.frame(CAR.align=cast_metrics[, CAR.align],
  projected.metrics)

projected.metrics <- data.table(merge(projected.metrics, 
  unique(read.counts[, .(CAR.align, len)]), 
  by = "CAR.align"))

projected.metrics[, CAR.annot := 'none']
projected.metrics[CAR.align %in% names(car_colors), CAR.annot := CAR.align]
projected.metrics[, CAR.annot := factor(
  CAR.annot, levels=c(names(car_colors)[1:7], 'none'))]

pct_exp <- scales::label_percent()(pca.dt[pc %in% c(1,2), var.norm])

# by name
pca.pos.sort.group.plot <- ggplot(projected.metrics,
    aes(x = -PC1, y = -PC2, label = CAR.annot,
      size=(CAR.align %in% names(car_colors)),
      color=CAR.annot)) +
  geom_point() +
  geom_text_repel(data=copy(projected.metrics)[
      CAR.annot == 'none', CAR.annot := ' '],
    size=3, color='grey20', box.padding = 0.5, segment.alpha=0) +
  scale_size_discrete(guide=F) +
  scale_color_manual('',
    labels=c(names(car_colors)[1:7], 'other costims'),
    values=c(car_colors[1:7], 'grey')) +
  labs(
    x=paste0('PC1 (',pct_exp[1],')'),
    y=paste0('PC2 (',pct_exp[2],')')) +
  theme_bw() +
  theme(axis.text = element_blank(), panel.grid=element_blank())

# by size
pca.pos.sort.size.plot <- ggplot(
    projected.metrics[,
      CAR.label := ifelse(CAR.annot == 'none', '', as.character(CAR.align))],
    aes(
      x = -PC1, y = -PC2, label = CAR.label, 
      size=(CAR.align %in% names(car_colors)), 
      fill=len/3)) +
  geom_point(pch=21, color='black') +
  scale_size_manual(guide=F, values=c(2,6)) +
  scale_fill_distiller(palette='RdYlBu') +
  geom_text_repel(data=projected.metrics,
    size=3, color='grey20', box.padding = 0.5, segment.alpha=0) +
  labs(
    x=paste0('PC1 (',pct_exp[1],')'),
    y=paste0('PC2 (',pct_exp[2],')')) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    panel.grid=element_blank(),
    legend.key.width=unit(0.4, 'lines'))

# by type

types <- fread(text="
  CAR.align,Activity,Family
  CRTAM,Costimulatory,IgSF
  CD28,Costimulatory,IgSF
  ICOS,Costimulatory,IgSF
  CXADR,Costimulatory,IgSF
  CD244,Costimulatory,IgSF
  CD7,Costimulatory,IgSF
  NTB-A,Costimulatory,IgSF
  CD200R,Costimulatory,IgSF
  NKR-P1A,Costimulatory,IgSF
  GITR,Costimulatory,TNFRSF
  CD30,Costimulatory,TNFRSF
  4-1BB,Costimulatory,TNFRSF
  CD40,Costimulatory,TNFRSF
  CD96,Unknown,IgSF
  CD300a,Unknown,CD300
  CD300f,Unknown,CD300
  CD2,Mixed,IgSF
  TIM1,Mixed,TIM
  NKG2D,Mixed,Lectin
  CD72,Mixed,Lectin
  DC-SIGN,Mixed,Lectin
  Siglec-3,Mixed,Lectin
  TACI,Mixed,TNFRSF
  BAFF-R,Mixed,TNFRSF
  BCMA,Mixed,TNFRSF
  LAG3,Inhibitory,IgSF
  CRACC,Inhibitory,IgSF
  TIGIT,Inhibitory,IgSF
  BTLA,Inhibitory,IgSF
  LAIR1,Inhibitory,IgSF
  PD1,Inhibitory,IgSF
  CTLA4,Inhibitory,IgSF
  ILT2,Inhibitory,IgSF
  ILT3,Inhibitory,IgSF
  ILT4,Inhibitory,IgSF
  KLRG1,Inhibitory,IgSF
  KIR3DL1,Inhibitory,IgSF
  KIR2DL1,Inhibitory,IgSF
  HAVCR2,Inhibitory,TIM
  TLT-1,Inhibitory,TREM")

pca_type_combined <- merge(types, projected.metrics, by='CAR.align', all=T)
pca_type_combined[,
  CAR.label := ifelse(CAR.annot == 'none', ' ', as.character(CAR.align))]

pca.pos.sort.type.plot <- ggplot(pca_type_combined,
    aes(x = -PC1, y = -PC2, label = CAR.label, shape=Activity, fill=Family,
      size=(CAR.align %in% names(car_colors)))) +
  geom_point() +
  scale_shape_manual(values=c(24,25,23,21)) +
  scale_size_manual(guide=F, values=c(2,5)) +
  guides(
    fill = guide_legend(override.aes = list(shape = 21, size=5)),
    shape = guide_legend(override.aes = list(size=5))) +
  scale_fill_manual(values=c(
    setNames(
      brewer.pal(12,'Paired')[c(5,3,7,9,10)],
      c('CD300','IgSF','Lectin','TIM','TREM')),
    setNames(brewer.pal(5,'Blues')[3],'TNFRSF')
    )) +
  geom_text_repel(data=pca_type_combined,
    size=3, color='grey20', box.padding = 0.4, segment.alpha=0) +
  labs(
    x=paste0('PC1 (',pct_exp[1],')'),
    y=paste0('PC2 (',pct_exp[2],')')) +
  theme_bw() +
  theme(axis.text = element_blank(), panel.grid=element_blank())

#explain components
# map component names back to variables
comp_names <- unique(cbind(
    pc_total_combined[,paste(assay, k.type, t.type, comparison, sort.group, sep='_')],
    pc_total_combined[,paste(assay, k.type, t.type, comparison, sort.group, sep='|')]))
indiv_components <- data.table(matrix(unlist(strsplit(comp_names[,2], '\\|')), ncol=5, byrow=T))
names(indiv_components) <- c('assay', 'k.type', 't.type', 'comparison', 'sort.group')
indiv_components <- data.table(measure=comp_names[,1], indiv_components)

component_rotations <- indiv_components[
  melt.data.table(data.table(combined_pca$rotation, keep.rownames=T), id.vars='rn'),
  on=c(measure='rn')]

indiv_component_plot <- ggplot(
  component_rotations[
    variable %in% c('PC1','PC2')][,
      list(diff_1v2= value[variable == 'PC1'] - value[variable == 'PC2']),
      by=.(assay, comparison, k.type, t.type)][, assay_group := assay][
        assay %in% c('CD69','IFNy','IL2'), assay_group := 'markers'][,
          list(value= sum(diff_1v2)), by=.(t.type, k.type, assay, assay_group)]) + 
  geom_bar(aes(x=assay, y=value, fill=assay_group), color='grey50', stat='identity') +
  facet_grid(k.type + assay_group ~ t.type, scales='free', space='free') +
  geom_hline(yintercept=0) +
  coord_flip() +
  labs(x='Pooled Assay Scores, grouped', y='<- More PC2     More PC1 ->')

plot_rotations <- component_rotations[variable %in% c('PC1','PC2')][,
  `:=`(min_val= min(value), max_val=max(value)), by=.(k.type, assay, t.type, comparison)][,
    measure_group := assay][
      grep('baseline', comparison), measure_group := 'expansion'][
        assay %in% c('CD69','IFNy','IL2'), measure_group := 'markers']

plot_summary <- copy(plot_rotations)[
  measure_group == 'markers', measure_group := assay][
  measure_group == 'expansion', measure_group := paste(assay, measure_group, sep='.')][,
  list(val_mean=mean(-value)), by=.(measure_group, k.type, t.type, variable)]

plot_summary[, measure_group := factor(
  measure_group,
  levels=c('IL2','IFNy','CD69','CTV3.expansion',
           'CTV2.expansion','CTV1.expansion','CTV2','CTV1'),
  labels=c('IL2','IFNy','CD69','Expansion:\nd0-d24','Expansion:\nd0-d16',
           'Expansion:\nd0-d4','Proliferation:\nd9-d16','Proliferation:\nd0-d3/4'))]

pca_factors <- ggplot(plot_summary[, k.type := factor(k.type, labels=c('CD19-','CD19+'))]) + 
  geom_bar(
    aes(x=measure_group, y=val_mean, fill=variable),
    stat='identity', position=position_dodge(0.7), color='grey30', size=0.5, alpha=0.8, width=0.7) +
  facet_grid(k.type~t.type, scales='free_y', space='free_y') +
  geom_hline(yintercept=0, size=0.5, linetype=2) +
  scale_y_continuous(breaks=c(-0.10, 0, 0.10)) +
  coord_flip() +
  labs(x='Measurement (mean DESeq2 LFC)',y="Principal Component\nContribution (rotation)") +
  theme_minimal() +
  theme(
    panel.border=element_rect(color='black', fill=NA),
    panel.grid=element_blank())

ratio.values <- pca_type_combined[, (max(PC1)-min(PC1))/(max(PC2)-min(PC2))]
```

### Fig2: cytokine panels - by cytokine, colored with donor - using this

```{r}
cyto.plot.data <- cyto.read.counts.all[k.type == 'pos' & sort.cond != 'neg' &
  assay %in% c('il2.ifng', 'il2','ifng') &
  sort.cond %in% c('il2','ifng') &
  car.sort.group.cells > 500]
cyto.plot.data[CAR.align == 'BTLA.a', CAR.align := 'BTLA']
cyto.plot.data[, n_measure := .N, by=.(CAR.align, sort.cond)]
cyto.plot.data <- cyto.plot.data[n_measure > 1]

cyto.plot.data[, CAR.axis := factor(paste(CAR.align,sort.cond, sep='|'))]
# cyto.plot.data[, CAR.axis := factor(
#   CAR.axis, cytokine_panel
#   levels=levels(CAR.axis),
#   labels=gsub('(.*)\\|.*', '\\1', levels(CAR.axis)))]

#fold-change from mean
cyto.plot.data[, l2fc.mean := log2(car.bin.pct.incl/mean(car.bin.pct.incl, na.rm=T)),
  by=.(sort.cond, donor, t.type)]

lfc.max = 3

cyto.plot.data[, l2fc.mean.donor := mean(l2fc.mean, na.rm=T),
  by=.(CAR.align, sort.cond, donor)]

cyto.plot.data[, l2fc.mean.subset := mean(l2fc.mean, na.rm=T),
  by=.(CAR.align, sort.cond, t.type)]


cd4_cd8_colors <- brewer.pal(9, 'YlOrRd')[c(4,6)]

donor_colors = c(
  brewer.pal(11, 'RdBu')[9],
  brewer.pal(11, 'PuOr')[9],
  brewer.pal(11, 'PRGn')[9])

cyto.plot.data[, donor := factor(donor, labels=paste0('Donor ', c('A','B','C')))]

# deal with OOB and CD4/CD8 together
cyto.plot.data[, pt.shape := factor(
  sign(l2fc.mean)*(abs(l2fc.mean) > 3),
  levels=c(-1,0,1),
  labels=c('OOB.lo','IB','OOB.hi'))]
cyto.plot.data[, l2fc.mean.disp := ifelse(
  abs(l2fc.mean) < 3,l2fc.mean, lfc.max*.98*sign(l2fc.mean))]
cyto.plot.data[pt.shape == 'IB', pt.shape := t.type]

pt.shape.types=c('OOB.lo'=25, 'CD4'=21, 'CD8'=22, 'OOB.hi'=24)
pt.shape.types_samesubset=c('OOB.lo'=25, 'CD4'=21, 'CD8'=21, 'OOB.hi'=24)

cyto.plot.data[,
  l2fc.mean.sort.cond := mean(l2fc.mean, na.rm=T),
  by=.(CAR.align, sort.cond)]

cyto.plot.data[,
  l2fc.sem.sort.cond := sd(l2fc.mean, na.rm=T)/sqrt(.N),
  by=.(CAR.align, sort.cond)]

cyto.plot.data[,
  l2fc.mean.subset := mean(l2fc.mean, na.rm=T),
  by=.(CAR.align, sort.cond)]


  do_wilcox_test <- function(df) {
  wilcox_test(df,
    sort.group.cond.rank ~ CAR.align,
    ref.group='all',
    alternative='g') %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
}

car_ranks <- cyto.plot.data[,
      `:=`(
        sort.group.cond.rank= rank(-car.bin.pct.incl),
        CAR.align.sort.cond= factor(CAR.align)),
      by=.(sort.cond, sort.group)][,
        n_measure := .N, by=.(CAR.align)][n_measure > 1][,
            CAR.align := factor(as.character(CAR.align))]

wilcoxon_stats <- car_ranks[, do_wilcox_test(.SD), by='sort.cond']
wilcoxon_stats <- wilcoxon_stats[, list(
  CAR.align=group2, p.adj, p.adj.signif, sort.cond,
  CAR.axis= factor(paste(group2,sort.cond, sep='|')))]

remove_t_type_sep_col_bold_signif <- function (breaks) {
  car_name <- unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))
  
  car_axis_colors <- rep('black', length(car_name))
  car_axis_colors[grep('CD28', car_name)] <- car_colors['CD28']
  car_axis_colors[grep('4-1BB', car_name)] <- car_colors['4-1BB']

  # bold CAR names if colored or if signif in wilcoxon_stats
  bold_cars <- breaks %in% wilcoxon_stats[p.adj.signif != 'ns', CAR.axis]
  car_name[bold_cars] <- 
      paste0('**',car_name[bold_cars],'**')
  
  car_color_span <- paste("<span style = 'color: ",
    car_axis_colors,
    ";'>",
    car_name,
    "</span>", sep = "")
  
  return(car_color_span)
}

cytokine_panel <- ggplot(cyto.plot.data, 
    aes(
      x = reorder(CAR.axis, l2fc.mean), 
      y = l2fc.mean.disp, fill=donor, color=donor,
      shape = pt.shape)) +
  geom_point(size=1.5) +
  geom_point(size=2.5, shape=21, color='grey30', fill=NA,
    aes(y=l2fc.mean.sort.cond)) +
  scale_shape_manual(values=pt.shape.types, breaks=c('CD4','CD8')) +
  scale_fill_manual(values=donor_colors, guide='none') +
  scale_color_manual(values=donor_colors) +
  facet_wrap(factor(sort.cond, labels=c("'IFN'~gamma",'IL2')) ~ ., scales='free', labeller=label_parsed) +
  geom_hline(yintercept=0, linetype=2, color='grey60') +
  scale_y_continuous(limits=c(-lfc.max,lfc.max), expand=c(0,0)) +
  scale_x_discrete(labels= remove_t_type_sep_col_bold_signif) +
  labs(y = 'Relative Cytokine Secretion,\nlog2 FC', 
       color='', shape='') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA),
        axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.box='horizontal', 
        strip.background = element_rect(colour="white", fill="white"),
        legend.justification=c(1,1),
        legend.position=c(0.25,1))


cytokine_panel_errorbars <- ggplot(cyto.plot.data, 
    aes(
      x = reorder(CAR.axis, l2fc.mean), 
      y = l2fc.mean.sort.cond, fill=donor, color=donor,
      ymin= l2fc.mean.sort.cond-l2fc.sem.sort.cond,
      ymax= l2fc.mean.sort.cond+l2fc.sem.sort.cond,
      shape = pt.shape)) +
  geom_point(size=2.5, shape=21, color='grey30', fill=NA,
    aes(y=l2fc.mean.sort.cond)) +
  geom_linerange(color='grey30', fill=NA) +
  scale_shape_manual(values=pt.shape.types, breaks=c('CD4','CD8')) +
  scale_fill_manual(values=donor_colors, guide='none') +
  scale_color_manual(values=donor_colors) +
  facet_wrap(factor(sort.cond, labels=c("'IFN'~gamma",'IL2')) ~ ., scales='free', labeller=label_parsed) +
  geom_hline(yintercept=0, linetype=2, color='grey60') +
  scale_y_continuous(limits=c(-lfc.max,lfc.max), expand=c(0,0), oob = squish) +
  scale_x_discrete(labels= remove_t_type_sep_col_bold_signif) +
  labs(y = 'Relative Cytokine Secretion,\nlog2 FC', 
       color='', shape='') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA),
        axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.box='horizontal', 
        strip.background = element_rect(colour="white", fill="white"),
        legend.justification=c(1,1),
        legend.position=c(0.25,1))

cytokine_subset_colors <- ggplot(cyto.plot.data, 
    aes(
      x = reorder(CAR.axis, l2fc.mean), 
      y = l2fc.mean.subset, color=t.type)) +
  geom_point(size=1.5) +
  geom_point(size=2.5, shape=21, color='grey30', fill=NA,
    aes(y=l2fc.mean.sort.cond)) +
  scale_color_manual(values=brewer.pal(9, 'YlOrRd')[c(4,6)], breaks=c('CD4','CD8')) +
  facet_wrap(factor(sort.cond, labels=c("'IFN'~gamma",'IL2')) ~ ., scales='free', labeller=label_parsed) +
  geom_hline(yintercept=0, linetype=2, color='grey60') +
  scale_y_continuous(limits=c(-3.5,3.5), expand=c(0,0)) +
  scale_x_discrete(labels= remove_t_type_sep_col_bold_signif) +
  labs(y = 'Relative Cytokine Secretion,\nlog2 FC', 
       color='', shape='') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA),
        axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.box='horizontal', 
        strip.background = element_rect(colour="white", fill="white"),
        legend.justification=c(1,1),
        legend.position=c(0.25,1))
```

### Fig2: early vs late plot

```{r}
read.counts[, car.abund.log := log10(car.abund)]

#assay_rep_list:
assay_rep_set <- unique(read.counts[
  data.table(
    assay=c('baseline','CTV1','CTV2','CTV3'), 
    prev_assay=c(NA,'baseline','CTV1','CTV2')), on='assay'][, 
  list(assay, sort.group, prev_assay, donor, batch, t.type, k.type)])

#map day0 
assay_rep_set <- rbind(
  assay_rep_set[assay=='baseline'][, k.type := 'pos'], 
  assay_rep_set[assay=='baseline'][, k.type := 'neg'], 
  assay_rep_set[assay!='baseline'])

#merge with prev copy to get prev sort.group correspondence
assay_rep_set <- assay_rep_set[, list(
    donor, batch, t.type, k.type,
    prev_assay=assay,prev.sort.group=sort.group)][
  assay_rep_set,
  on=c('donor','batch','t.type','k.type','prev_assay')]

# use baseline from prolif2 always
assay_rep_set[assay == 'CTV1', 
  prev.sort.group := gsub('prolif1','prolif2',prev.sort.group)]

#merge with prev measurements
prev_measure <- unique(read.counts[, list(
  car.abund.prev=car.abund, prev.sort.group=sort.group, CAR.align)])[
    assay_rep_set, on=c('prev.sort.group')]

#merge with orig read counts
sprinter.read.counts <- prev_measure[, list(
    car.abund.prev, prev.sort.group, CAR.align, sort.group)][
      read.counts, on=c('sort.group','CAR.align')]

#calculate relative prev
sprinter.read.counts[, car.abund.rel.prev := car.abund/car.abund.prev]

# expansion vs proliferation
car_label_set <- c('BAFF-R','CD30','4-1BB','TACI',
                   'CD28','KLRG1','CD40','NTB-A','CD2')

plot_set <- unique(
  sprinter.read.counts[assay %in% c('CTV1', 'CTV2')][, 
    car.lbl := ifelse(
      CAR.align %in% car_label_set, as.character(CAR.align), '')][,
    list(car.lbl, rlog_car_score, car.abund.rel.baseline, 
         car.abund.rel.prev, k.type, assay, t.type, donor, 
         CAR.align, batch, sort.group, avg.divisions)])

early_v_late_indiv_dt <- dcast(
    plot_set, car.lbl+k.type+t.type + donor + CAR.align + batch ~ assay, 
    value.var=c('rlog_car_score','car.abund.rel.baseline','car.abund.rel.prev'))[
      k.type=='pos'][, late_v_early := log2(
        car.abund.rel.prev_CTV1/car.abund.rel.prev_CTV2)]

# early vs late stats
early_v_late_lm_dt <- early_v_late_indiv_dt[,
  data.table(
    summary(lm(late_v_early ~ 0+CAR.align))$coeff, keep.rownames=T)]
setnames(early_v_late_lm_dt, c('rn','Pr(>|t|)'),c('CAR.align','p.val'))
early_v_late_lm_dt[, CAR.align := gsub('CAR.align','',CAR.align)]

# total exp stats
total_exp_lm_dt <- early_v_late_indiv_dt[,
  data.table(
    summary(lm(log2(car.abund.rel.baseline_CTV2) ~ 0+CAR.align))$coeff, keep.rownames=T)]
setnames(total_exp_lm_dt,c('rn','Pr(>|t|)'),c('CAR.align','p.val'))
total_exp_lm_dt[, CAR.align := gsub('CAR.align','',CAR.align)]

sprinter_dt <- merge(
  total_exp_lm_dt,
  early_v_late_lm_dt,
  by=c('CAR.align'),
  suffixes=c(".total",".evl"))

# set most labels to blank
sprinter_dt <- sprinter_dt[,
    car.lbl := ifelse((p.val.total < 0.05 & Estimate.total > 0) | 
    (CAR.align %in% c('4-1BB','CD28','BAFF-R','TACI')), CAR.align, '')]

# add additional blank points to make ggrepel avoid annotation boxes
sprinter_lbls_w_blanks <- rbind(
  sprinter_dt,
  data.table(
    car.lbl='',
    Estimate.evl =rep(c(-1.3,1.3),2),
    Estimate.total =rep(c(-.85,.85), each=2)),
  fill=T)

sprinter_plot <- ggplot(sprinter_dt,
  aes(y=-Estimate.evl, x=Estimate.total, label=car.lbl)) + 
  geom_point(pch=21, aes(
    color=(p.val.evl > 0.05 | car.lbl != ''),
    fill=factor(as.numeric((p.val.evl < 0.05))*sign(Estimate.evl),
      labels=c('later', 'n.s.', 'earlier')),
    size=ifelse(Estimate.total > 0, -log10(p.val.total),0))) +
  scale_color_manual(values=c('white','black'), guide='none') +
  scale_fill_manual(
    values=c(later=hcl.colors(8,'PRGn')[7],ns='white',earlier=hcl.colors(8,'PRGn')[2]),
    na.value='white',
    limits=c('earlier','later')) +
  geom_text_repel(data=sprinter_lbls_w_blanks, seed=10, point.padding=0.1) +
  annotate(geom='label', x=-Inf, y=Inf, label='later/poor', 
    label.r=unit(0,'pt'), size=3, vjust=.99, hjust=.01, color=hcl.colors(8,'PRGn')[7]) +
  annotate(geom='label', x=-Inf, y=-Inf, label='earlier/poor', 
    label.r=unit(0,'pt'), size=3, vjust=.01, hjust=.01, color=hcl.colors(8,'PRGn')[2]) +
  annotate(geom='label', x=Inf, y=Inf, label='later/potent', 
    label.r=unit(0,'pt'), size=3, vjust=.99, hjust=.99, color=hcl.colors(8,'PRGn')[7]) +
  annotate(geom='label', x=Inf, y=-Inf, label='earlier/potent', 
    label.r=unit(0,'pt'), size=3, vjust=.01, hjust=.99, color=hcl.colors(8,'PRGn')[2]) +
  geom_vline(xintercept=0, linetype=2, color='grey60') +
  geom_hline(yintercept=0, linetype=2, color='grey60') +
  scale_size_continuous(range=c(2,6)) +
  #scale_x_continuous(expand=c(0.2,0.1)) +
  scale_y_continuous(expand=c(0.08,0.04)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    legend.position= 'bottom',
    legend.box = "vertical",
    panel.border = element_rect(colour = "black")) +
  labs(
    y='Late vs Early Expansion,\nlog2(d3-14 / d0-3)',
    x='Overall expansion, log2(d0-d14) FC from mean',
    fill='Significant late/early expansion,\np<0.05:',
    size='Overall rel. expansion, -log10(p):')
```

## Fig2: abudance over time plot

```{r}
prolif_data <- read.counts[
  (k.type =='pos' | is.na(k.type)) & batch != 'post-cytof'][,
    list(batch, donor, 
      mean.car.abund.rel.baseline= mean(log2(car.abund.rel.baseline)), 
      sd.car.abund.rel.baseline= 2^sd(log2(car.abund.rel.baseline))), 
    by=c('CAR.align', 't.type','assay')]

# # add in mixed linear model p-values
prolif_data <- unique(prolif_data)[unique(pooled_prolif_p_values
  [t.type != 'both']), on=c('CAR.align', 't.type')]
prolif_data <- add_significance(prolif_data, p.col='padj')

#reorder CAR names to put CD28 and 41BB in front (at end of levels list)
new_car_levels  <- prolif_data[, c(
  levels(CAR.align)[!(levels(CAR.align) %in% c('4-1BB','CD28'))], 
  '4-1BB','CD28')]

prolif_data[, CAR.align := factor(CAR.align, new_car_levels)]

prolif_data[, CAR.type := 'other']
prolif_data[CAR.align == '4-1BB', CAR.type := '4-1BB']
prolif_data[CAR.align == 'CD28', CAR.type := 'CD28']
prolif_data[, CAR.type := factor(
  CAR.type,levels=c('other','4-1BB','CD28'))]


top_num <- 6

label_data <- unique(prolif_data[assay=='CTV3', 
  list(t.type, CAR.align, mean.car.abund.rel.baseline,
       assay, CAR.type, padj, padj.signif)])[, 
    car_order := rank(mean.car.abund.rel.baseline), by='t.type'][
      car_order %in% c(1:(1+top_num),(40-top_num):40) | 
        CAR.align %in% c('CD28','4-1BB')]

#color labels pink vs green
label_data[, CAR.topbot := CAR.type]
label_data[car_order %in% 1:(1+top_num) & CAR.type == 'other', 
  CAR.topbot :='top']
label_data[car_order %in% (40-top_num):40 & CAR.type == 'other',
  CAR.topbot :='bot']
label_data[, CAR.topbot := factor(CAR.topbot)]
label_data <- label_data[padj.signif != 'ns' | CAR.align %in% c('CD28','4-1BB')]

# add topbot data from labels to prolif data to color lines
abund_plot_data <- label_data[,
  list(CAR.align,t.type,CAR.topbot)][
    prolif_data[order(CAR.type)], on=c('CAR.align','t.type')]
abund_plot_data[is.na(CAR.topbot), CAR.topbot := 'other']


fig2_abundance_over_time_vert <- ggplot(
  abund_plot_data[rev(order(CAR.topbot))], aes(
    y=mean.car.abund.rel.baseline, x=assay,
    color=CAR.topbot,
    group=CAR.align,
    label=CAR.align)) + 
  # line
  geom_line(aes(size=CAR.topbot)) +
  scale_size_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
    values=c(1.5,1.5,0.5,0.8,0.5), guide=F) +
  scale_color_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
    values=c(receptor_cols, outlier_cols[1], "#40B549", 'grey40'),
    guide=F) +
  #points
  new_scale('size') +
  geom_point(aes(size=CAR.topbot, color=CAR.topbot, group=CAR.align)) +
  scale_size_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom','Other'),
    values=c(2,2,1,1.5,1), guide=F) +
  # text
  new_scale('color') +
  geom_text_repel(data=label_data,
    aes(color=CAR.topbot, label=paste0('"',CAR.align,'"^"', padj.signif, '"')),
    parse=T,
    nudge_x = 1.2,
    direction = "y",
    hjust = 1,
    size = 2.7,
    segment.size = 0.3,
    segment.color='grey') +
  scale_color_manual('',
    labels=c('4-1BB', 'CD28','Top','Bottom'),
    values=c(receptor_cols, outlier_cols),
    guide=F) +
  facet_grid(t.type ~ .) +
  scale_x_discrete(
    'Days of continuous CD19 stimulation', 
    expand = c(0, 0.4, 0, 1.4), 
    labels=c('Baseline', 'd3-4', 'd14-16', 'd24')) +
  scale_y_continuous('Relative Expansion,\n log2 FC from mean ') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA), panel.grid=element_blank())
```

### Fig2: CD4 vs CD8s (new version, w/ mean prolif and diff)

```{r, fig.width=7, fig.height=4}

cast_4v8 <- dcast(data.dt[
    k.type == 'pos' & 
        ((ref_set == 'baseline' & test_set == 'A' & assay != 'CTV1') | 
        (ref_set == 'CD' & test_set == 'A' & assay == 'CTV1'))], 
      CAR.align + assay + CAR.type ~ t.type, 
    value.var = c("log2FoldChange", "padj.disp"))

# rename facets
cast_4v8[, assay := factor(assay, 
    labels=assay_fig_labels)]

# label significant hits
cast_4v8[, CAR.type.size := CAR.type]
cast_4v8[
    ((padj.disp_CD4 > -log10(0.05) & log2FoldChange_CD4 > 0) | 
    (padj.disp_CD8 > -log10(0.05) & log2FoldChange_CD8 > 0)), 
  CAR.type.size := 'other_sig']

cast_4v8[, CAR.label.sig := '']
cast_4v8[CAR.type.size != 'other', CAR.label.sig := CAR.align]

cast_4v8[, mean_lfc := rowMeans(cbind(log2FoldChange_CD4, log2FoldChange_CD8))]
cast_4v8[, rel_8 := log2FoldChange_CD8 - log2FoldChange_CD4]

cast_4v8[CAR.type.size == 'other_sig' & xor(padj.disp_CD4 > -log10(0.05), 
  padj.disp_CD8 > -log10(0.05)), CAR.type.size := 'other_sig_diff']

# add false points to get ggrepel to avoid plotting underneath the skewed lbls
cast_4v8_w_blanks <- rbind(
  cast_4v8,
  data.table(CAR.label.sig='', rel_8 =rep(c(-1,1),3), mean_lfc=rep(-1,6),
   assay = rep(levels(cast_4v8$assay),each= 2)),
  fill=T)

prolif_4v8_rel_vert <- ggplot(cast_4v8_w_blanks[order(CAR.type)], 
  aes(y=rel_8, x=mean_lfc,
      color=CAR.type,
      label=CAR.label.sig,
      size=CAR.type.size)) + 
  annotate(geom='label', x=-Inf, y=Inf, label='CD8 Skewed', 
    label.r=unit(0,'pt'), size=3, vjust=.99, hjust=.01) +
  annotate(geom='label', x=-Inf, y=-Inf, label='CD4 Skewed', 
    label.r=unit(0,'pt'), size=3, vjust=.01, hjust=.01) +
  geom_point() +
  facet_grid(assay ~ ., scales='free_x') +
  scale_color_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28'),
    values=c('grey20', receptor_cols), guide=F) +
  scale_size_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28', 'New Receptors', 'New Receptors'),
    values=c(0.8,0.8,0.8,2.5,2.5), guide=F) +
  scale_x_continuous(limits=c(-2.5,3.25), breaks=c(-2:3)) +
  scale_y_continuous(limits=c(-2.25,1.9)) +
  geom_hline(yintercept=0, linetype='dashed', alpha=.7) +
  geom_text_repel(size=3, point.padding=0.3, segment.color='grey', max.overlaps=20) +
  theme_minimal() + 
  labs(y='log2 CD8:CD4 proliferation ratio', 
       x='Rel. proliferation/expansion, log2 FC from mean') +
  theme(panel.border=element_rect(fill=NA), panel.grid=element_blank()) +
  guides(size=F)

ggsave(here::here('..','figs','pooled','mixed_4v8_rel.pdf'), prolif_4v8_rel,
       height=2.5, width=7, useDingbats=FALSE)

```

# Raw bin plot (fig 2A)

```{r}
subset_barplot_reads <- read.counts[
  assay == 'CTV1' & batch == 'prolif2' & donor == 'd1' & k.type == 'pos']

subset_barplot_reads <- subset_barplot_reads[,
  CAR.ranked.sep := factor(
    paste(CAR.align,t.type,sep='|'))]

subset_barplot_reads[,
  CAR.ranked.sep := factor(CAR.ranked.sep, levels=levels(
    reorder(
      CAR.ranked.sep[k.type == 'pos'], 
      rlog_car_score[k.type == 'pos'])))]

remove_t_type_sep <- function (breaks)
  unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))

remove_t_type_sep_col <- function (breaks) {
  car_name <- unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))
  
  car_axis_colors <- rep('black', length(car_name))
  car_axis_colors[grep('CD28', car_name)] <- car_colors['CD28']
  car_axis_colors[grep('4-1BB', car_name)] <- car_colors['4-1BB']

  # bold CAR names if colored
  car_name[car_axis_colors != 'black'] <- paste0('**',car_name[car_axis_colors != 'black'],'**')
  
  car_color_span <- paste("<span style = 'color: ",
    car_axis_colors,
    ";'>",
    car_name,
    "</span>", sep = "")
  
  return(car_color_span)
}

barplot_stacked_plot <- ggplot(subset_barplot_reads) +
  geom_bar(aes(y=car.bin.pct, x=CAR.ranked.sep, fill=bin),
    stat='identity', width=1) +
  scale_fill_brewer(palette = "YlGn") +
  facet_wrap(. ~ t.type, scales = 'free_x', ncol=1) +
  labs(y='% sorted into each bin', x='CAR Costimulatory Domains, ranked') +
  scale_x_discrete(expand = c(0, 0), labels=remove_t_type_sep) +
  scale_y_continuous(limits = c(0,1), expand = expansion(mult = c(0, 0)), 
                     breaks=c(0, .5, 1), labels=percent) +
  theme_bw(base_size=18) + theme(axis.ticks = element_blank()) + 
  theme(
    axis.text.x = element_text(size=9, angle = 90, vjust = 0.5, hjust=1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    panel.border = element_rect(colour = "black"))

ggsave(here::here('..','figs','pooled','fig2a_stackbar_single_donor.pdf'), 
       barplot_stacked_plot, height=6, width=9, useDingbats=FALSE)

other_car_colors <- unique(subset_barplot_reads$CAR.align)[
  unique(subset_barplot_reads$CAR.align) %ni% c('4-1BB','CD28')]
other_car_colors <- setNames(rep(NA,length(other_car_colors)),other_car_colors)

#element_rect(fill=brewer.pal(4,'YlGn')[1])

barplot_stacked_plot_horiz <- ggplot(subset_barplot_reads) +
  geom_bar(aes(y=car.bin.pct, x=CAR.ranked.sep, fill=bin),
    stat='identity', width=1) +
  scale_fill_brewer(palette = "YlGn") +
  facet_wrap(t.type ~ ., scales = 'free_x', nrow=1) +
    labs(y='% of cells sorted into proliferative bins', 
       x='CAR Costimulatory Domains, ranked') +
  scale_x_discrete(expand = c(0, 0), labels=remove_t_type_sep_col) +
  scale_y_continuous(limits = c(-0.0175,1), expand = expansion(mult = c(0, 0)), 
                     breaks=c(0, .5, 1), labels=percent) +
  coord_cartesian(clip='off') +
  annotate(geom='rect', xmin=-Inf, xmax=Inf, ymin=0, ymax=1, color='black', fill=NA, size=0.25) +
  theme_bw(base_size=12) + 
  theme(
    axis.ticks = element_blank(),
    axis.text.x = element_markdown(size=9, angle = 90, vjust = 0.5, hjust=1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank()) +
  geom_point(data=subset_barplot_reads[CAR.align %in% c('4-1BB','CD28')],
    aes(x=CAR.ranked.sep, y=-0.0175, color=CAR.align),
      pch=17,
      size=2.8) +
  scale_color_manual(, 
    values=c(other_car_colors, car_colors, 
    '4-1BB'=car_colors[['4-1BB']]), guide='none')

ggsave(here::here('..','figs','pooled','fig2a_stackbar_single_donor_horiz.pdf'), 
       barplot_stacked_plot_horiz, height=6, width=9, useDingbats=FALSE)

```

# CTV1 and CTV2 barplot

```{r}
subset_barplot_reads <- read.counts[
  assay %in% c('CTV1') & batch %in% c('prolif2') & donor %in% c('d1','d2') & k.type == 'pos']

subset_barplot_reads <- subset_barplot_reads[,
  CAR.ranked.sep := factor(
    paste(CAR.align,t.type,batch,donor,sep='|'))]

subset_barplot_reads[,
  CAR.ranked.sep := factor(CAR.ranked.sep, levels=levels(
    reorder(
      CAR.ranked.sep[k.type == 'pos'], 
      rlog_car_score[k.type == 'pos'])))]

remove_t_type_sep <- function (breaks)
  unlist(lapply(breaks, function(str)
    strsplit(str, '|',fixed=T)[[1]][1]))

barplot_stacked_multi_plot <- ggplot(subset_barplot_reads) +
  annotate(geom='rect', xmin=-Inf, xmax=Inf, ymin=0, ymax=1, color='black',
    fill=brewer.pal(4, 'YlGn')[1], size=0.25) +
  geom_bar(aes(y=car.bin.pct, x=CAR.ranked.sep, fill=bin),
    stat='identity', width=1) +
  scale_fill_brewer(palette = "YlGn") +
  facet_wrap(sort.group ~ t.type, scales = 'free_x', ncol=2) +
  labs(y='', x='') +
  scale_x_discrete(expand = c(0, 0), labels=remove_t_type_sep) +
  scale_y_continuous(limits = c(-0.025,1), expand = expansion(mult = c(0, 0)), 
                     breaks=c(0, .5, 1), labels=percent) +
  theme_bw(base_size=18) + 
  theme(
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    panel.border = element_blank(),
    strip.text= element_blank(),
    legend.position = "none") +
  coord_cartesian(clip='off') +
  annotate(geom='rect', xmin=-Inf, xmax=Inf, ymin=0, ymax=1, color='black',
    fill=NA, size=0.25) +
  geom_point(data=subset_barplot_reads[CAR.align %in% c('4-1BB','CD28')],
    aes(x=CAR.ranked.sep, y=-0.025, color=CAR.align),
      pch=17,
      size=2.8) +
  scale_color_manual(
    values=c(other_car_colors, car_colors, '4-1BB'=car_colors[['4-1BB']]), guide='none')

ggsave(here::here('..','figs','pooled','fig2a_stackbar_both_donors.pdf'), 
       barplot_stacked_multi_plot, height=6, width=9, useDingbats=FALSE)

```

## Fig2: Compile panels

```{r}

plot_a <- plot_grid(
  barplot_stacked_plot_horiz+theme(legend.position = "none"), 
  (barplot_stacked_multi_plot / plot_spacer()) + plot_layout(heights=c(4.5,1)), 
  rel_widths=c(5,1),
  ncol = 2)

plot_b <- cytokine_subset_colors + 
  guides(shape = guide_legend(override.aes = list(fill = 'grey30') )) +
  theme(legend.justification = c(0, 1), legend.position = c(0, 1),
        legend.margin = margin(t=0, unit='cm'),
        plot.margin = unit(c(1,1,1,0.8), "cm"),
        panel.grid=element_blank())

plot_c <- fig2_abundance_over_time_vert + 
  theme(panel.grid=element_blank())

plot_d <- sprinter_plot + 
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    legend.position= 'bottom',
    legend.text = element_text(size=9, hjust=0),
    legend.title = element_text(size=9, hjust=0),
    legend.box = "vertical",
    panel.border = element_rect(colour = "black"),
    plot.margin = unit(c(0.2,0.2,0.2,0.2), "cm"),
    legend.title.align=0,
    legend.margin=margin(t = -0.3, unit='cm'),
    legend.box.just = "left",
    legend.justification = 'left')

plot_e <- prolif_4v8_rel_vert + theme(
  plot.margin = unit(c(0,1,0.5,0.8), "cm"),
  panel.grid=element_blank())

plot_abcde <- plot_grid(
  plot_a,
  plot_b,
  plot_grid(
    plot_c, plot_e, labels=c('C','D'),
    ncol=2, rel_widths=c(1.1, 1)), ncol=1, 
  nrow=3, 
  rel_heights=c(0.8, 0.9, 1.3),
  labels=c('A','B',''))

ggsave(here::here('..','figs','compiled','fig2.newdraft.pdf'), plot_abcde,
       height=10, width=11, useDingbats=FALSE)  
```

### Fig3: assembled

```{r}
fig3 <- plot_grid(
  fig_3_single_dendro,
  plot_grid(
    pca.pos.sort.type.plot + theme(
      axis.title=element_text(size=unit(9,'pt')),
      legend.margin=unit(0,"cm"),
      plot.margin=margin(5,5,5,25, unit='pt'),
      axis.title.y = element_blank(),
      legend.text = element_text(size=unit(7,'pt')),
      legend.title = element_text(size=unit(8,'pt'))),
    color_car_strips(car.abund.plot + theme(
        axis.title=element_text(size=unit(9,'pt')),
        panel.border=element_rect(fill=NA),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.box='horizontal', 
        strip.background = element_rect(colour="white", fill="white"),
        legend.justification=c(1,0),
        legend.position=c(1,0.8))),
    ncol=2, nrow=1, rel_widths = c(1.55,3), labels=c('B','C')),
  rel_heights=c(1.4,1),
  nrow=2, ncol=1, labels=c('A',''))

ggsave(here::here('..','figs','compiled','fig3.newdraft.pdf'), fig3,
       height=7, width=11, useDingbats=FALSE)
```

## FigS1: Load Exhaustion Data 
```{r}

data_dir <- here::here('..','data')

if (reload_data | !exists('exh_dt'))
  #load(file.path(s3.data.dir, 'exh.Rdata'))
  load(file.path(local.data.dir, 'exh.Rdata'))

exh_dt[, car := factor(car, levels=c(car_order,'untrans'))]

# some replicates appear to be incorrect - possibly due to evaporation or error in measurement.
# we will identify which wells are way off of the median value for the set of 3 replicates:
exh_means <- exh_dt[grep('_t',variable) & !is.na(donor), list(
  mean_val=mean(value, na.rm=T)),
  by=c('well','plate','day','variable','donor','row','col','k562','car','t_type')]

exh_means[, c('med_val','med_diff') := list(
  median(mean_val),
  (mean_val-median(mean_val))/median(mean_val)),
  by=c('day','car','k562','variable','donor','t_type')]

exh_means[, oob := abs(med_diff) >= 0.25]
exh_means[, n_oob := sum(oob), list(donor,k562,day,t_type,car,well,plate,variable)]

outlier_wells <- unique(exh_means[n_oob == 1, list(donor,k562,day,t_type,car,well,plate)])

exh_dt <- outlier_wells[, outlier_well := T][exh_dt, on=.(donor,k562,day,t_type,car,well,plate)]
exh_dt[is.na(outlier_well), outlier_well := F]

plots <- list()

cd4_marker_pos_threshold <- list(
  tim3_t= 666,
  gfp_t=1250,
  zombie_t=750,
  myc_t=1400,
  lag3_t= 950, 
  pd1_t= 1400,
  cd39_t= 1100)

cd8_marker_pos_threshold <- list(
  tim3_t= 666,
  gfp_t=1500,
  zombie_t=750,
  myc_t=1400,
  lag3_t= 800, 
  pd1_t= 1400,
  cd39_t= 1250)

censor_negative_min = -1500
```

### Fig1: Exh panel

```{r}
# donor 1 CD8 exh data for main text:
fig1_exh_panel <- stacked_exh_plot(day_dt[t_type == 'cd8' & day %in% c(0,15,24,33) & donor == 1][,
    donor := factor(paste0('Donor ',donor))]) +
  labs(x='Day') +
  scale_y_continuous(expand=c(0,0), labels=label_percent(), breaks=c(0,.5,1)) +
  guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5,
    label.position = "bottom")) +
  theme(
    legend.position = 'top',
    strip.text.x = element_blank(),
    panel.spacing.y = unit(0.8, "lines"),
    panel.background = element_rect(fill='grey70'),
    legend.key.size = unit(0.8,'lines'),
    plot.subtitle = element_text(hjust=0.5, margin=margin(0,0,0,0,'pt')),
    axis.title.y=element_blank())

ggsave(here::here('..','figs','compiled','fig1_exh.pdf'), fig1_exh_panel,
       height=4, width=1.5, useDingbats=FALSE)
```

## FigS1D: compare abundance and length to expansion and proliferation

```{r}
# mean measures by car
car_len_abund_comparisons <- read.counts[
  k.type == 'pos'][,
  `:=`(
    mean_rlog_car_score= mean(rlog_car_score[assay=='CTV1'], na.rm=T),
    meanlog2.abund.rel.baseline= mean(log2(car.abund.rel.baseline[assay=='CTV3']), na.rm=T),
    baseline.car.abund= mean(log2(baseline.car.abund), na.rm=T)),
  by=c('CAR.align')]


d3_ctv <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=mean_rlog_car_score, x= baseline.car.abund)) + 
  labs(
    y='CTV score, d3',
    x='log2 initial abundance',
    subtitle='Init. Lib. Abund. vs d3 Prolif.')

d24_exp <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=meanlog2.abund.rel.baseline, x= baseline.car.abund)) + 
  labs(
    y='Relative Expansion, d24',
    x='log2 initial abundance',
    subtitle='Init. Lib. Abund. vs d24 Exp.')

d3_ctv_len <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=mean_rlog_car_score, x= log2(len))) + 
  labs(
    y='CTV score, d3',
    x='log2 domain length',
    subtitle='Domain Length vs d3 Prolif.')

d24_exp_len <- ggplot(car_len_abund_comparisons) + 
  geom_point(aes(y=meanlog2.abund.rel.baseline, x= log2(len))) + 
  labs(
    y='Relative Expansion, d24',
    x='log2 domain length',
    subtitle='Domain Length vs d24 Exp.')

len_abund_vs_outcome <- (d3_ctv | d24_exp) / (d3_ctv_len | d24_exp_len) +
  plot_annotation(title='No correlation between initial library abundance\nand proliferation/expansion:')

len_abund_vs_outcome_horiz <- (d3_ctv | d24_exp | d3_ctv_len | d24_exp_len) +
  plot_annotation(title='No correlation between initial library abundance\nand proliferation/expansion:')


supp_S1D_len_abund_vs_outcome <- len_abund_vs_outcome_horiz &
  theme_minimal() + 
  theme(panel.border=element_rect(fill=NA),
        panel.grid = element_blank())

```

### FigS1: 28, BB, Zeta exhaustion data

```{r}

s1_car_set <- c('CD28','41BB','zeta','untrans')
s1_car_colors <- c(
  car_colors[intersect(names(car_colors), s1_car_set)],
  untrans='grey50')

get_marker_counts <- function(exh_dt, exh_vars=c('tim3_t','lag3_t', 'pd1_t','cd39_t')) {
  marker_ct <- exh_dt[!outlier_well & variable %in% exh_vars,
    list(n_pos_marker= sum(gt_thresh)),
    by=c('well','plate','day','Time','car','donor','k562','t_type')]
  
  marker_ct <- marker_ct[, N := .N, by=c('well','plate','day','car','k562','t_type')][N > 1]
  
  marker_ct <- map_day_0(marker_ct)
  
  marker_freq <- marker_ct[, data.frame(table(factor(n_pos_marker, levels= 0:length(exh_vars)))), 
    by=c('well','plate','day','car','donor','k562','t_type')]
  names(marker_freq)[names(marker_freq) == 'Var1'] <- 'n_pos_marker'
  
  marker_freq[ n_pos_marker != 7, pct := Freq/sum(Freq), 
    by=c('well','plate','day','car','donor','k562','t_type')]
  
  return(marker_freq)
}

if (!exists('day_dt'))
  day_dt <- get_marker_counts(
    exh_dt,
    exh_vars=c('tim3_t','lag3_t', 'pd1_t','cd39_t'))[
      day %in% c(0,6,15,24,33) &
      car %in% s1_car_set &
      k562 == 'cd19+']

stacked_exh_plot <- function(day_dt_subset) {
  
  day_dt_subset <- day_dt_subset[, 
      c('mean_pct','sd_pct') := list(mean(pct), sd(pct)), 
      by=c('n_pos_marker','t_type','donor','day','k562','car')][,
        t_type := toupper(t_type)]
  
  day_dt_subset[, car := factor(car)]
  day_dt_subset[, pct := as.numeric(pct)]
  
  ggplot(unique(day_dt_subset[, list(car, mean_pct, n_pos_marker, day, t_type, donor)])) +
    geom_bar(aes(x=factor(day, levels=rev(levels(factor(day)))), y=mean_pct, fill=n_pos_marker),
      stat='identity', width=1, color='black', size=0.2) +
    facet_grid(car ~ donor, scales='free') +
    scale_y_continuous(expand=c(0,0), labels=label_percent(), breaks=c(.5,1)) +
    scale_x_discrete(expand=c(0,0), limits=rev) +
    scale_fill_manual(
      'Exh. Markers per cell', values=brewer.pal(9,'RdPu')[c(1,2,5,7,9)]) +
    labs(x="Days in Culture", y='Fraction of Cells') +
    theme_minimal() +
    theme(
      panel.grid=element_blank(),
      panel.border=element_rect(color='black', size=0.2, fill=NA))

}

plots$exh_stacked_bars <- ((
  stacked_exh_plot(day_dt[t_type == 'cd4' & day %in% c(0,15,24,33)][,
      donor := factor(paste0('Donor ',donor))]) +
    labs(subtitle='CD4', tag='B') +
    theme(
      panel.background = element_rect(fill='grey70'),
      legend.position='bottom',
      plot.subtitle = element_text(hjust=0.5, margin=margin(0,0,0,0,'pt')),
      axis.title=element_blank())
  ) / (
  stacked_exh_plot(day_dt[t_type == 'cd8' & day %in% c(0,15,24,33)][,
      donor := factor(paste0('Donor ',donor))]) +
    labs(subtitle='CD8', x='Day') +
    theme(
      panel.background = element_rect(fill='grey70'),
      legend.position='bottom',
      plot.subtitle = element_text(hjust=0.5, margin=margin(0,0,0,0,'pt')),
      axis.title.y=element_blank())
  )) + 
  plot_layout(guides='collect')

indiv_marker_means <- map_day_0(exh_dt)[(
    variable %in% c('tim3_t','lag3_t', 'pd1_t','cd39_t') &
    outlier_well==F &
    k562 == 'cd19+' &
    car %in% s1_car_set
  ),
  list(
    mean_val=mean(value),
    sem_val=sqrt(var(value)/.N)),
  by=.(donor, k562, day, t_type, car, variable)]

indiv_marker_means[, variable := factor(variable)]
indiv_marker_means[, variable := factor(variable,
  labels=toupper(gsub('_t','',levels(variable))))]

plot_indiv_marker_t_type <- function(indiv_marker_mean_subset) {
  ggplot(indiv_marker_mean_subset, 
       aes(x=day, color=car, y=mean_val)) + 
  geom_line() +
  geom_linerange(
    aes(
      ymin=mean_val-sem_val/2,
      ymax=mean_val+sem_val/2,
      group=interaction(car, day, variable, donor))) +
  geom_point() +
  scale_color_manual(values=s1_car_colors) +
  facet_grid(variable ~ donor, scale='free_y') +
  theme_minimal() +
  theme(
    panel.border = element_rect(fill=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position='bottom',
    strip.background = element_blank())
}

plots$marker_means <- (
  plot_indiv_marker_t_type(indiv_marker_means[t_type == 'cd4'][, 
      donor := factor(paste0('Donor ',donor))][,
      variable := factor(variable, labels=levels(variable))]) +
    labs(subtitle='CD4', tag='A') +
    theme(
      plot.subtitle = element_text(hjust=0.5, margin=margin(0,0,0,0,'pt')),
      axis.title=element_blank())
  ) | (
  plot_indiv_marker_t_type(indiv_marker_means[t_type == 'cd8'][, 
      donor := factor(paste0('Donor ',donor))]) +
    labs(subtitle='CD8', x='Day') +
    theme(
      plot.subtitle = element_text(hjust=0.5, margin=margin(0,0,0,0,'pt')),
      axis.title.y=element_blank())
  ) + 
  plot_layout(guides='collect')
```

### Exhaustion stats (CD28, zeta, BB)
These stats were used for calculating significance for Figure 1E.

```{r, eval=F}
exh_stat_dt <- get_marker_counts(
    exh_dt,
    exh_vars=c('tim3_t','lag3_t', 'pd1_t'))[
      k562 == 'cd19+' & car %in% c('41BB','CD28','zeta','untrans')][,
      donor := factor(paste0('Donor ',donor))][, 
        list(
          mean.markers= sum(as.numeric(n_pos_marker) * pct)/sum(pct),
          pct.sum =sum(pct)),
        by=.(plate, day, car, donor, k562, t_type)][,
        day := factor(day, ordered=T)]

ggplot(exh_stat_dt,
  aes(x=day, y=mean.markers, 
      color=car, group=interaction(car, donor, t_type))) + 
  geom_line() +
  geom_point() + 
  facet_wrap(donor ~ t_type) +
  scale_color_manual(values=car_colors)

aov_formula <- mean.markers~day*car + Error(donor)

overall_model <- aov(
  formula=aov_formula,
  data=exh_stat_dt[t_type == 'cd8'])

post_hoc_pval <- function(car_pair) {
  model <- aov(
    formula=aov_formula, 
    data=exh_stat_dt[car %in% car_pair])
  aov_df <- data.table(summary(model)$`Error: Within`[[1]], keep.rownames = T)
  
  aov_df[, car_a := car_pair[1]]
  aov_df[, car_b := car_pair[2]]
  
  return(aov_df)
}

#cars ordered by mean exhaustion rank
cars_exh_order <- as.character(
  exh_stat_dt[, 
    list(car, rank=rank(mean.markers)), by=.(donor,day,t_type)][,
      list(mean.rank= mean(rank)), by='car'][order(mean.rank), car])

all_pairs <- rbindlist(
  apply(combn(cars_exh_order, 2), 2, post_hoc_pval))
all_pairs[, p.adj := p.adjust(`Pr(>F)`, method='BH'), by=rn]
all_pairs[, car_a := factor(car_a, levels=cars_exh_order)]
all_pairs[, car_b := factor(car_b, levels=cars_exh_order)]
all_pairs[, rn := str_trim(rn)]
all_pairs <- add_significance(all_pairs, p.col='p.adj')

# showing both car and car:day effects
ggplot(all_pairs[rn %in% c('car','day:car')],
    aes(x=car_a, y=car_b)) + 
  geom_tile(
    aes(fill=-log10(p.adj))) + 
  geom_tile(
    data=all_pairs[rn %in% c('car','day:car')][p.adj < 0.05],
    color='black', fill=NA, size=1) +
  scale_fill_distiller(palette="RdPu", direction=1) +
  facet_wrap(~rn) + 
  scale_color_manual(values=c('white','black')) + 
  theme_minimal() + 
  plot_theme + 
  theme(legend.position='right')

# only car effect, with significance *s

ggplot(all_pairs[rn == 'day:car'],
    aes(x=car_a, y=car_b, 
        fill=-log10(p.adj),
      label=p.adj.signif)) + 
  geom_tile(color='white') +
  scale_fill_gradientn(colors=brewer.pal(9, 'RdPu')[1:7], 
    guide=guide_colorbar(
      barheight=0.7,
      barwidth=4,
      direction = "horizontal",
      title.position = "top")) +
  geom_text() + 
  labs(x='', y='Overall Exhaustion Marker Rank', fill='-log10(p),\nRM ANOVA') +
  theme_minimal() + 
  theme(panel.grid=element_blank(),
    legend.position=c(1,0.05),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    legend.justification=c(1,0))

```


## Replicates (pairs) comparisons (not using)

Comparisons on x/y, all combos
```{r, eval=F}

x_group <- 'prolif2_d1_3d_CTV1_CD4_neg'
y_group <- 'prolif2_d1_3d_CTV1_CD4_pos'

cast_comparison <- function(
    comp.df, x_group, y_group, value_col='CAR.score', xycols=c('x','y'),
    rescale= 'combined') {
  
  #message(paste(x_group, y_group, sep=', '))
  
  cast_groups <- dcast(
    unique(comp.df[sort.group %in% c(x_group, y_group), 
        list(CAR.align, sort.group, get(value_col))]), 
    CAR.align ~ sort.group, value.var = 'V3')[, 
      `:=`(x.group= x_group, y.group= y_group)]
  
  names(cast_groups)[c(2,3)] <- xycols
  
  stopifnot(rescale %in% c('combined','separate'))
  
  # rescle == combined:
  # rescale both x and y to (0,1) on same scale
  xy_scalemin = cast_groups[, min(c(x,y), na.rm=T)]
  xy_scalemax = cast_groups[, max(c(x,y), na.rm=T)]
  tryCatch({
    cast_groups[, x_scaled_comb := scales::rescale(
      x, from=c(min(c(x,y), na.rm=T), max(c(x,y), na.rm=T)))]
    cast_groups[, y_scaled_comb := scales::rescale(
      y, from=c(min(c(x,y), na.rm=T), max(c(x,y), na.rm=T)))]
  }, error = function(e) {
    message(e)
    cast_groups[, x_scaled_comb := NaN]
    cast_groups[, y_scaled_comb := NaN]}
  )
    
  # rescle == separate:
  # rescale x and y to (0,1) on individual scales
  tryCatch({
    cast_groups[, x_scaled_sep :=  scales::rescale(x)]
    cast_groups[, y_scaled_sep :=  scales::rescale(y)]
  }, error = function(e) {
    message(e)
    cast_groups[, x_scaled_sep := NaN]
    cast_groups[, y_scaled_sep := NaN]}
  )
}

# use malanhanobis distance to collapse replicates, 
# then use median absolute deviation to identify outliers
#https://www.r-craft.org/r-news/combined-outlier-detection-with-dplyr-and-ruler/

maha_dist <- . %>% select_if(is.numeric) %>%
    mahalanobis(center = colMeans(.), cov = cov(.))

isnt_out_maha <- function(tbl, isnt_out_f, ...) {
  tbl %>% maha_dist() %>% isnt_out_f(...)
}

isnt_out_mad <- function(x, thres = 3, na.rm = TRUE) {
  abs(x - median(x, na.rm = na.rm)) <= thres * mad(x, na.rm = na.rm)
}

top_x_mad <- function(x, top=5, na.rm = TRUE) {
  mads <- abs(x - median(x, na.rm = na.rm))
  top_mads <- order(-mads)[1:top]
  return(!(1:length(mads) %in% top_mads))
}


plot_all_reps <- function(df=read.counts, value_col, df_only=F, ordinal=F) {
  
  all_rep_comparisons <- df[
    grepl('CTV', assay)][order(assay)][order(t.type)][order(k.type)][, 
    data.table(matrix(combn(unique(sort.group), 2), ncol=2, byrow=T)),
    by=c('assay','t.type','k.type')]
  
  names(all_rep_comparisons)[4:5] <- c('x.group','y.group')
  
  all_rep_comparisons <- all_rep_comparisons[, 
    cast_comparison(df, x.group, y.group, value_col=value_col)[, 
      c('x.group','y.group') := NULL],
    by=c('assay','t.type','k.type','x.group','y.group')]
  
  # combine with baseline abundance as color
  baseline.abund <- df[
    assay=='baseline', list(mean.baseline= mean(car.abund.baseline, na.rm=T)),
    by=c('t.type','CAR.align')][,
      list(CAR.align, 
        rel.baseline.log= log10(mean.baseline/mean(mean.baseline))), 
      by=c('t.type')]
  
  all_rep_comparisons <-all_rep_comparisons[
    baseline.abund, on=c('t.type','CAR.align')]
  
  all_rep_comparisons[, rep_pair := paste(
    gsub('(prolif\\d_d\\d).*','\\1', x.group), 
    gsub('(prolif\\d_d\\d).*','\\1', y.group), 
    sep='\n')]
  
  all_rep_comparisons[
    rep_pair == "prolif2_d2\nprolif1_d2", rep_pair := "prolif1_d2\nprolif2_d2"] 
  
  all_rep_comparisons[, assay_kt := paste(assay,t.type,k.type, sep='\n')]
  
  # use malanhanobis distance to collapse replicates, 
  # then use median absolute deviation to identify outliers
  all_rep_comparisons[,
    outlier := {   
      
      nonsingular <- apply(tibble(x_scaled_sep, y_scaled_sep), 2, 
        function (x) var(x) > 0 & !is.na(var(x)))
      non_singular_sd <- tibble(x_scaled_sep, y_scaled_sep)[,
        (nonsingular)]
      !isnt_out_maha(non_singular_sd, top_x_mad)
    }, by=.(rep_pair, k.type, t.type, assay)]
  
  all_rep_comparisons[, label_outliers := ''][outlier == T, 
    label_outliers := CAR.align]
  
  all_rep_comparisons[, label_arrayed := ''][
    CAR.align %in% array.list & !outlier, 
    label_arrayed := CAR.align]
  
  non_na_comps <- all_rep_comparisons[, 
    !any(is.na(list(var(x_scaled_sep), var(y_scaled_sep)))), 
    by=.(rep_pair, k.type, t.type, assay)]
  
  if (df_only) return(all_rep_comparisons)
  
  r_squareds <- all_rep_comparisons[
    non_na_comps[V1==T], on=.(rep_pair, k.type, t.type, assay)][V1 == T][, 
      list(r=sqrt(summary(lm(x_scaled_sep ~ y_scaled_sep))$r.squared)), 
      by=.(rep_pair, k.type, t.type, assay)]
  
  r_squareds[, r_lbl := paste('r=',as.character(round(r, 2)))]
  
  return(ggplot(data=all_rep_comparisons, 
      aes(x_scaled_sep, y_scaled_sep)) + 
    geom_point(shape = 21, colour = "grey30", aes(fill=rel.baseline.log)) + 
    geom_text_repel(size=2.5, color='grey20', aes(label=label_outliers)) +
    geom_text_repel(size=2.5, color='lightsalmon4', aes(label=label_arrayed)) + 
    theme_bw() + 
    scale_fill_distiller(palette='BrBG', 
      limit=c(-1,1) * max(abs(all_rep_comparisons$rel.baseline.log))) +
    geom_label(size=2.5, aes(x=Inf, y=-Inf, label=r_lbl), data=r_squareds,
      hjust=1, vjust=0) +
    facet_grid(rep_pair ~  k.type + t.type + assay))

}

assay_fig_labels <- c(
  'Initial\nProliferation\n(d0-d3/d4)', 
  'Relative\nExpansion\n(d0-d16)', 
  'Relative\nExpansion,\n(d0-d24)')

# combined rlog car score for CTV1 & CTV2, baseline abundance for CTV3
fig2_rep_comp <- rbind(
    plot_all_reps(
        df=read.counts[batch != 'post-cytof'], 
        value_col='rlog_car_score', df_only=T)[assay == 'CTV1'],
    plot_all_reps(
        df=read.counts[batch != 'post-cytof' & CAR.align != 'TNR8'], 
        value_col='car.abund.rel.baseline', df_only=T)[assay != 'CTV1'])

fig2_rep_comp[, assay := factor(assay, labels=assay_fig_labels)]

fig2_rep_comp[, x.rank := rank(x_scaled_sep), by=.(assay, x.group, y.group)]
fig2_rep_comp[, y.rank := rank(y_scaled_sep), by=.(assay, x.group, y.group)]

replicate_comp_plot <- ggplot(data=fig2_rep_comp, 
       aes(x_scaled_sep, y_scaled_sep)) + 
    geom_point(aes(color=interaction(rep_pair, k.type)), size=0.8) +
    scale_x_continuous(labels=scales::number_format(accuracy = .1), breaks=c(0,0.5,1)) +
    scale_y_continuous(labels=scales::number_format(accuracy = .1), breaks=c(0,0.5,1)) +
    theme_minimal() +
    coord_fixed() +
    theme(panel.border=element_rect(fill=NA), panel.grid=element_blank()) + 
    facet_grid(. ~t.type + assay) +
    scale_color_manual('', 
        values=brewer.pal(11,'PRGn')[c(2,3,4,10,9,8)],
        labels=c(
            'CD19-, A v B', 'CD19-, A v C', 'CD19-, B v C',
            'CD19+, A v B', 'CD19+, A v C', 'CD19+, B v C')) +
    labs(x='Scaled replicate 1', y='Scaled replicate 2')

supp_S2_1_replicate_comp_plot <- replicate_comp_plot
```

### Fig S1C: Replicate Comparison Ranking

```{r}
## all reps

all_reps <- rbind(
  # CTV1: early proliferation
  unique(read.counts[grepl('CTV', assay) & k.type == 'pos' & assay == 'CTV1'][,
    list(CAR.align, assay, t.type, donor, batch,
         metric=rlog_car_score, sort.group)]),
  # CTV2/3: late expansion
  unique(read.counts[grepl('CTV', assay) & k.type == 'pos' & assay %in% c('CTV2','CTV3')][, 
    list(CAR.align, assay, t.type, donor, batch,
         metric=car.abund.rel.baseline, sort.group)]),
  #cytokines
  cyto.plot.data[,
    list(CAR.align, assay=sort.cond, t.type, donor, batch='prolif1', 
        metric=car.bin.pct.incl, sort.group)])[,
  car.score.rank := rank(metric), by= c('sort.group','assay')][,
  car.score.rank.mean := mean(car.score.rank), by='CAR.align'][,
  car.score.rank.rev := rank(-metric), by=c('sort.group','assay')]

all_reps[, donor := factor(paste0(donor, batch), labels=c('A','B','C')), by=assay]

#add assay means
all_reps <- all_reps[, car.score.rank.mean.assay := mean(car.score.rank),
  by=.(assay, CAR.align)]
all_reps <- all_reps[, CAR.align.assay := reorder(factor(
  paste0(CAR.align, assay)),
  car.score.rank.mean.assay)]

kruskal_test_all_reps <- kruskal.test(
  car.score.rank ~ CAR.align, data = all_reps)

ordered_cars <- all_reps[grepl('CTV',assay), levels(reorder(droplevels(CAR.align), car.score.rank.rev))]

#CD28 and 4-1BB ranks
which(rev(ordered_cars) %in% c('CD28', '4-1BB'))

#p values for 4-1BB and CD28, MWU
copy(all_reps)[, is_CD28 := CAR.align == 'CD28'][, wilcox_test(.SD, car.score.rank ~ is_CD28)]$p
copy(all_reps)[, is_41BB := CAR.align == '4-1BB'][, wilcox_test(.SD, car.score.rank ~ is_41BB)]$p

assay_kw_signif <- all_reps[, 
    kruskal.test(car.score.rank ~ CAR.align, data = .SD), by=.(assay)][,
    signif_label := paste0('K-W H= ',round(statistic),'\np= ',signif(p.value,2))]

s1c_ordinal_plots <- ggplot(all_reps) + geom_jitter(
    aes(
      x=as.numeric(reorder(CAR.align.assay, car.score.rank.mean))/5, 
      y=car.score.rank, 
      color=interaction(donor, t.type)),
    width=0.5, height=0.5, size=1) + 
  theme_minimal() + 
  scale_color_manual('', 
    values=brewer.pal(11,'PRGn')[c(2,3,4,10,9,8)],
    labels=paste(
      'Donor',
      rep(c('A','B','C'),2),
      rep(c('CD4','CD8'), each=3))) +
  coord_flip() +
  theme_minimal() +
  facet_grid(.~factor(assay,
      levels=c(paste0('CTV',1:3),c('ifng','il2')),
      labels=c('d3 Prolif.','d0-16 Rel. Expansion', 'd0-24 Rel. Expansion',
               'IFNg','IL2')),
    scales='free') +
  theme(panel.border=element_rect(fill=NA),
    axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5),
    panel.grid = element_blank(),
    legend.title = element_blank(),
    legend.background =element_rect(colour="white", fill="white"),
    legend.box='horizontal', 
    strip.background = element_rect(colour="white", fill="white")) + 
  labs(x='Average Rank across Donors', y='Individual Rank for each Donor') +
  geom_text(data=assay_kw_signif, aes(label=signif_label),
           x=39, y=1, vjust=1, hjust=0, size=2.5)

```



## Fig S1 Assemble

```{r}

supp_S1_1 <- (
  (
    (
      plots$marker_means &
        guides(color=guide_legend(name='CAR',
          direction = "horizontal",
          title.position = "top",
          label.position = "bottom")) &
        theme(legend.position = 'bottom')
    ) | (
      plots$exh_stacked_bars &
        guides(fill=guide_legend(name='CAR',
          direction = "horizontal",
          title.position = "top",
          label.position = "bottom",
          keywidth=unit(1, 'lines'),
          keyheight=unit(1, 'lines'))) &
        theme(legend.position = 'bottom'))
  ) / 
    (s1c_ordinal_plots + theme(aspect.ratio = 1, legend.position = 'bottom')) /
    (supp_S1D_len_abund_vs_outcome & theme(aspect.ratio = 1))
  ) + 
  plot_layout(
    heights=c(2,1,1))
  
ggsave(here::here('..','figs','compiled','figS1.newdraft.pdf'), supp_S1_1,
       height=13, width=11, useDingbats=FALSE)

```


### FigS2: CD69 flowseq
```{r}

cd69.pct.pos <- dcast(
  read.counts[batch == "post-cytof" & assay == 'CD69'][
    bin %in% c('C','D'), list(pct_pos= sum(car.bin.pct)), 
      by=c('assay','CAR.align', 'k.type','t.type')],
  assay + CAR.align + t.type ~ k.type, value.var='pct_pos')

cd69.pct.pos[, CAR.align.assay := factor(paste(CAR.align,t.type,sep='|'))]
cd69.pct.pos[, CAR.align.assay := factor(
  CAR.align.assay, 
  levels=levels(CAR.align.assay),
  labels=cd69.pct.pos[, gsub('(.*)\\..*', '\\1', levels(CAR.align.assay))])]

cd69_panel <- ggplot(cd69.pct.pos,
  aes(x = reorder(CAR.align.assay, -(neg-pos)),
      xend = reorder(CAR.align.assay,  -(neg-pos)),
      y = neg,
      yend = pos)) + 
  geom_segment(color="grey") +
  geom_point(
    aes(y = pos, color='CD19+'), 
    size=2.5) +
  geom_point(
    aes(y = neg, color='CD19-'), 
    size=2.5) + 
  scale_color_manual(
      values=c('CD19+'=cd19_val[2], 'CD19-'='grey30')) + 
  facet_wrap(~ t.type, scales='free_x', ncol=1) +
  scale_y_continuous(label=percent) +
  scale_x_discrete(labels= remove_t_type_sep_col) +
  labs(y = 'CD69 % Positive') +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA),
        axis.text.x = element_markdown(angle = 90, hjust = 1, vjust=0.5, size=rel(0.8)),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        strip.background = element_rect(colour="white", fill="white"),
        legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(0,1.01))

ggsave(here::here('..','figs','pooled','cd69_panel.pdf'), cd69_panel,
       height=3, width=8, useDingbats=FALSE)

supp_S2_2_cd69_panel <- cd69_panel
```

### FigS2: Volcano Plot

```{r}
max_pval_y <- 6.5

#CD28 and 4-1BB Colors
receptor_cols <- RColorBrewer::brewer.pal(9, 'YlGnBu')[c(6,8)]

# label significant hits
data.dt[, CAR.type.size := CAR.type]
data.dt[CAR.type == 'other' & 
    ((padj < 0.05 & log2FoldChange > 0) | 
    (padj < 0.05 & log2FoldChange > 0)), 
  CAR.type.size := 'other_sig']

data.dt[, CAR.label.sig := '']
data.dt[CAR.type.size != 'other',
    CAR.label.sig := CAR.align]

# plot subset - CTV1, CTV2, and cumulative of all 3
data.subset.dt <- data.dt[
    k.type == 'pos' & 
        ((ref_set == 'baseline' & test_set == 'A' & assay != 'CTV1') | 
        (ref_set == 'CD' & test_set == 'A' & assay == 'CTV1'))]

# rename facets
data.subset.dt[, assay := factor(assay, labels=assay_fig_labels)]

# max out at log10-15
data.subset.dt[, padj.disp := -log10(padj)]
data.subset.dt[, CAR.pvalmax := padj.disp > max_pval_y]
data.subset.dt[padj.disp > max_pval_y, padj.disp := max_pval_y-0.18]

data.subset.dt[, order := ifelse(CAR.type=="other", 2, 1)]

interbin_volcano <- ggplot(data.subset.dt[order(CAR.type)], 
  aes(
    x=log2FoldChange, y=padj.disp,
    color=CAR.type.size,
    fill=CAR.type.size,
    label=CAR.label.sig,
    size=CAR.type.size,
    shape=CAR.pvalmax)) + 
  geom_point() + 
  facet_grid(t.type ~ assay) +
  scale_color_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28', 'New Receptors'),
    values=c('grey50', receptor_cols, outlier_cols[2]),
    guide=F) +
  scale_fill_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28', 'New Receptors'),
    values=c('grey50', receptor_cols, outlier_cols[2]),
    guide=F) +
  scale_shape_manual(values=c(21,24), guide=F) +
  scale_size_manual('',
    labels=c('New Receptors', '4-1BB', 'CD28', 'New Receptors'),
    values=c(1.5,3,3,2.5), guide=F) +
  geom_hline(yintercept=-log10(0.05), linetype = 'dashed', alpha=.3) +
  labs(x='Relative Proliferation/Expansion,\nlog2 FC', 
       y='-log10(p)') + 
  scale_y_continuous(limits=c(0, max_pval_y), expand=c(.05,.05,0,0)) +
  scale_x_continuous(limits=c(-3.25,3.25), breaks=c(-3:3)) +
  geom_text_repel(
    size=2,
    point.padding=0.1,
    text.padding=0.1,
    segment.color='grey',
    max.overlaps=Inf) +
  theme_minimal() +
  theme(panel.border=element_rect(fill=NA), panel.grid=element_blank())

supp_S2_5_interbin_volcano <- interbin_volcano
```

### FigS2: CD19 neg pos comparison
```{r}

pos_neg_comparisons <- read.counts[grepl('CTV', assay), {
  if (length(unique(sort.group)) == 2)
    cast_comparison(
      .SD,
      unique(sort.group[k.type == 'neg']),
      unique(sort.group[k.type == 'pos']))
  },
  by=c('donor','assay','batch','t.type')][assay == 'CTV1']

pos_neg_comparisons[, y_scaled_comb_rank := rank(y_scaled_comb), by='y.group']
pos_neg_comparisons[, x_scaled_comb_rank := rank(x_scaled_comb), by='x.group']

pos_neg_comparisons_melt <- melt(
  pos_neg_comparisons[, 
    xy_ratio := x_scaled_comb/y_scaled_comb], measure.vars= grep(
  '^[xy][^.]', names(pos_neg_comparisons), perl=T, value=T)
  )[, 
    `:=`(var_mean= mean(value, na.rm=T), var_sd= sd(value, na.rm=T)), 
    by=c('t.type','CAR.align','variable')]

cast_left <- paste(names(pos_neg_comparisons_melt)[1:6], collapse = ' + ')
cast_right <- 'variable'
cast_formula <- paste(cast_left, cast_right, sep= ' ~ ')

pos_neg_assay_avg <- dcast(
  pos_neg_comparisons_melt, cast_formula, value.var='var_mean')

pos_neg_assay_plot <- unique(
  pos_neg_assay_avg[, list(t.type, y_scaled_comb, x_scaled_comb, y_scaled_comb_rank, x_scaled_comb_rank, CAR.align)])
pos_neg_assay_plot[, label.few := '']
pos_neg_assay_plot[
  CAR.align %in% c('BAFF-R','CD30','4-1BB','TACI','CD28','KLRG1','CD40','NTB-A','CD2'),
  label.few := CAR.align]

# stats
pos_neg_stats <- rbind(
  data.table(
    test='spearman',
    t.type=c('CD4','CD8'),
    rbind(
      with(pos_neg_assay_plot[t.type=='CD4'],
        cor.test(y_scaled_comb_rank, x_scaled_comb_rank, method='spearman'))[
          c('estimate','p.value')],
      with(pos_neg_assay_plot[t.type=='CD8'],
        cor.test(y_scaled_comb_rank, x_scaled_comb_rank, method='spearman'))[
          c('estimate','p.value')])),
  unique(pos_neg_assay_plot[,
    cor.test(y_scaled_comb, x_scaled_comb, method='pearson'), by='t.type'][,
      list(t.type, estimate, p.value, test='pearson')])
)
pos_neg_stats$estimate <- unlist(pos_neg_stats$estimate)
pos_neg_stats$p.value <- unlist(pos_neg_stats$p.value)

fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}

# with axis cuts
ggplot_pos_neg_assay <- ggplot(pos_neg_assay_plot, 
       aes(x=x_scaled_comb, y=y_scaled_comb, label=label.few)) + 
    geom_point(color='grey30') + 
    facet_grid( ~ t.type, scale='free') + 
    scale_x_continuous(expand=expansion(mult=c(.1,.1), add=0), limits=c(0,NA)) +
    scale_y_continuous(limits=c(0.48, 0.985), expand=expansion(mult=c(.05,.05), add=0)) +
    geom_text_repel(size=2.25, max.overlaps=Inf) +
    geom_text(data=pos_neg_stats[test=='pearson'], 
      mapping=aes(label=paste0("r == ", round(estimate,2))), 
      position=position_nudge(x = 1),
      x=-Inf, y=Inf,
      size=2.8, vjust=1, hjust=0, parse=TRUE) +
    geom_text(data=pos_neg_stats[test=='pearson'], 
      mapping=aes(label=paste0("p == ", fancy_scientific(signif(p.value,2)))), 
      position=position_nudge(x = 1),
      x=-Inf, y=Inf,
      size=2.8, vjust=1.8, hjust=0, parse=TRUE) +
    labs(
        #title='Scaled Mean Proliferation',
        y='CD19+ Proliferation', 
        x='CD19- Proliferation') +
    theme_minimal() +
    theme(
      panel.spacing.y=unit(1.5, "lines"),
      panel.border=element_rect(color='black', fill=NA),
      panel.grid=element_blank())

ggplot_pos_neg_assay_rank <- ggplot(pos_neg_assay_plot, 
       aes(x=x_scaled_comb_rank, y=y_scaled_comb_rank, label=label.few)) + 
    geom_point(color='grey30') + 
    facet_grid( ~ t.type, scale='free') + 
    geom_text_repel(size=2.25, max.overlaps=Inf) +
    geom_text(data=pos_neg_stats[test=='spearman'], 
      mapping=aes(label=paste0("rho == ", round(estimate,2))), 
      position=position_nudge(x = 1),
      x=-Inf, y=Inf,
      size=2.8, vjust=1, hjust=0, parse=TRUE) +
    geom_text(data=pos_neg_stats[test=='spearman'], 
      mapping=aes(label=paste0("p == ", fancy_scientific(signif(p.value,2)))), 
      position=position_nudge(x = 1),
      x=-Inf, y=Inf,
      size=2.8, vjust=1.8, hjust=0, parse=TRUE) +
    labs(
        #title='Proliferation Rank',
        y='CD19+ mean rank', 
        x='CD19- mean rank') +
    theme_minimal() +
    theme(
      #strip.text=element_blank(),
      panel.spacing.y=unit(1.5, "lines"),
      panel.border=element_rect(color='black', fill=NA),
      panel.grid=element_blank())

#supp_S2_4_pos_neg <- ggplot_pos_neg_assay
supp_S2_4_pos_neg <- (ggplot_pos_neg_assay / ggplot_pos_neg_assay_rank)

```


### Assemble Supp Fig 2

```{r}


supp_fig_2 <- plot_grid(
  plot_grid(
    plot_grid(
      supp_S2_5_interbin_volcano,
      supp_S2_2_cd69_panel,
      ncol=2, labels=c('A','B')),
    cytokine_panel,
    plot_grid(
      supp_S2_4_pos_neg & theme(aspect.ratio=1),
      sprinter_plot,
      nrow=1,
      rel_widths=c(1.3,1),
      labels=c('D','E'), align='vh'),
    ncol=1,
    rel_heights=c(2,1,2.2),
    labels=c('A','C','','')))
  
fig_scaling <- 1
ggsave(here::here('..','figs','compiled','figS2.newdraft.pdf'), supp_fig_2,
       height=14*fig_scaling, width=11*fig_scaling, useDingbats=FALSE)

```
### Supp Fig 3 - expression differences

```{r}

if (reload_data | !exists('diff_dt')) {
  #load(file.path(s3.data.dir, 'diff.sampled.Rdata'))
  #diff_dt <- fread(file.path(s3.data.dir, 'diff.sampled.csv.gz'))
  load(file.path(local.data.dir, 'diff.sampled.Rdata'))
  diff_dt <- fread(file.path(local.data.dir, 'diff.sampled.csv.gz'))
}

map_day_0_diff <- function(df) {
  return(rbind(
    df[k562!='none'],
    df[k562=='none'],
    df[k562=='none' & day == 0][, k562 := 'cd19+'],
    df[k562=='none' & day == 0][, k562 := 'cd19-']
  ))
}

diff_cast_dt <- dcast(diff_dt[!is.na(value)],
  t_type + donor + k562 + day + car + well + plate + event_id + col + row ~ variable,
  value.var='value')

mfi_myc_summ <- melt(diff_cast_dt[!is.na(car) & car != 'untrans', list(
  myc_to_gfp=mean(myc_t/gfp_t),
  gfp=mean(gfp_t),
  myc=mean(myc_t)),
  by=c('car','donor','day','k562','t_type')], id.vars=c('car','donor','day','k562','t_type'))[,
    value_norm := value/mean(value), by=c('variable')]

myc_average_exp <- map_day_0_diff(mfi_myc_summ)[,
  list(car, value_norm_day= value/mean(value)), 
  by=c('variable','day','donor','k562','t_type')][,
    list(car, value_avg= value_norm_day/mean(value_norm_day)),
  by=c('variable','k562','t_type')][,
    car := factor(car, levels=car_order)][variable == 'myc']

pca_car_colors <- copy(car_colors)
names(pca_car_colors) <- gsub('TNR8', 'CD30', names(pca_car_colors))

supp_S3_2_gfp_myc <- ggplot(myc_average_exp[,
    k562 := factor(
      k562,
      levels=levels(factor(k562)),
      labels=c('CD19+ K562s','CD19- K562s','No K562s'))
    ][car == 'TNR8', car := 'CD30'][,
    car := factor(car, levels=gsub('TNR8','CD30',car_order))],
    aes(
      x=car, fill=car, y=value_avg, color=car, group=(interaction(car,k562)))) + 
  geom_boxplot(alpha=0.8) +
  scale_y_continuous() +
  scale_color_manual('', values=pca_car_colors) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  scale_fill_manual('', values=pca_car_colors) +
  facet_grid( ~  k562, scale='free') +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill=NA, color='grey30'),
    legend.position='none',
    strip.background = element_blank()) +
  labs(y='CAR : GFP Ratio\n(relative to mean)',
       x='CAR', title='') +
  geom_hline(yintercept=1, linetype=3, color='grey50')

```


### Assemble Supp Fig 3

```{r}

fig_s3a <- pca_factors + 
    theme(
      legend.position=c(-.6,0),
      legend.justification=c(0,0)) + 
    labs(fill='')

fig_s3b <- pca.pos.sort.size.plot + 
       coord_fixed(ratio.values) + 
       labs(fill='Domain Length (aa)') +
       theme(legend.position='bottom',
             legend.key.width = unit(1.2,'lines'),
             legend.key.height = unit(0.5, 'lines'))
      
fig_s3c <- supp_S3_2_gfp_myc

fig_s3d <- color_car_strips(car.abund.data.neg.plot + 
  theme(
    axis.title=element_text(size=unit(9,'pt')),
    panel.border=element_rect(fill=NA),
    panel.grid = element_blank(),
    legend.title = element_blank(),
    legend.box='horizontal', 
    strip.background = element_rect(colour="white", fill="white"),
    legend.justification=c(1,0),
    legend.position=c(1,0.6)))

fig_s3 <- ((fig_s3a |
  (((fig_s3b | fig_s3c) + plot_layout(widths=c(1,1.5))) / 
     fig_s3d ) + plot_layout(heights=c(1.5, 1))) + plot_layout(widths=c(1,3.5))) / 
     plot_spacer() + 
  plot_annotation(tag_levels = 'A')

ggsave(here::here('..','figs','compiled','figS3.newdraft.pdf'), fig_S3,
       height=13, width=11, useDingbats=FALSE)


```