---
title: "Arrayed Screen Differentiation Exploration"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, 
    encoding = encoding, output_dir = here::here(,'..','html'))})
---

# Load Data 
```{r}
library(ggplot2)
library(data.table)
library(scales)
library(RColorBrewer)
library(ggridges)
library(tidyverse)

output_dir <- file.path(here::here('..','figs'))

data_dir <- here::here('..','data')
load(file.path(data_dir, 'exh.sampled.Rdata'))
diff_dt <- fread(file.path(data_dir, 'exh.sampled.csv.gz'))

cd4_marker_pos_threshold <- list(
  tim3_t= 666,
  gfp_t=1250,
  zombie_t=750,
  myc_t=1400,
  lag3_t= 950, 
  pd1_t= 1400,
  cd39_t= 1100)

cd8_marker_pos_threshold <- list(
  tim3_t= 666,
  gfp_t=1500,
  zombie_t=750,
  myc_t=1400,
  lag3_t= 800, 
  pd1_t= 1400,
  cd39_t= 1250)

censor_negative_min = -1500
redo_umap_clustering = F
redo_umap_param_search = F

#only redo param search if doing clustering also:
redo_umap_clustering = redo_umap_clustering && redo_umap_clustering

# map day 0 to both plus and minus
map_day_0 <- function(df) {
  return(rbind(
    df[k562!='none'],
    df[k562=='none'][, k562 := 'cd19+'],
    df[k562=='none'][, k562 := 'cd19-']
  ))
}

exh_cast_dt <- dcast(exh_dt[!is.na(car) & !is.na(donor) & !is.na(event_id)], 
    car + donor + day + well + plate + event_id + row + col + t_type + k562 ~ 
      variable, value.var='value')
```

### CD27 expression in CD8s is maintined in BAFF-R

```{r}

#display all colors: display.brewer.all(type="seq")
blues <- brewer.pal(9,'Blues')[c(5,8)] #41bb, cd28
greens <- brewer.pal(9,'Greens')[c(5,8)] #baffr, taci
oranges <- brewer.pal(9,'Oranges')[c(4,6)] #cd40, tnr8
red <- brewer.pal(9,'PuRd')[c(7)] #klrg1
purple <- brewer.pal(9,'Purples')[c(7)] #zeta
grey <- 'grey30' #unt
car_colors <- c(
  blues[1], greens[1], greens[1], blues[2], oranges[1], 
  red, greens[2], oranges[2], grey, purple)
car_linetypes <- c(1,1,3,1,1,1,1,1,1,1)

ggplot(map_day_0(diff_dt)[
        variable=='cd27_t' & k562 == 'cd19+' & value > censor_negative_min]) + 
    geom_density(aes(x=value, color=factor(day), group=factor(day))) +
    scale_color_manual(values=brewer.pal(9,'Greens')[rev(c(4,5,6,7,8))]) + 
    facet_grid(car ~ t_type + donor)

ggplot(map_day_0(diff_dt)[
    variable=='cd27_t' & k562 == 'cd19+' & value > censor_negative_min & 
        t_type == 'cd8']) + 
    geom_boxplot(aes(x=day, y=value, fill=factor(day), group=factor(day)),
        outlier.shape = NA) +
    scale_fill_manual(values=brewer.pal(9,'Greens')[rev(c(4,5,6,7,8))]) + scale_y_continuous(limits=c(1000,3000)) +
    facet_grid(t_type + donor ~ car)

cd27_lines <- ggplot(
    map_day_0(diff_dt)[variable=='cd27_t' & k562 == 'cd19+' &
        value > censor_negative_min & t_type == 'cd8'][,
            list(val_sd= sd(value), val_mean= mean(value)),
        by=.(car, day, donor, variable, t_type)][
            car %in% c('BAFF-R','41BB','CD28','TACI','untrans') & 
                (donor == 2 | (donor == 1 & day < 33))], 
    aes(x=day, y=val_mean/1000, 
        ymin=(val_mean-val_sd)/1000, 
        ymax=(val_mean+val_sd)/1000, color=car, group=car)) + 
    geom_point() + 
    geom_line() + 
    facet_grid(~ donor, scale='free_x') + 
    scale_x_continuous(breaks=c(0,6,15,24,33)) +
    scale_y_continuous(limits=c(1.25,2.4)) +
    scale_color_manual(
        values=c(blues[1], greens[1], blues[2], greens[2], grey)) +
    theme_minimal(base_size=9) + 
    geom_vline(xintercept=-Inf) + 
    geom_hline(yintercept=-Inf) +
    labs(y='CD27 log MFI (AU)', x='Days in culture')

cd27_ridges <- ggplot(map_day_0(diff_dt)[
    variable=='cd27_t' & k562 == 'cd19+' & value > -500 & t_type == 'cd8'][
        car %in% c('BAFF-R','41BB','CD28','TACI','untrans') & 
            (donor == 2 | (donor == 1 & day < 33))]) + 
    geom_density_ridges(aes(x=value, y=factor(day), fill=factor(day), 
      height =..ndensity..)) +
    scale_fill_manual(values=brewer.pal(9,'Greens')[rev(c(4,5,6,7,8))]) +
    facet_grid(car~donor, scales='free') + 
    scale_x_continuous(limits=c(-500, 3000)) +
    scale_y_discrete(expand=c(0,0), limits= rev(levels(factor(diff_dt$day)))) +
    theme_minimal(base_size=9) + 
    geom_vline(xintercept=-Inf) + 
    geom_hline(yintercept=-Inf) +
    labs(x='CD27 log MFI (AU)', y='Days in culture')

cd27_ridges
cd27_lines

ggsave(paste0(output_dir, '/cd27_line_plot.pdf'),
       cd27_lines, width = 6, height = 3,
       units = 'in', dpi = 300)

ggsave(paste0(output_dir, '/cd27_ridge_plot.pdf'),
       cd27_ridges, width = 3, height = 10,
       units = 'in', dpi = 300)

```

## Two Dimensional Diff Plots

```{r fig.height=13, fig.width=20}

xy_hist_plot <- function(df, x_var, y_var, t_type) {
  
  if (t_type == 'cd4') {
    df_threshold = cd4_marker_pos_threshold
  } else if (t_type == 'cd8') {
    df_threshold = cd8_marker_pos_threshold
  }
  
  x_thresh = df_threshold[[x_var]]
  y_thresh = df_threshold[[y_var]]
  
  quadrants = df[, list(
    total=.N, 
    x_pos= get(x_var) > x_thresh, 
    y_pos=get(y_var) > y_thresh), by=.(car, day)]
  
  
  pct_plot = unique(quadrants[, pct := .N/total, by=.(car, day, x_pos, y_pos)])
  
  x_vals = c(-1,-1,1,1)
  y_vals = c(-1,1,-1,1)
  
  pct_plot <- pct_plot[
    data.table(
      x_pos= x_vals==1, y_pos= y_vals==1,
      x=Inf*x_vals, y=Inf*y_vals,
      hjust=1*(x_vals==1),
      vjust=1*(y_vals==1)),
    on=.(x_pos, y_pos)]
  
  ggplot(df) + 
    geom_density_2d(aes_string(x=x_var, y=y_var)) +
    facet_grid(car ~ day) +
    geom_vline(xintercept=x_thresh) +
    geom_hline(yintercept=y_thresh) +
    geom_text(data=pct_plot, aes(
      x=x, y=y, 
      label=percent(pct, 0.01),
      hjust=hjust, vjust=vjust), size=3) +
    theme_bw()
}

plot_grid(
  xy_hist_plot(
    df=map_day_0(diff_cast_dt)[k562 == 'cd19+' & t_type == 'cd4' & donor == 1],
    x_var='cd45ra_t',
    y_var='cd27_t',
    t_type='cd4'),
  xy_hist_plot(
    df=map_day_0(diff_cast_dt)[k562 == 'cd19+' & t_type == 'cd4' & donor == 2],
    x_var='cd45ra_t',
    y_var='cd27_t',
    t_type='cd4'), 
  ncol=2, labels=c('Donor 1','Donor 2'))

plot_grid(
  xy_hist_plot(
    df=map_day_0(diff_cast_dt)[k562 == 'cd19+' & t_type == 'cd4' & donor == 1],
    x_var='cd45ra_t',
    y_var='ccr7_t',
    t_type='cd4'),
  xy_hist_plot(
    df=map_day_0(diff_cast_dt)[k562 == 'cd19+' & t_type == 'cd4' & donor == 2],
    x_var='cd45ra_t',
    y_var='ccr7_t',
    t_type='cd4'), 
  ncol=2, labels=c('Donor 1','Donor 2'))
```

## PCA Diff Plots

```{r}
df_subset <- map_day_0(exh_cast_dt)[k562 == 'cd19+' & t_type == 'cd8']

pca <- prcomp(
  df_subset[, grep('\\d+_t',names(df_subset), value=T), with=F])

df_subset <- cbind(df_subset, pca$x)

ggplot(df_subset) + 
    geom_density_2d(
        aes(
            x=PC1, y=PC2, 
            color=factor(day), 
            group=factor(day))) + 
    facet_grid(car ~ donor + day)




```

# Exh UMAP


For now, skip straight to high dimensional plotting.

## UMAP Functions
```{r}

library(ggplot2)
library(data.table)
library(scales)
library(RColorBrewer)
library(reticulate)
library(uwot)
library(gridExtra)
#library(spatstat)
library(ggdendro)
library(cowplot)
library(tidyverse)

#grab channel map from cd4 data (cd4/8 are the same)
channel_map <- exh_opt_cd4$channel_map

# for now use all variables in the channel map except cd4/8
umap_vars <- c(paste0(names(channel_map),'_t'))
umap_vars <- umap_vars[umap_vars != 'cd_t']

sample_for_umap <- function(df, sample_size) {
  umap_dt <- diff_dt[!is.na(event_id), 
    .SD[event_id %in% sample(unique(event_id), 
      min(sample_size, length(unique(event_id))))],
    by=c('well', 'plate', 'day')]
}

cast_for_umap <- function(df) {
  #cast it so variables are columns and
  #subset sample umap data on variables
  umap_cast_dt <- dcast(df, 
    event_id + well + plate + day + t_type ~ variable, 
    value.var='value')
  return(umap_cast_dt)
}

scale_for_umap <- function(df, umap_vars, censor_negative_min) {
  umap_dt_in <- df[, ..umap_vars]
  
  # for points that are very negative, trim the to below the cutoff
  umap_dt_in[umap_dt_in < censor_negative_min] <- censor_negative_min
  
  # scale each input column
  umap_dt_in[ , 
    (names(umap_dt_in)) := lapply(.SD, scale), .SDcols = names(umap_dt_in)]
  
  return(umap_dt_in)
}
  
# check scales:
# ggplot(melt(cbind(umap_dt_in, id=1:nrow(umap_dt_in)), id.var='id'), 
#        aes(x=value)) +
#     geom_density() +
#     facet_grid(variable~.) +
#     scale_fill_distiller(palette='RdYlBu') +
#     theme_bw()

```
## Choose UMAP Params

First, find parameters with a small sample.
```{r eval=redo_umap_param_search}
# use umap canned functions
umap_dt <- sample_for_umap(exh_dt, 20)
umap_cast_dt <- cast_for_umap(umap_dt)
umap_dt_in <- scale_for_umap(umap_cast_dt, umap_vars, censor_negative_min)

### Run across parameter list

#umap parameter list
min_dist_set <- c(0.0001, 0.005, 0.1)
n_neighbor_set <- c(3,6,15,30)
umap_params <- data.table(expand.grid(min_dist_set, n_neighbor_set))

# run umap via uwot library
umap_out <- umap_params[, {
  print(paste0('Running: ',.BY)); 
  as.data.table(umap(umap_dt_in, min_dist=as.numeric(.BY[1]),
    n_neighbors=as.numeric(.BY[2]), verbose=F, n_threads=8, n_trees=50))
  }, by=names(umap_params)]

# rename the columns
names(umap_out) <- c('min_dist','n_neighbor','umap_1','umap_2')

# add the umap output to the input dt
umap_fcs_dt <- cbind(umap_cast_dt[, 1:length(names(umap_cast_dt))], umap_out)
umap_fcs_dt[, id := 1:.N]

umap_fcs_dt <- unique(umap_dt[, 
  .(donor, car, k562, t_type, day, well, event_id)])[
    umap_fcs_dt, on=.(t_type,well,day,event_id)]



```

```{r eval=redo_umap_param_search, fig.width=20, fig.height=10}
color_points <- ggplot(umap_fcs_dt[car %in% c('CD28','41BB','KLRG1','BAFF-R','Zeta')], 
    aes(x=umap_1, y=umap_2, color=interaction(car, t_type))) +
  geom_point(size=0.1, alpha=0.1) + 
  guides(colour = guide_legend(override.aes = list(alpha=1, size=3))) +
  facet_grid(min_dist~n_neighbor) +
  theme_bw()

color_density <- ggplot(umap_fcs_dt, aes(x=umap_1, y=umap_2)) +
  geom_hex(bins = 70) +
  scale_fill_continuous(
    type = "viridis", limits=c(0,30), oob=scales::squish) +
  facet_grid(min_dist~n_neighbor) +
  theme_bw()

grid.arrange(color_points, color_density, ncol=2)
```

Choosing neighbors == 15 and min_dist == 1e-4. 

## Checking UMAP Param space

Scaling this to 200 events per well and checking CAR/condition separation.

```{r, eval=redo_umap_clustering}
library(ggplot2)
library(data.table)
library(scales)
library(RColorBrewer)
library(reticulate)
library(uwot)
library(gridExtra)
library(spatstat)
library(ggdendro)
library(cowplot)
library(tidyverse)
use_condaenv(condaenv = 'r-reticulate', required = TRUE)

run_umap <- function(dt, chosen_dist, chosen_n_neighbor, sample_n, umap_vars) {
  
  umap_dt <- dt[!is.na(event_id), 
    .SD[event_id %in% sample(unique(event_id), 
      min(sample_n, length(unique(event_id))))],
    by=c('well', 'plate', 'day')]
  
  #grab channel map from cd4 data (cd4/8 are the same)
  #channel_map <- diff_opt_cd4$channel_map
  
  # for now use all variables in the channel map
  #umap_vars <- c(paste0(names(channel_map),'_t'))
  
  #cast it so variables are columns and
  #subset sample umap data on variables
  umap_cast_dt <- data.table(dcast(umap_dt, 
    event_id + well + plate + day + t_type ~ variable, 
    value.var='value'))
  
  umap_dt_in <- umap_cast_dt[, ..umap_vars]
  
  # scale each input column
  umap_dt_in[ , 
    (names(umap_dt_in)) := lapply(.SD, scale), .SDcols = names(umap_dt_in)]
  
  # run umap via uwot library
  umap_out <- umap(umap_dt_in, min_dist=chosen_dist,
      n_neighbors=chosen_n_neighbor, verbose=T, n_threads=8, n_trees=50,
      ret_nn=T)
  
  # nearest neighbors in original space
  nearest_neighbors <- umap_out$nn$euclidean$idx
  neighbor_dist <- umap_out$nn$euclidean$dist
  edge_weights <- scales::rescale(-neighbor_dist, to=c(0,1))
  edge_weights <- edge_weights - min(edge_weights)
  umap_out <- data.table(umap_out$embedding)
  
  #nearest neighbors in embedding
  # umap_nn <- cbind(1:nrow(umap_fcs_dt), 
  #     nnwhich(umap_fcs_dt[, list(umap_1, umap_2)], k=c(1:3)))
  # umap_nd <- cbind(rep(0, nrow(umap_fcs_dt)), 
  #     nndist(umap_fcs_dt[, list(umap_1, umap_2)], k=c(1:3)))
  # umap_ew <- scales::rescale(-umap_nd, to=c(0,1))
  # umap_ew <- umap_ew - min(umap_ew)
      
  # rename the columns
  names(umap_out) <- c('umap_1','umap_2')
  
  # add the umap output to the input dt
  umap_fcs_dt <- cbind(umap_cast_dt[, 1:length(names(umap_cast_dt))], umap_out)
  umap_fcs_dt[, id := 1:.N]
  
  # add back the annotations
  umap_fcs_dt <- unique(umap_dt[, 
    .(donor, car, k562, t_type, day, well, event_id)])[
      umap_fcs_dt, on=.(t_type,well,day,event_id)]
  
  print(edge_weights)
  source_python(here::here('py/leiden_clust.py'))
  clust_membership <- leiden_clust(nearest_neighbors,
    edge_weights=scales::rescale(-neighbor_dist, to=c(0,1)))
  umap_fcs_dt[, cluster := factor(clust_membership+1)]
  
  return(umap_fcs_dt)
}

umap_vars <- c("tim3_t","lag3_t","cd39_t","pd1_t" )
umap_fcs_dt <- run_umap(exh_dt, 0.005, 15, 2500, umap_vars)

# save the clustering data
fwrite(umap_fcs_dt, 
  compress='gzip', 
  file=file.path(here::here('..','data','exh.sampled.umap.csv.gz')))

# sync output back to s3
system('aws s3 sync \\ 
  --exclude "*" \\
  --include "*.Rdata" \\
  --include "*.csv*" \\
  ~/local-tcsl/data s3://roybal-tcsl//lenti_screen_compiled_data/data')
```
