---
title: "Differentiation Plots for Arrayed Screens"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here('..','html'))})
---

# Load Data 
```{r}
library(ggplot2)
library(data.table)
library(scales)
library(RColorBrewer)
library(reticulate)
library(uwot)
library(gridExtra)
library(spatstat)
library(ggdendro)
library(cowplot)
library(tidyverse)
library(grid)
library(gridExtra) # add to AWS
library(cowplot) # add to AWS
library(slingshot) # add to AWS
library(viridis) # add to AWS
library(reshape2)
library(RColorBrewer)
library(ggrepel)
#library(monocle3)
use_condaenv(condaenv = 'r-reticulate', required = TRUE)

data_dir <- here::here('..','data')
load(file.path(data_dir, 'diff.sampled.Rdata'))
diff_dt <- fread(file.path(data_dir, 'diff.sampled.csv.gz'))

censor_negative_min = -1500
redo_umap_clustering = F
redo_umap_param_search = F

#only redo param search if doing clustering also:
redo_umap_clustering = redo_umap_clustering && redo_umap_clustering

# map day 0 to both plus and minus
map_day_0 <- function(df) {
  return(rbind(
    df[k562!='none'],
    df[k562=='none'][, k562 := 'cd19+'],
    df[k562=='none'][, k562 := 'cd19-']
  ))
}
```

# UMAP

For now, skip straight to high dimensional plotting.

## UMAP Functions
```{r}

#grab channel map from cd4 data (cd4/8 are the same)
channel_map <- diff_opt_cd4$channel_map

# for now use all variables in the channel map except cd4/8
# removing zombie, cd_t, myc_t, gfp_t
umap_vars <- c(paste0(names(channel_map),'_t'))
umap_vars <- umap_vars[umap_vars != c('cd_t', 'zombie_t', 'myc_t', 'gfp_t')]

sample_for_umap <- function(df, sample_size) {
  umap_dt <- diff_dt[!is.na(event_id), 
    .SD[event_id %in% sample(unique(event_id), 
      min(sample_size, length(unique(event_id))))],
    by=c('well', 'plate', 'day')]
}

cast_for_umap <- function(df) {
  #cast it so variables are columns and
  #subset sample umap data on variables
  umap_cast_dt <- dcast(df, 
    event_id + well + plate + day + t_type ~ variable, 
    value.var='value')
  return(umap_cast_dt)
}

scale_for_umap <- function(df, umap_vars, censor_negative_min) {
  umap_dt_in <- df[, ..umap_vars]
  
  # for points that are very negative, trim the to below the cutoff
  umap_dt_in[umap_dt_in < censor_negative_min] <- censor_negative_min
  
  # scale each input column
  umap_dt_in[ , 
    (names(umap_dt_in)) := lapply(.SD, scale), .SDcols = names(umap_dt_in)]
  
  return(umap_dt_in)
}
  
# check scales:
# ggplot(melt(cbind(umap_dt_in, id=1:nrow(umap_dt_in)), id.var='id'), 
#        aes(x=value)) +
#     geom_density() +
#     facet_grid(variable~.) +
#     scale_fill_distiller(palette='RdYlBu') +
#     theme_bw()
```
## Choose UMAP Params

First, find parameters with a small sample.
```{r eval=redo_umap_param_search}
# use umap canned functions
umap_dt <- sample_for_umap(diff_dt, 20)
umap_cast_dt <- cast_for_umap(umap_dt)
umap_dt_in <- scale_for_umap(umap_cast_dt, umap_vars, censor_negative_min)

### Run across parameter list

#umap parameter list 
min_dist_set <- c(0.0001, 0.005, 0.1)
n_neighbor_set <- c(3,6,15,30)
learning_rate_set <- c(0.1, 0.25, 0.5, 1) 


umap_params <- data.table(expand.grid(min_dist_set, n_neighbor_set, learning_rate_set))

# run umap via uwot library
umap_out <- umap_params[, {
  print(paste0('Running: ',.BY)); 
  as.data.table(umap(umap_dt_in, min_dist=as.numeric(.BY[1]),
    n_neighbors=as.numeric(.BY[2]), learning_rate=as.numeric(.BY[3]), verbose=F, n_threads=8, n_trees=100))
  }, by=names(umap_params)]

# rename the columns
names(umap_out) <- c('min_dist','n_neighbor', 'learning_rate', 'umap_1','umap_2')

# add the umap output to the input dt
umap_fcs_dt <- cbind(umap_cast_dt[, 1:length(names(umap_cast_dt))], umap_out)
umap_fcs_dt[, id := 1:.N]

umap_fcs_dt <- unique(umap_dt[, 
  .(donor, car, k562, t_type, day, well, event_id)])[
    umap_fcs_dt, on=.(t_type,well,day,event_id)]
```

```{r eval=redo_umap_param_search, fig.width=30, fig.height=15}
color_points <- ggplot(umap_fcs_dt[car %in% c('CD28','41BB','KLRG1','BAFF-R','Zeta')], 
    aes(x=umap_1, y=umap_2, color=interaction(car, t_type))) +
  geom_point(size=0.1, alpha=0.1) + 
  guides(colour = guide_legend(override.aes = list(alpha=1, size=3))) +
  facet_grid(min_dist~n_neighbor+learning_rate) +
  theme_bw()

color_density <- ggplot(umap_fcs_dt, aes(x=umap_1, y=umap_2)) +
  geom_hex(bins = 70) +
  scale_fill_continuous(
    type = "viridis", limits=c(0,30), oob=scales::squish) +
  facet_grid(min_dist~n_neighbor+learning_rate) +
  theme_bw()

grid.arrange(color_points, color_density, ncol=2)
```

Choosing neighbors == 30 and min_dist == 1e-4. 

## Checking UMAP Param space

Scaling this to 200 events per well and checking CAR/condition separation.

```{r}
chosen_dist=0.0001
chosen_n_neighbor=100 # default is 15
chosen_learning_rate=0.1

umap_dt <- diff_dt[!is.na(event_id), 
  .SD[event_id %in% sample(unique(event_id), 
    min(1000, length(unique(event_id))))], #modify minimum parameter for points per well
  by=c('well', 'plate', 'day')]

#grab channel map from cd4 data (cd4/8 are the same)
channel_map <- diff_opt_cd4$channel_map

# for now use all variables in the channel map
umap_vars <- c(paste0(names(channel_map),'_t'))

# remove cd_t, zombie_t, myc_t, gfp_t from umap_vars
umap_vars <- umap_vars[!umap_vars %in% c('cd_t', 'zombie_t', 'myc_t', 'gfp_t')]
umap_vars # check umap vars

# filter by cd4/8, cd19+ and day 0 only
umap_cd4_dt <- umap_dt %>% filter(t_type=='cd4', k562 %in% c('none', 'cd19+'))
umap_cd8_dt <- umap_dt %>% filter(t_type=='cd8', k562 %in% c('none', 'cd19+'))
```
### CD4 UMAP
```{r, eval=redo_umap_clustering}
umap_cd4_cast_dt <- dcast(umap_cd4_dt, 
  event_id + well + plate + day + t_type ~ variable, 
  value.var='value')

umap_cd4_dt_in <- umap_cd4_cast_dt[, umap_vars]

# scale each input column
umap_cd4_dt_in <- scale(umap_cd4_dt_in)

# run umap via uwot library (added learning rate and n_sgd_threads=1)
set.seed(123)
umap_cd4_out <- umap(umap_cd4_dt_in, min_dist=chosen_dist,
    n_neighbors=chosen_n_neighbor, learning_rate=chosen_learning_rate, verbose=T, n_threads=8, n_trees=100, n_sgd_threads=1,
    ret_nn=T)

# nearest neighbors in original space
nearest_neighbors_cd4 <- umap_cd4_out$nn$euclidean$idx
neighbor_dist_cd4 <- umap_cd4_out$nn$euclidean$dist
edge_weights_cd4 <- scales::rescale(-neighbor_dist_cd4, to=c(0,1))
edge_weights_cd4 <- edge_weights_cd4 - min(edge_weights_cd4)
umap_cd4_out <- data.table(umap_cd4_out$embedding)

#nearest neighbors in embedding
umap_nn <- cbind(1:nrow(umap_fcs_dt), 
    nnwhich(umap_fcs_dt[, list(umap_1, umap_2)], k=c(1:3)))
umap_nd <- cbind(rep(0, nrow(umap_fcs_dt)), 
    nndist(umap_fcs_dt[, list(umap_1, umap_2)], k=c(1:3)))
umap_ew <- scales::rescale(-umap_nd, to=c(0,1))
umap_ew <- umap_ew - min(umap_ew)

# rename the columns
names(umap_cd4_out) <- c('umap_1','umap_2')

# add the umap output to the input dt
umap_cd4_fcs_dt <- cbind(umap_cd4_cast_dt[, 1:length(names(umap_cd4_cast_dt))], umap_cd4_out)
umap_cd4_fcs_dt <- tibble::rowid_to_column(umap_cd4_fcs_dt, "id")

# add back the annotations
umap_cd4_fcs_dt <- umap_cd4_fcs_dt %>% left_join(distinct(umap_cd4_dt, donor, car, k562, t_type, day, well, event_id), by=c('t_type', 'day', 'well', 'event_id'))
```

### CD8 UMAP
```{r, eval=redo_umap_clustering}
umap_cd8_cast_dt <- dcast(umap_cd8_dt, 
  event_id + well + plate + day + t_type ~ variable, 
  value.var='value')

umap_cd8_dt_in <- umap_cd8_cast_dt[, umap_vars]

# scale each input column
umap_cd8_dt_in <- scale(umap_cd8_dt_in)

# run umap via uwot library (added learning rate and n_sgd_threads=1)
set.seed(123)
umap_cd8_out <- umap(umap_cd8_dt_in, min_dist=chosen_dist,
    n_neighbors=chosen_n_neighbor, learning_rate=chosen_learning_rate, verbose=T, n_threads=8, n_trees=100, n_sgd_threads=1,
    ret_nn=T)

# nearest neighbors in original space
nearest_neighbors_cd8 <- umap_cd8_out$nn$euclidean$idx
neighbor_dist_cd8 <- umap_cd8_out$nn$euclidean$dist
edge_weights_cd8 <- scales::rescale(-neighbor_dist_cd8, to=c(0,1))
edge_weights_cd8 <- edge_weights_cd8 - min(edge_weights_cd8)
umap_cd8_out <- data.table(umap_cd8_out$embedding)

#nearest neighbors in embedding
umap_nn <- cbind(1:nrow(umap_fcs_dt), 
    nnwhich(umap_fcs_dt[, list(umap_1, umap_2)], k=c(1:3)))
umap_nd <- cbind(rep(0, nrow(umap_fcs_dt)), 
    nndist(umap_fcs_dt[, list(umap_1, umap_2)], k=c(1:3)))
umap_ew <- scales::rescale(-umap_nd, to=c(0,1))
umap_ew <- umap_ew - min(umap_ew)

# rename the columns
names(umap_cd8_out) <- c('umap_1','umap_2')

# add the umap output to the input dt
umap_cd8_fcs_dt <- cbind(umap_cd8_cast_dt[, 1:length(names(umap_cd8_cast_dt))], umap_cd8_out)
umap_cd8_fcs_dt <- tibble::rowid_to_column(umap_cd8_fcs_dt, "id")

# add back the annotations
umap_cd8_fcs_dt <- umap_cd8_fcs_dt %>% left_join(distinct(umap_cd8_dt, donor, car, k562, t_type, day, well, event_id), by=c('t_type', 'day', 'well', 'event_id'))

```

## Clustering via leiden/reticulate

```{python, eval=redo_umap_clustering}
import leidenalg
import igraph as ig
import pandas as pd
import numpy as np
import math

# first, for CD4
edges_cd4 = []
n_neighbors_cd4 = r.nearest_neighbors_cd4.shape[1]
for i, nn_row in enumerate(r.nearest_neighbors_cd4):
    edge_weights_cd4 = r.edge_weights_cd4[i]
    for j in range(1, n_neighbors_cd4):
        edges_cd4.append((int(nn_row[0]), int(nn_row[j]), edge_weights_cd4[j]))
      
knn_graph_cd4 = ig.Graph.TupleList(edges_cd4, directed=True, weights=True)

part_cd4 = leidenalg.find_partition(knn_graph_cd4, 
    leidenalg.ModularityVertexPartition, weights='weight')
#part = leidenalg.find_partition(knn_graph, 
#    leidenalg.CPMVertexPartition, weights='weight', 
#    resolution_parameter=0.05)

cluster_membership_cd4 = [x for _,x in sorted(zip(knn_graph_cd4.vs['name'],part_cd4.membership))]

# again for CD8
edges_cd8 = []
n_neighbors_cd8 = r.nearest_neighbors_cd8.shape[1]
for i, nn_row in enumerate(r.nearest_neighbors_cd8):
    edge_weights_cd8 = r.edge_weights_cd8[i]
    for j in range(1, n_neighbors_cd8):
        edges_cd8.append((int(nn_row[0]), int(nn_row[j]), edge_weights_cd8[j]))
      
knn_graph_cd8 = ig.Graph.TupleList(edges_cd8, directed=True, weights=True)

part_cd8 = leidenalg.find_partition(knn_graph_cd8, 
    leidenalg.ModularityVertexPartition, weights='weight')
#part = leidenalg.find_partition(knn_graph, 
#    leidenalg.CPMVertexPartition, weights='weight', 
#    resolution_parameter=0.05)

cluster_membership_cd8 = [x for _,x in sorted(zip(knn_graph_cd8.vs['name'],part_cd8.membership))]

```

```{r, eval=redo_umap_clustering, fig.width=15, fig.height=10}
# convert into data tables
umap_cd4_fcs_dt <- as.data.table(umap_cd4_fcs_dt)
umap_cd8_fcs_dt <- as.data.table(umap_cd8_fcs_dt)

#add new column to umap 
umap_cd4_fcs_dt[, cluster := factor(py$cluster_membership_cd4+1)]
umap_cd8_fcs_dt[, cluster := factor(py$cluster_membership_cd8+1)]
```

```{r, eval=redo_umap_clustering}

# save the clustering data
fwrite(umap_cd4_fcs_dt, 
  compress='gzip', 
  file=file.path(here::here('..','data','diff.sampled.umap_cd4.csv.gz')))

fwrite(umap_cd8_fcs_dt, 
  compress='gzip', 
  file=file.path(here::here('..','data','diff.sampled.umap_cd8.csv.gz')))

# sync output back to s3
system('aws s3 sync \\
  --exclude "*" \\
  --include "*.Rdata" \\
  --include "*.csv*" \\
  ~/local-tcsl/data s3://roybal-tcsl//lenti_screen_compiled_data/data')
```

# PCA (in the works)
```{r}
# pca_cd8 <- prcomp(umap_cd8_dt_in) #data already scaled from above umap preprocessing
#summary(pca_cd8) #PC1,PC2 explain ~55% of variance
#names(pca_cd8)

#plot(pca_cd8$x[,1:2], main="first PCA")

#pca_cd8$x %>% 
  #as.data.frame %>%
  #ggplot(aes(x=PC1,y=PC2)) + geom_point(size=1) +
  #theme_bw(base_size=20)

# pca_cd8_df <- as.data.frame(pca_cd8$x), 
# pca_cd8_df %>% head()

# umap_cd4_dt_in


```



## Cluster Analysis

One issue with these clusters is that they are not contiguous in the UMAP embedding:

### CD8 UMAP Plot
```{r, fig.width=15, fig.height=12}
umap_cd8_fcs_dt <- fread(
  file.path(
    here::here('..','data','diff.sampled.umap_cd8.csv.gz')))
umap_cd8_fcs_dt[, cluster := factor(cluster)]

umap_cd8_cluster_dt <- umap_cd8_fcs_dt[, list(
  mean_umap_1= mean(umap_1), 
  mean_umap_2= mean(umap_2), 
  size= .N), by=cluster]

# whole plot
ggplot() + 
  geom_point(data=umap_cd8_fcs_dt, 
    aes(x=umap_1, y=umap_2, color=cluster), size=0.2, alpha=0.2) +
  geom_label(data=umap_cd8_cluster_dt, aes(
    x=mean_umap_1, y=mean_umap_2,
    label=paste(cluster,size,sep='\n'), 
    color=cluster), alpha=0.3) +
  theme_minimal()

# individual plots
ggplot() + 
  geom_point(data=umap_cd8_fcs_dt, 
    aes(x=umap_1, y=umap_2, color=cluster), size=0.2, alpha=0.2) +
  geom_label(data=umap_cd8_cluster_dt, aes(
    x=mean_umap_1, y=mean_umap_2,
    label=paste(cluster,size,sep='\n'), 
    color=cluster), alpha=0.3) +
  facet_wrap(~cluster) +
  theme_bw()
```
### CD4 UMAP Plot
```{r, fig.width=15, fig.height=12}
umap_cd4_fcs_dt <- fread(
  file.path(
    here::here('..','data','diff.sampled.umap_cd4.csv.gz')))
umap_cd4_fcs_dt[, cluster := factor(cluster)]

umap_cd4_cluster_dt <- umap_cd4_fcs_dt[, list(
  mean_umap_1= mean(umap_1), 
  mean_umap_2= mean(umap_2), 
  size= .N), by=cluster]

# whole plot
ggplot() + 
  geom_point(data=umap_cd4_fcs_dt, 
    aes(x=umap_1, y=umap_2, color=cluster), size=0.2, alpha=0.2) +
  geom_label(data=umap_cd4_cluster_dt, aes(
    x=mean_umap_1, y=mean_umap_2,
    label=paste(cluster,size,sep='\n'), 
    color=cluster), alpha=0.3) +
  theme_minimal()

# individual plots
ggplot() + 
  geom_point(data=umap_cd4_fcs_dt, 
    aes(x=umap_1, y=umap_2, color=cluster), size=0.2, alpha=0.2) +
  geom_label(data=umap_cd4_cluster_dt, aes(
    x=mean_umap_1, y=mean_umap_2,
    label=paste(cluster,size,sep='\n'), 
    color=cluster), alpha=0.3) +
  facet_wrap(~cluster) +
  theme_bw()
```

```{r, fig.width=30, fig.height=20}
color_plots <- function(umap_df) {
  cd4_colors = brewer.pal('Greens', n=9)[c(2,4,6,8,9)]
  cd8_colors = brewer.pal('Oranges', n=9)[c(2,4,6,8,9)]
  color_time <- ggplot(umap_df[][, cd19 := (k562=='cd19+')], 
      aes(x=umap_1, y=umap_2, color=interaction(day, t_type))) +
    geom_point(size=0.1, alpha=0.5) +
    facet_grid(car~cd19) +
    scale_color_manual(values=c(cd4_colors, cd8_colors)) +
    guides(colour = guide_legend(override.aes = list(alpha=1, size=3)))
  
  color_density <- ggplot(umap_df, aes(x=umap_1, y=umap_2)) +
    geom_hex(bins = 70) +
    facet_grid(car~cd19) +
    scale_fill_continuous(
      type = "viridis", limits=c(0,30), oob=scales::squish) +
    theme_bw()
  
  color_markers <- ggplot(
      melt(umap_df, measure.vars=umap_vars)[, 
        value.scaled := scale(value), by='variable'][,
          cd19 := (k562=='cd19+')], 
      aes(x=umap_1, y=umap_2, z=value.scaled)) +
    stat_summary_hex(bins = 70) +
    facet_grid(variable~cd19) +
    scale_fill_distiller(palette='RdYlBu', limits=c(-3, 3), oob=scales::squish) +
    theme_bw()
  
  grid.arrange(color_time, color_density, color_markers, ncol=3)
}

#cd8
color_plots(umap_cd8_fcs_dt)

#cd4
color_plots(umap_cd4_fcs_dt)
```

```{r, fig.width=15, fig.height=8}
dendrogram <- function(umap_fcs_dt, cell_type) {
  cluster_pct_vars <- c(paste0(
    names(channel_map),'_t'), 
    'FSC-A','SSC-A','donor','cd19')
  
  max_cluster <- max(as.numeric(umap_fcs_dt$cluster), na.rm=T)
  
  umap_var_melt_dt <-melt(
    umap_fcs_dt[!is.na(donor)], measure.vars= c(umap_vars,'donor', 'cd19', 'myc_t', 'gfp_t')) #add back in previously removed vars
  
  # Use z scale to plot params
  umap_var_melt_dt[, value_scaled := scale(value), by=variable]
  param_cluster_pct_dt <- umap_var_melt_dt[
    !is.na(cluster) & variable %in% cluster_pct_vars,
    list(
      clust_mean = mean(value_scaled),
      clust_sd = sd(value_scaled)),
        by=c('variable','cluster')]
  
  # Make dendrogram
  cluster_mean_cast <- dcast(
    param_cluster_pct_dt, 
    variable ~ cluster, 
    value.var='clust_mean')
  
  cluster_dendro_m <- as.matrix(cluster_mean_cast[, -c(1)])
  rownames(cluster_dendro_m) <- unlist(cluster_mean_cast[,1])
  dendro_clusters <- as.dendrogram(hclust(d = dist(x = t(cluster_dendro_m))))
  dendro_vars <- as.dendrogram(hclust(d = dist(x = cluster_dendro_m)))
  
  # Create dendrogram plot
  dendro_vars_plot <- ggdendrogram(data = dendro_vars, rotate = TRUE) + 
    theme(axis.text.y = element_text(size = 6))
  dendro_clusters_plot <- ggdendrogram(data = dendro_clusters, rotate = TRUE) + 
    theme(axis.text.y = element_text(size = 6))
  
  # column order
  cluster_order <- order.dendrogram(dendro_clusters)
  cluster_names <- colnames(cluster_dendro_m[, cluster_order])
  param_cluster_pct_dt[, cluster:= factor(cluster, levels = cluster_names)]
  
  # row order
  var_order <- order.dendrogram(dendro_vars)
  var_names <- row.names(cluster_dendro_m[var_order, ])
  param_cluster_pct_dt$variable <- factor(
      param_cluster_pct_dt$variable,
      levels = var_names)
  
   #annotate clusters as per variable values (removed gfp, myc, cd, donor)
  '%notin%' <- Negate('%in%')
  annotations <- param_cluster_pct_dt %>%
  mutate(variable=gsub("\\_t$", "", variable)) %>%
  filter(variable %notin% c('gfp','myc','cd','donor')) %>%
  mutate(cluster=as.character(cluster)) %>%
  group_by(cluster) %>% 
  mutate(rk = rank(-abs(clust_mean))) %>%
  filter(rk <= 3) %>% 
  arrange(cluster, rk) %>%
  mutate(variable = ifelse(clust_mean > 0, paste(variable,'+',sep = ""), paste(variable,'-',sep = ""))) %>%
  summarise(variable = paste(variable, collapse=", ")) %>% 
  data.table::transpose()
  colnames(annotations) <- as.character(unlist(annotations[1,]))
  annotations = annotations[-1, ]
  annotations <- annotations[cluster_names]
  row.names(annotations) <- NULL 
  
  # heatmap
  var_heatmap <- ggplot(param_cluster_pct_dt) + 
      geom_tile(aes(x=cluster, y=variable, fill=clust_mean), color='black') + geom_text(aes(x=cluster, y=variable, label=round(clust_mean, 2)), size=3, color='black') +
      scale_fill_distiller(palette='PiYG', limits=c(-3,3), oob=scales::squish) +
      theme_minimal() # + theme(axis.text.x = element_text(angle=90, hjust=1))

  # col dendro
  dendro_data_col <- dendro_data(dendro_clusters, type = "rectangle")
  dendro_col <- axis_canvas(var_heatmap, axis = "x") + 
      geom_segment(data = segment(dendro_data_col), 
          aes(x = x, y = y, xend = xend, yend = yend))
  plot_dendroheat <- insert_xaxis_grob(var_heatmap, 
              dendro_col, grid::unit(0.2, "null"), position = "top")
  dendro_plot <- ggdraw(plot_dendroheat) + draw_label(label=cell_type, x = 0.95, y = 0.95, hjust = 0.95, vjust = 0.95, size=40) 
  
  # annotation table
  tt <- ttheme_default(core=list(fg_params = list(fontsize=10), colhead=list(fg_params = list(fontsize=10, parse=TRUE))))
  annotations[1,] = str_wrap(annotations[1,], 0)
  tbl <- tableGrob(annotations, rows=NULL, theme=tt)
  # tbl <- strwrap(tbl, width = 10, simplify = FALSE)
  tbl$widths <- unit(rep(1/ncol(tbl), ncol(tbl)), "npc")
  
  # print plot and table
  return(list(grid.arrange(dendro_plot, tbl, nrow = 2, as.table = TRUE), cluster_names))
} 

#cd8
cd8_dendro <- dendrogram(umap_cd8_fcs_dt, 'cd8')
cd8_colnames <- cd8_dendro[[2]]

#cd4
cd4_dendro <- dendrogram(umap_cd4_fcs_dt, 'cd4')
cd4_colnames <- cd4_dendro[[2]]
```

### Percent of cells in day, by cluster

```{r, echo=F, fig.width=15, fig.height=8}
day_cluster <- function(umap_fcs_dt, cell_type, cluster_names) {
  day_cluster_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(day,cluster, t_type, k562))][,
      list(day, pct=N/sum(N)), by=c('k562','cluster','t_type')]
  
  day_cluster_pct_dt[, cluster := factor(cluster, levels = cluster_names)]
  
  ggplot(day_cluster_pct_dt[k562=='cd19+'], aes(
      x=cluster, 
      y=factor(day, levels=c(0,6,15,24,33)), 
      fill=pct)) + 
    geom_tile( 
        color='black') +
    geom_text(aes(label=round(pct*100)), size=3, color='white') +
    facet_grid(k562+t_type~., scales='free') +
    scale_fill_viridis_b(direction=1, limits=c(0, 1)) +
    labs(y='Day', subtitle='(cols sum to 100%)') +
    theme_minimal() + ggtitle(paste('Percent of ',as.character(cell_type),' cells in each cluster by day'))
}

#cd8
day_cluster(umap_cd8_fcs_dt, 'CD8', cd8_colnames)

#cd4
day_cluster(umap_cd4_fcs_dt, 'CD4', cd4_colnames)
```

### Percent of cells in cluster, by day

```{r, echo=F, fig.width=15, fig.height=8}
day_cluster_total <- function(umap_fcs_dt, cell_type, cluster_names) {
  day_cluster_total_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(day,cluster, t_type, k562))][,
      list(cluster, pct=N/sum(N)), by=c('k562','day','t_type')]
  
  day_cluster_total_pct_dt[, cluster := factor(cluster, levels = cluster_names)]
  
  ggplot(day_cluster_total_pct_dt[k562=='cd19+'], aes(
      x=cluster, 
      y=factor(day, levels=c(0,6,15,24,33)), 
      fill=pct)) + 
    geom_tile(color='black') +
    geom_text(aes(label=round(pct*100)), size=3, color='white') +
    facet_grid(k562+t_type~., scales='free') +
    scale_fill_viridis_b('',direction=1, limits=c(0, 0.6)) +
    labs(title='Percent of cells in each day by cluster', y='Day',
      subtitle='(rows sum to 100%)') +
    theme_minimal() + ggtitle(paste('Percent of ',as.character(cell_type),' cells in each day by cluster'))
}

#cd8
#day_cluster_total(umap_cd8_fcs_dt, 'CD8', cd8_colnames)

#cd4
#day_cluster_total(umap_cd4_fcs_dt, 'CD4', cd4_colnames)

```

### CAR differences per cluster
#### summed across cols
```{r, echo=F, fig.width=15, fig.height=8}

car_per_day_per_cluster <- function(umap_fcs_dt, cell_type, cluster_names) {
  car_cluster_day_indiv_k562_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(car,cluster,k562,t_type, day))][,
      list(car, pct=N/sum(N)), by=c('cluster','k562','t_type','day')][,
        pct_delta := scale(pct), by=c('cluster','k562','t_type', 'day')]
  
  car_cluster_day_indiv_k562_pct_dt[,
    cluster := factor(cluster, levels = cluster_names)]
  
  car_cluster_day_indiv_k562_pct_dt$day <- factor(car_cluster_day_indiv_k562_pct_dt$day, levels=c('0','6','15','24','33'))
  
  min_col <- brewer.pal(name='BrBG', n=11)[2]
  
  ggplot(car_cluster_day_indiv_k562_pct_dt[k562=='cd19+']) + 
      geom_tile(aes(x=cluster, y=car, fill=pct_delta), color='black') +
      facet_grid(day~k562, scales='free') +
      scale_fill_distiller('',
        palette='BrBG', 
        direction=1,
        #limits=c(-2.8, 2.8), 
        na.value=min_col) +
      labs(title='CAR enrichment\nper day per cluster', y='CAR',
        subtitle='CAR enrichment in cluster\n(z-score) (summed across cols)') +
      theme_minimal() + ggtitle(paste(as.character(cell_type),' CAR enrichment\nper day per cluster'))
}

#cd8
car_per_day_per_cluster(umap_cd8_fcs_dt, "CD8", cd8_colnames)

#cd4
car_per_day_per_cluster(umap_cd4_fcs_dt, "CD4", cd4_colnames)
```

```{r, echo=F, fig.width=15, fig.height=8}
car_per_day_per_cluster_total <- function(umap_fcs_dt, cell_type, cluster_names) {
  car_cluster_day_indiv_k562_total_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(car,cluster,k562,t_type,day))][,
      list(cluster, pct=N/sum(N)), by=c('car','k562','t_type','day')][,
        pct_delta := scale(pct), by=c('car','k562','t_type','day')]
  
  car_cluster_day_indiv_k562_total_pct_dt[,
  cluster := factor(cluster, levels = cluster_names)]
  
  car_cluster_day_indiv_k562_total_pct_dt$day <- factor(car_cluster_day_indiv_k562_total_pct_dt$day, levels=c('0','6','15','24','33'))
  
  min_col <- brewer.pal(name='BrBG', n=11)[2]
  
  ggplot(car_cluster_day_indiv_k562_total_pct_dt[k562=='cd19+']) + 
      geom_tile(aes(x=cluster, y=car, fill=pct_delta), color='black') +
      facet_grid(day~k562, scales='free') +
      scale_fill_distiller('',
        palette='BrBG', 
        direction=1,
        #limits=c(-2.8, 2.8), 
        na.value=min_col) +
      labs(title='CAR enrichment\nper day per cluster', y='CAR',
        subtitle='CAR enrichment in cluster\n(z-score) (summed across rows)') +
      theme_minimal() + ggtitle(paste(as.character(cell_type),' CAR enrichment\nper day per cluster')) 
}

#cd8
#car_per_day_per_cluster_total(umap_cd8_fcs_dt, "CD8", cd8_colnames)

#cd4
#car_per_day_per_cluster_total(umap_cd4_fcs_dt, "CD4", cd4_colnames)
```

#### Bar charts showing how each CAR changes cluster by day
```{r fig.width=15, fig.height=8}

car_clusters_by_day <- function(umap_fcs_dt, cell_type, cluster_names) {
  car_cluster_day_indiv_k562_bar_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(car,cluster,k562,t_type,day))][,
      list(day, pct=N/sum(N)), by=c('car','k562','t_type','cluster')][,
        pct_delta := scale(pct), by=c('car','k562','t_type','cluster')]
  
  car_cluster_day_indiv_k562_bar_pct_dt[,
  cluster := factor(cluster, levels = cluster_names)]
  
  car_cluster_day_indiv_k562_bar_pct_dt$day <- factor(car_cluster_day_indiv_k562_bar_pct_dt$day, levels=c('0','6','15','24','33'))
  
  #filter for CD19+, only CARs > 10% in a cluster in a given day
  car_cluster_day_indiv_k562_bar_pct_dt %>% 
    filter(k562=='cd19+', pct > 0.10) %>%
    ggplot() + geom_bar(aes(x=cluster, y=pct), color='black', stat='identity') +
      facet_grid(car~day, scales='free_x', space='free_x') +
      labs(title='CAR enrichment (>10%)\nper cluster per day', y='CAR',
        subtitle='CAR enrichment in cluster\n(absolute %) (summed across rows by cluster number)') + ggtitle(paste(as.character(cell_type),' CAR enrichment (>10%)\nper cluster per day'))
}

#cd8
car_clusters_by_day(umap_cd8_fcs_dt, "CD8", cd8_colnames)

#cd4
car_clusters_by_day(umap_cd4_fcs_dt, "CD4", cd4_colnames)
```

#### Bar charts showing distribution of each CAR in clusters per day 
```{r fig.width=15, fig.height=8}
car_clusters_by_square <- function(umap_fcs_dt, cell_type, cluster_names) {
  car_cluster_day_indiv_k562_total_pct_dt <- map_day_0(umap_fcs_dt)[,
      data.table(table(car,cluster,k562,t_type,day))][,
        list(cluster, pct=N/sum(N)), by=c('car','k562','t_type','day')][,
          pct_delta := scale(pct), by=c('car','k562','t_type','day')]
    
    car_cluster_day_indiv_k562_total_pct_dt[,
    cluster := factor(cluster, levels = cluster_names)]
    
    car_cluster_day_indiv_k562_total_pct_dt$day <- factor(car_cluster_day_indiv_k562_total_pct_dt$day, levels=c('0','6','15','24','33'))
  #filter for  CD19+, only CARs > 10% in a cluster in a given day
  car_cluster_day_indiv_k562_total_pct_dt %>% 
    filter(k562=='cd19+', pct>.10) %>%
    ggplot() + geom_bar(aes(x=cluster, y=pct), color='black', stat='identity') +
      facet_grid(car~day, scales='free_x', space='free_x') +
      labs(title='CAR enrichment (>10%)\nper day per cluster', y='CAR',
        subtitle='CAR enrichment in cluster\n(z-score) (summed across CAR x day (each square))') + ggtitle(paste(as.character(cell_type),' CAR enrichment (>10%)\nper day per cluster'))
}

#cd8
car_clusters_by_square(umap_cd8_fcs_dt, "CD8", cd8_colnames)

#cd4
car_clusters_by_square(umap_cd4_fcs_dt, "CD4", cd4_colnames)

```

## Grouped Plots

### CD8 Cluster Plots
```{r, fig.width=15, fig.height=30}
all_plots <- function(umap_fcs_dt, cell_type) {
  cluster_pct_vars <- c(paste0(
    names(channel_map),'_t'), 
    'FSC-A','SSC-A','donor','cd19')
  
  max_cluster <- max(as.numeric(umap_fcs_dt$cluster), na.rm=T)
  
  umap_var_melt_dt <-melt(
    umap_fcs_dt[!is.na(donor)], measure.vars= c(umap_vars,'donor', 'cd19', 'myc_t', 'gfp_t')) #add back in previously removed vars
  
  # Use z scale to plot params
  umap_var_melt_dt[, value_scaled := scale(value), by=variable]
  param_cluster_pct_dt <- umap_var_melt_dt[
    !is.na(cluster) & variable %in% cluster_pct_vars,
    list(
      clust_mean = mean(value_scaled),
      clust_sd = sd(value_scaled)),
        by=c('variable','cluster')]
  
  # Make dendrogram
  cluster_mean_cast <- dcast(
    param_cluster_pct_dt, 
    variable ~ cluster, 
    value.var='clust_mean')
  
  cluster_dendro_m <- as.matrix(cluster_mean_cast[, -c(1)])
  rownames(cluster_dendro_m) <- unlist(cluster_mean_cast[,1])
  dendro_clusters <- as.dendrogram(hclust(d = dist(x = t(cluster_dendro_m))))
  dendro_vars <- as.dendrogram(hclust(d = dist(x = cluster_dendro_m)))
  
  # Create dendrogram plot
  dendro_vars_plot <- ggdendrogram(data = dendro_vars, rotate = TRUE) + 
    theme(axis.text.y = element_text(size = 6))
  dendro_clusters_plot <- ggdendrogram(data = dendro_clusters, rotate = TRUE) + 
    theme(axis.text.y = element_text(size = 6))
  
  # column order
  cluster_order <- order.dendrogram(dendro_clusters)
  cluster_names <- colnames(cluster_dendro_m[, cluster_order])
  param_cluster_pct_dt[, cluster:= factor(cluster, levels = cluster_names)]
  
  # row order
  var_order <- order.dendrogram(dendro_vars)
  var_names <- row.names(cluster_dendro_m[var_order, ])
  param_cluster_pct_dt$variable <- factor(
      param_cluster_pct_dt$variable,
      levels = var_names)
  
   #annotate clusters as per variable values (removed gfp, myc, cd, donor)
  '%notin%' <- Negate('%in%')
  annotations <- param_cluster_pct_dt %>%
  mutate(variable=gsub("\\_t$", "", variable)) %>%
  filter(variable %notin% c('gfp','myc','cd','donor', 'cd19')) %>%
  mutate(cluster=as.character(cluster)) %>%
  group_by(cluster) %>% 
  mutate(rk = rank(-abs(clust_mean))) %>%
  filter(rk <= 3) %>% 
  arrange(cluster, rk) %>%
  mutate(variable = ifelse(clust_mean > 0, paste(variable,'+',sep = ""), paste(variable,'-',sep = ""))) %>%
  summarise(variable = paste(variable, collapse=", ")) %>% 
  data.table::transpose()
  colnames(annotations) <- as.character(unlist(annotations[1,]))
  annotations = annotations[-1, ]
  annotations <- annotations[cluster_names]
  row.names(annotations) <- NULL 
  
  # heatmap
  var_heatmap <- ggplot(param_cluster_pct_dt) + 
      geom_tile(aes(x=cluster, y=variable, fill=clust_mean), color='black') + geom_text(aes(x=cluster, y=variable, label=round(clust_mean, 2)), size=3, color='black') +
      scale_fill_distiller(palette='PiYG', limits=c(-3,3), oob=scales::squish) +
      theme_minimal() # + theme(axis.text.x = element_text(angle=90, hjust=1))

  # col dendro
  dendro_data_col <- dendro_data(dendro_clusters, type = "rectangle")
  dendro_col <- axis_canvas(var_heatmap, axis = "x") + 
      geom_segment(data = segment(dendro_data_col), 
          aes(x = x, y = y, xend = xend, yend = yend))
  plot_dendroheat <- insert_xaxis_grob(var_heatmap, 
              dendro_col, grid::unit(0.2, "null"), position = "top")
  dendro_plot <- ggdraw(plot_dendroheat) # + draw_label(label=cell_type, x = 0.95, y = 0.95, hjust = 0.95, vjust = 0.95, size=40) 
  
  # annotation table
  tt <- ttheme_default(core=list(fg_params = list(fontsize=10), colhead=list(fg_params = list(fontsize=10, parse=TRUE))))
  annotations[1,] = str_wrap(annotations[1,], 0)
  tbl <- tableGrob(annotations, rows=NULL, theme=tt)
  # tbl <- strwrap(tbl, width = 10, simplify = FALSE)
  tbl$widths <- unit(rep(1/ncol(tbl), ncol(tbl)), "npc")
  
  # CLUSTERS BY DAY
  day_cluster_total_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(day,cluster, t_type, k562))][,
      list(cluster, pct=N/sum(N)), by=c('k562','day','t_type')]
  
  day_cluster_total_pct_dt[, cluster := factor(cluster, levels = cluster_names)]
  
  clusters_by_day <- ggplot(day_cluster_total_pct_dt[k562=='cd19+'], aes(
      x=cluster, 
      y=factor(day, levels=c(0,6,15,24,33)), 
      fill=pct)) + 
    geom_tile(color='black') +
    geom_text(aes(label=round(pct*100)), size=3, color='white') +
    facet_grid(k562+t_type~., scales='free') +
    scale_fill_viridis_b('',direction=1, limits=c(0, 0.6)) +
    labs(title='Percent of cells in each day by cluster', y='Day',
      subtitle='(rows sum to 100%)') +
    theme_minimal() + ggtitle(paste('Percent of ',as.character(cell_type),' cells in each day by cluster'))

  # CARs BY CLUSTER/DAY
  car_cluster_day_indiv_k562_total_pct_dt <- map_day_0(umap_fcs_dt)[,
    data.table(table(car,cluster,k562,t_type,day))][,
      list(cluster, pct=N/sum(N)), by=c('car','k562','t_type','day')][,
        pct_delta := scale(pct), by=c('car','k562','t_type','day')]
  
  car_cluster_day_indiv_k562_total_pct_dt[,
  cluster := factor(cluster, levels = cluster_names)]
  
  car_cluster_day_indiv_k562_total_pct_dt$day <- factor(car_cluster_day_indiv_k562_total_pct_dt$day, levels=c('0','6','15','24','33'))
  
  min_col <- brewer.pal(name='BrBG', n=11)[2]
  
  cars_by_cluster <- ggplot(car_cluster_day_indiv_k562_total_pct_dt[k562=='cd19+']) + 
      geom_tile(aes(x=cluster, y=car, fill=pct_delta), color='black') +
      facet_grid(day~k562, scales='free') +
      scale_fill_distiller('',
        palette='BrBG', 
        direction=1,
        #limits=c(-2.8, 2.8), 
        na.value=min_col) +
      labs(title='CAR enrichment\nper day per cluster', y='CAR',
        subtitle='CAR enrichment in cluster\n(z-score) (summed across rows)') +
      theme_minimal() + ggtitle(paste(as.character(cell_type),' CAR enrichment\nper day per cluster')) 
  
  # print all plots and table
  plot_grid(dendro_plot, clusters_by_day, cars_by_cluster, tbl, labels = c('A', 'B', 'C', 'D'), nrow=4, rel_heights = c(1/8, 1/8, 1/2, 1/8))
} 

all_plots(umap_cd8_fcs_dt, 'CD8')

umap_cd8_fcs_dt %>% head()

# manually determined 
cd8_diff_by_cluster <- data.frame("cluster" = c(1:12), "cell_type" = c('TEM', 'TSCM', 'TCM', 'TSCM', 'TEM', 'TTM', 'TEM', 'TTM', 'TEM', 'TEM', 'TN', 'TN'))
cd8_diff_by_cluster$cluster <- as.factor(cd8_diff_by_cluster$cluster)
umap_cd8_fcs_dt <- left_join(umap_cd8_fcs_dt, cd8_diff_by_cluster, by='cluster')

```

### CD4 Cluster Plots
```{r, fig.width=15, fig.height=30}
cd4_diff_by_cluster <- data.frame("cluster" = c(1:14), "cell_type" = c('TEM', 'TEM', 'TEM', 'TEM', 'TCM', 'TN', 'TN', 'TCM', 'TSCM', 'TEM', 'TCM', 'TCM', 'TSCM','TEM'))
cd4_diff_by_cluster$cluster <- as.factor(cd4_diff_by_cluster$cluster)
umap_cd4_fcs_dt <- left_join(umap_cd4_fcs_dt, cd4_diff_by_cluster, by='cluster')

all_plots(umap_cd4_fcs_dt, 'CD4')
```
## Temporal UMAP Visualizations 


### Visualize parameters on umap
```{r}
viz.umap <- function(dat,dr,param.name,limits=NULL){
  ColVal <- pull(dat %>% select(param.name))
  if(is.null(limits)){
    Lim <- quantile(ColVal,probs=seq(0,1,0.01))[c(2,100)]
    p <- ggplot(dat, aes(x = umap_1, y =umap_2)) +geom_point(aes(color = ColVal), size=0.1)+theme_classic()+scale_color_distiller(name=param.name, palette = "RdYlBu", limits=Lim, oob=squish)+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank())+ggtitle(param.name)
  } else {
    p <- ggplot(dat, aes(x = umap_1, y = umap_2)) +geom_point(aes(color = ColVal), size=0.1)+theme_classic()+scale_color_distiller(name=param.name, palette = "RdYlBu", limits=c(limits[1],limits[2]), oob=squish)+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank())+ggtitle(param.name)
  }
  p
}
```

#### CD8 markers 
```{r fig.width=15, fig.height=11}
param_test <- colnames(umap_cd8_fcs_dt)[c(7, 9:13)]

as.data.frame(umap_cd8_fcs_dt)[, param_test]

#function to visualize diffusion map
visualize_params_umap <- function(df) {
  parameters <- colnames(df)[c(7, 9:13)]
  for (i in 1:length(parameters)) {
  p <- viz.umap(dat=df, param.name=parameters[i])
  print(p)
  }
}

umap_marker_heatmap <- function(df) {
plot_grid(viz.umap(dat=df, param.name='ccr7_t'), viz.umap(dat=df, param.name='cd27_t'), viz.umap(dat=df, param.name='cd45ra_t'), viz.umap(dat=df, param.name='cd45ro_t'), viz.umap(dat=df, param.name='cd62l_t'), viz.umap(dat=df, param.name='cd95_t'), row=2)
}

umap_marker_heatmap(umap_cd8_fcs_dt)
```
#### CD4 markers
```{r fig.width=15, fig.height=11}
umap_marker_heatmap(umap_cd4_fcs_dt)
```


#### visualize leiden clusters on umap
```{r}
visualize_clusters_umap <- function(df) {
  label_clusters_umap <- df %>% group_by(cluster, cell_type) %>% select(umap_1, umap_2) %>% summarize_all(mean)
  ggplot(df, aes(x=umap_1, y=umap_2, color=as.factor(cluster)))+geom_point(size=0.1)+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank())+geom_label_repel(aes(label=cell_type), data=label_clusters_umap)+guides(colour=FALSE)+ggtitle('Leiden clusters')
}
  
visualize_clusters_umap(umap_cd8_fcs_dt)
visualize_clusters_umap(umap_cd4_fcs_dt)
```

#### visualize days UMAP
```{r}
visualize_days_umap <- function(df) {
  label_days_umap <- df %>% group_by(day) %>% select(umap_1, umap_2) %>% summarize_all(mean)
  ggplot(df, aes(x=umap_1, y=umap_2, color=as.factor(day)))+geom_point(size=0.1)+theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank())+geom_label_repel(aes(label=day), data=label_days_umap)+guides(colour=FALSE)+ggtitle('UMAP by Day') + scale_color_viridis(discrete=TRUE)
}

#cd8
visualize_days_umap(umap_cd8_fcs_dt)

#cd4
visualize_days_umap(umap_cd4_fcs_dt)
```

#### visualize CARS UMAP
```{r fig.width=15, fig.height=11}
visualize_cars_umap <- function(df) {
  label_cars_umap <- df %>% group_by(car) %>% select(umap_1, umap_2) %>% summarize_all(mean)
  ggplot(df, aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = ..level..)) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank()) + guides(colour = guide_legend(override.aes = list(size=5)))
}


#ALL CARS
visualize_all_cars_by_day_umap <- function(df) {
  plot_grid(df %>% filter(day==0) %>% visualize_cars_umap()
, df %>% filter(day==6) %>% visualize_cars_umap(), df %>% filter(day==15) %>% visualize_cars_umap(), df %>% filter(day==24) %>% visualize_cars_umap(), df %>% filter(day==33) %>% visualize_cars_umap(), labels = c('Day 0', 'Day 6', 'Day 15', 'Day 24', 'Day 33'), nrow=3)
}

#ALL CARS - CD8
visualize_all_cars_by_day_umap(umap_cd8_fcs_dt)

#ALL CARS - CD4
visualize_all_cars_by_day_umap(umap_cd4_fcs_dt)
```
#### Contour plot - visualize CARS by day, cluster
##### CD8
```{r fig.width=15, fig.height=11}
'%notin%' <- Negate('%in%')

#cars x day contour
umap_cd8_fcs_dt %>% ggplot(., aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = stat(nlevel)), contour_var = "ndensity") + facet_grid(car ~ day) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank()) + ggtitle("CAR Densities by Day")

# clusters x day contour
umap_cd8_fcs_dt %>% ggplot(., aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = stat(nlevel)), contour_var = "ndensity") + facet_grid(cluster ~ day) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank()) + ggtitle("Cluster Densities by Day")


# clusters by car contour
umap_cd8_fcs_dt %>% ggplot(., aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = stat(nlevel)), contour_var = "ndensity") + facet_grid(cluster ~ car) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank())+ ggtitle("CAR Densities by Cluster")
```

##### CD4 
```{r fig.width=15, fig.height=11}
#cars x day contour
umap_cd4_fcs_dt %>% ggplot(., aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = stat(nlevel)), contour_var = "ndensity") + facet_grid(car ~ day) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank()) + ggtitle("CAR Densities by Day")

# clusters x day contour
umap_cd4_fcs_dt %>% ggplot(., aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = stat(nlevel)), contour_var = "ndensity") + facet_grid(cluster ~ day) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank()) + ggtitle("Cluster Densities by Day")


# clusters by car contour
umap_cd4_fcs_dt %>% ggplot(., aes(x=umap_1, y=umap_2))+geom_point(size=0.1, color= "#DCDCDC") + geom_density2d(aes(colour = stat(nlevel)), contour_var = "ndensity") + facet_grid(cluster ~ car) + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.title = element_blank())+ ggtitle("CAR Densities by Cluster")


```

#### frequencies (checking for donor batch effect)
```{r}
#cd8
cluster_by_donor <- function(df, num_clusters){
  counts <- as.data.frame.matrix(table(df$donor,df$cluster))
  counts$donor <- c(1:2)
  counts %>% group_by(donor) %>% gather(1:num_clusters, key=cluster, value=samples) %>% ggplot(., aes(x="",y=samples, fill=factor(donor))) + facet_grid(.~cluster) + geom_bar(width = 1, stat='identity', position=position_dodge())
}

cluster_by_donor(umap_cd8_fcs_dt, 12) #imbalances in clusters 1,10,2,4,7,9

#cd4
cluster_by_donor(umap_cd4_fcs_dt, 14) #imbalances in clusters 14,7

```

#### CD8 correlation
```{r fig.width=15, fig.height=11}
library(Hmisc)
library(ggcorrplot) 
#corr plot
#biomarker_values <- umap_cd8_fcs_dt[,c(7,9:13)]
#corr <- round(cor(biomarker_values), 2)
#p.mat <- cor_pmat(biomarker_values) #pvals

# visualize - positive corrs in red, neg in blue Color intensity and size are proportion to corr coefficients (legend).
#ggcorrplot(corr, method = "circle", hc.order = TRUE, type = "upper", outline.col = "white",ggtheme = ggplot2::theme_gray, colors = c("#6D9EC1", "white", "#E46726"), p.mat = p.mat)

biomarker_corr <- function(df) {
  biomarkers <- df %>% 
  select('ccr7_t', 'cd27_t', 'cd45ra_t', 'cd45ro_t', 'cd62l_t', 'cd95_t') 
  corr <- round(cor(biomarkers),2)
  p.mat <- cor_pmat(biomarkers)
  ggcorrplot(corr, method = "circle", type = "upper", outline.col = "white", p.mat = p.mat, ggtheme= ggplot2::theme_gray, colors = c("#6D9EC1", "white", "#E46726"))
}

biomarker_corr_by_day <- function(df) {
  plot_grid(df %>% filter(day==0) %>% biomarker_corr()
, df %>% filter(day==6) %>% biomarker_corr(), df %>% filter(day==15) %>% biomarker_corr(), df %>% filter(day==24) %>% biomarker_corr(), df %>% filter(day==33) %>% biomarker_corr(), labels = c('Day 0', 'Day 6', 'Day 15', 'Day 24', 'Day 33'), nrow=3)
}

#cd8
biomarker_corr_by_day(umap_cd8_fcs_dt)
```

#### CD4 correlation
```{r fig.width=15, fig.height=11}
biomarker_corr_by_day(umap_cd4_fcs_dt)
```







